---
title: "Human T cell data - compare DEGs between CD4, CD8, MAIT, NKT, gdT lineages"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


## 1. IMPORT

```{r import-librairies, message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(Seurat)
library(RColorBrewer)
library(DESeq2)
library(pheatmap)
library(ggrepel)
library(cowplot)
# library(clusterProfiler)
# library(enrichplot)
```

```{r import-data}
# Import seurat object
seur.human <- readRDS("~/Projects/HumanThymusProject/data/raw_data/human_data/filtered_seurat_harmony.11.22.22.RDS")
seur.filt <- readRDS("~/Projects/HumanThymusProject/data/CLUSTER/seurat-integration/input/seur_filt_22-12-19.rds")

# Sanity check
print(seur.human) # 81,378 cells (it's the whole seurat object)

# Quick visualization
DimPlot(seur.human, reduction="UMAP_50", group.by="new_clusters")
```




## 2. FINDMARKERS

### 2.1. Identify CD4, CD8, MAIT, NKT, gdT cells
```{r identify-cell-types}
# Get vectors of cell names
cells.cd4thy <- rownames(seur.human@meta.data[seur.human@meta.data$group.ident=="CD4_Thymus",])
cells.cd8thy <- rownames(seur.human@meta.data[seur.human@meta.data$group.ident=="CD8_Thymus",])
cells.maithy <- rownames(seur.human@meta.data[seur.human@meta.data$group.ident=="MAIT_Thymus",])
cells.nkthy <- rownames(seur.human@meta.data[seur.human@meta.data$group.ident=="NKT_Thymus",])
cells.gdthy <- rownames(seur.human@meta.data[seur.human@meta.data$group.ident=="GD_Thymus",])

# sanity check
# table(seur.human@meta.data$group.ident)
# length(cells.cd4thy)
# length(cells.cd8thy)
# length(cells.maithy)
# length(cells.nkthy)
# length(cells.gdthy)
```


### 2.2. Calculate DEGs
```{r deg-thy-cd4vscd8}
# DEGs on SCTransformed data
FindMarkers(seur.human, ident.1="CD4_Thymus", ident.2="CD8_Thymus", logfc.threshold = 0.25)

# SCTransform vignette
seur.human.sctprep <- PrepSCTFindMarkers(seur.human)
seur.human.sctprep <- subset(seur.human.sctprep, idents = c("CD4_Thymus", "CD8_Thymus"))
degs.sct <- FindMarkers(seur.human.sctprep, assay = "SCT", ident.1 = "CD4_Thymus", ident.2 = "CD8_Thymus", recorrect_umi = FALSE)
degs.sct %>%
  arrange(avg_log2FC)

# Try on log normalized data
counts <- GetAssayData(object = seur.human, slot = "counts", assay="RNA")
metadata <- seur.human@meta.data
initseurat <- CreateSeuratObject(counts = counts, meta.data = metadata, min.cells=1) # keep only genes expressed in at least 1 cell
seur.norm <- NormalizeData(object=initseurat, normalization.method = "LogNormalize")
seur.norm <- ScaleData(object=seur.norm, features=rownames(seur.norm))
Idents(seur.norm) <- "group.ident"
levels(Idents(seur.norm))
degs.lognorm <- FindMarkers(seur.norm, ident.1="CD4_Thymus", ident.2="CD8_Thymus", logfc.threshold = 0.4)
degs.lognorm %>%
  arrange(avg_log2FC)

# Look at other comparisons
FindMarkers(seur.norm, ident.1="CD4_Thymus", ident.2="NKT_Thymus", logfc.threshold = 0.4) %>% arrange(avg_log2FC)
FindMarkers(seur.norm, ident.1="CD4_Thymus", ident.2="MAIT_Thymus", logfc.threshold = 0.4) %>% arrange(avg_log2FC)
FindMarkers(seur.norm, ident.1="CD8_Thymus", ident.2="NKT_Thymus", logfc.threshold = 0.4) %>% arrange(avg_log2FC)
FindMarkers(seur.norm, ident.1="CD8_Thymus", ident.2="MAIT_Thymus", logfc.threshold = 0.4) %>% arrange(avg_log2FC)
FindMarkers(seur.norm, ident.1="NKT_Thymus", ident.2="MAIT_Thymus", logfc.threshold = 0.4) %>% arrange(avg_log2FC)

# Look at DEGs across batches from the same cell type
test <- subset(seur.norm, subset= group.ident=="CD8_Thymus")
Idents(test) <- "Batch"
levels(Idents(test))
FindMarkers(test, ident.1="B", ident.2="C", logfc.threshold = 0.4) %>% arrange(avg_log2FC)
```

```{r deg-thy-cd4vsrest}
# Thymic CD4 T cells vs rest
seur.thymus <- subset(seur.norm, subset= Tissue=="Thymus")
degs.lognorm.cd4 <- FindMarkers(seur.thymus, ident.1="CD4_Thymus", logfc.threshold = 0.4)
degs.lognorm.cd4 %>%
  arrange(avg_log2FC)


# Thymic CD8 T cells vs rest
degs.lognorm.cd8 <- FindMarkers(seur.thymus, ident.1="CD8_Thymus", logfc.threshold = 0.4)
degs.lognorm.cd8 %>%
  arrange(avg_log2FC) # 274


# Thymic NKT T cells vs rest
degs.lognorm.cd8 <- FindMarkers(seur.thymus, ident.1="NKT_Thymus", logfc.threshold = 0.4)
degs.lognorm.nkt %>%
  arrange(avg_log2FC)
```




## 3.  PSEUDOBULK DESEQ2

### 3.1. Prepare data
Following the tutorial from : https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html

```{r deseq2}
## ****************
## IDENTIFY GROUPS FOR DE ANALYSIS
# Create SingleCellExperiment object
counts <- seur.filt@assays$RNA@counts
metadata <- seur.filt@meta.data

# Subset metadata only to information we're interested in
metadata <- metadata[,c("Tissue", "cell.ident", "Batch", "old_clusters")]
colnames(metadata) <- c("tissue_id", "cell_id", "batch_id", "cluster_id")
print(metadata) # sanity check

# Create a "groups" df that will define how we aggregate counts
groups <- metadata %>%
  # ---
  # EXAMPLE 1: Compare thymic CD4 vs thymic CD8 cells (all clusters, but separate batches)
  # filter(tissue_id =="Thymus" & cell_id %in% c("CD4", "CD8")) %>%
  # select(cell_id, batch_id) # keep only relevant columns for grouping
  # ---
  # EXAMPLE 2: Compare peripheral CD4 vs peripheral CD8 cells in cluster 12 (separate batches)
  filter(tissue_id=="PBMC" & cell_id %in% c("CD4", "CD8") & cluster_id=="12") %>%
  select(cell_id, batch_id) # keep only relevant columns for grouping
  # ---
  # EXAMPLE 3: Compare thymic NKT vs peripheral NKT cells in cluster 12 (separate batches)
  # filter(cell_id=="NKT" & cluster_id=="12") %>%
  # select(tissue_id, batch_id) # keep only relevant columns for grouping
print(groups) # sanity check

## ****************
## AGGREGATE COUNTS
# Subset to keep only cells of interest (defined in "groups") & put cells in rows (genes as columns)
count.agg <- t(counts[, which(colnames(counts) %in% rownames(groups))])
nrow(groups) == nrow(count.agg) # verify nb of cells in "groups" (rows) is same nb of cells in "count.agg" (rows)

# Aggregate counts based on columns in "groups"
count.agg <- Matrix.utils::aggregate.Matrix(count.agg, groupings = groups, fun = "sum")
# dim(count.agg)
# count.agg[1:6, 1:6]
# table(colSums(count.agg) == 0) # check if any gene has total count of 0

# Final counts for DESeq (groups as columns, genes as rows)
counts.deseq <- data.frame(t(count.agg))
counts.deseq <- counts.deseq[rowSums(counts.deseq)!=0,] # remove genes that have total count of 0


## ****************
## PREPARE METADATA DF FOR DESEQ
# Have metadata df where rownames are the colnames of "counts.deseq"
metadf.deseq <- groups %>%
  distinct() %>%
  unite(sample_id, remove=FALSE)
rownames(metadf.deseq) <- metadf.deseq$sample_id

# Check that the row names of metadf.deseq are the same as the column names of counts.deseq in order to use as input to DESeq2
metadf.deseq <- metadf.deseq[match(colnames(counts.deseq), rownames(metadf.deseq)),]
all(rownames(metadf.deseq) == colnames(counts.deseq))
```




### 3.2. Differential expression analysis


Now we can create our DESeq2 object to prepare to run the DE analysis. We need to include the counts, metadata, and design formula for our comparison of interest. In the design formula we should also include any other columns in the metadata for which we want to regress out the variation (e.g. batch, sex, age, etc.). We only have our comparison of interest, which is stored as the group_id in our metadata data frame.

```{r create-deseq2-object}
# Create DESeq2 object
dds <- DESeqDataSetFromMatrix(counts.deseq, 
                              colData = metadf.deseq, 
                              design = ~ cell_id + batch_id) # TO MODIFY BASED ON WHAT YOU WANT TO COMPARE (keep batch_id though)
```


The next step in the DESeq2 workflow is QC, which includes sample-level and gene-level steps to perform QC checks on the count data to help us ensure that the samples/replicates look good.
To explore the similarity of our samples, we will be performing sample-level QC using Principal Component Analysis (PCA) and hierarchical clustering methods. Sample-level QC allows us to see how well our replicates cluster together, as well as, observe whether our experimental condition represents the major source of variation in the data. Performing sample-level QC can also identify any sample outliers, which may need to be explored further to determine whether they need to be removed prior to DE analysis.
When using these unsupervised clustering methods, normalization and log2-transformation of the counts improves the distances/clustering for visualization. DESeq2 uses median of ratios method for count normalization and a regularized log transform (rlog) of the normalized counts for sample-level QC as it moderates the variance across the mean, improving the clustering.


```{r deseq2-qc}
# Transform counts for data visualization
rld <- rlog(dds, blind=TRUE)
head(assay(rld))

# Plot PCA
DESeq2::plotPCA(rld, intgroup = "sample_id")
DESeq2::plotPCA(rld, intgroup = "cell_id")
# PC1 separates the CD4 & CD8

# Correct for batch effect, and then run PCA
counts_corrected <- limma::removeBatchEffect(assay(rld), metadf.deseq$batch_id)
rv <- matrixStats::rowVars(counts_corrected) # variance of each gene
select_rv <- order(rv, decreasing = TRUE)[seq_len(500)] # get the positions of the top 500 most variable genes?...
pca <- prcomp(t(counts_corrected[select_rv,])) # run pca on top 500 HVG
percentVar <- pca$sdev^2/sum(pca$sdev^2)
counts_pca <- data.frame(PC1 = pca$x[, 1], PC2 = pca$x[, 2],
                         batch = metadf.deseq$batch_id,
                         condition = metadf.deseq$cell_id) # MODIFY "condition" BASED ON WHAT YOU ARE COMPARING
# Plot
ggplot(counts_pca, aes(x = PC1, y = PC2, color = condition, shape=batch)) +
  geom_point(size = 4) +
  xlab(paste0("PC1: ", round(percentVar[1] * 100), "% variance")) +
  ylab(paste0("PC2: ", round(percentVar[2] * 100), "% variance")) +
  coord_fixed(expand=TRUE) +
  scale_y_continuous(expand = c(0.1,0.1))


# Compute pairwise correlation values on the batch-effect-corrected matrix
rld_cor <- cor(counts_corrected[select_rv,])
# Plot heatmap
pheatmap(rld_cor, annotation = metadf.deseq[, c("cell_id"), drop=F])
```


Differential expression analysis with DESeq2 involves multiple steps as displayed in the flowchart below in blue. Briefly, DESeq2 will model the raw counts, using normalization factors (size factors) to account for differences in library depth. Then, it will estimate the gene-wise dispersions and shrink these estimates to generate more accurate estimates of dispersion to model the counts. Finally, DESeq2 will fit the negative binomial model and perform hypothesis testing using the Wald test or Likelihood Ratio Test.
All of these steps are performed by running the single DESeq() function on our DESeq2 object created earlier.

```{r, fig.width=5, fig.height=5}
# Run DESeq2 differential expression analysis
dds <- DESeq(dds)

# Plot dispersion estimates
# jpeg("~/Projects/HumanThymusProject/data/human-thymus/HumanData_06_DEGs-btw-Tlineages/dispersion_estimate.jpeg", width=1500, height=1500, res=200)
plotDispEsts(dds)
# dev.off()

# Output results of Wald test for contrast for CD4 vs CD8
contrast <- c("cell_id", "CD4", "CD8") # MODIFY BASED ON WHAT YOU ARE COMPARING
res <- results(dds, 
               contrast = contrast,
               alpha = 0.05)
res <- lfcShrink(dds,
                 type="ashr",
                 contrast = contrast,
                 res=res)
res # condition CD4 vs CD8 (high log2FC means upregulated in CD4)

# Turn the results object into a tibble for use with tidyverse functions
result.df <- res %>%
  data.frame() %>%
  rownames_to_column(var="gene") %>%
  as_tibble()


## KEEP ONLY SIGNIFICANT GENES ##
# Set threshold
padj_cutoff <- 0.05

# Subset the significant results
result.sig <- dplyr::filter(result.df, padj < padj_cutoff) %>%
  dplyr::arrange(padj)

# Check significant genes output
table(result.sig$log2FoldChange > 0)
result.sig %>%
  dplyr::arrange(log2FoldChange)
# Sanity check
# result.sig %>%
#   filter(gene %in% c("RUNX3", "ZBTB7B"))
```


### 3.3. Heatmap

```{r heatmap}
# Extract normalized counts for only the significant genes
counts.norm <- counts(dds, normalized = TRUE)
counts.norm.sig <- data.frame(counts.norm) %>%
  rownames_to_column(var = "gene") %>%
  dplyr::filter(gene %in% result.sig$gene)
rownames(counts.norm.sig) <- counts.norm.sig$gene
        
# Set a color palette
heat_colors <- brewer.pal(6, "YlOrRd")

# Run pheatmap using the metadata data frame for the annotation
path.plots <- "~/Projects/HumanThymusProject/data/human-thymus/HumanData_06_DEGs-btw-Tlineages"
# jpeg(file.path(path.plots, "heatmap_CD4CD8.jpeg"), height = 2000, width = 2000, res = 300)
pheatmap(counts.norm.sig[,-1], 
    color = heat_colors,
    cluster_rows = T,
    cluster_cols = T,
    show_rownames = T,
    fontsize_row = 2,
    annotation = metadf.deseq[, c("batch_id", "cell_id")], # MODIFY BASED ON WHAT YOU ARE COMPARING
    border_color = NA,
    fontsize = 10,
    main= "Thymic cells",
    scale = "row",
    height = 20)
# dev.off()
```


### 3.4. Volcano plot

```{r volcano, fig.height=5, fig.width=5}
## Obtain logical vector where TRUE values denote padj values < 0.05 and fold change > 1.5 in either direction
results.df.thresh <- result.df %>%
  mutate(threshold = padj < 0.05 & abs(log2FoldChange) >= 1)

# Genes to highlight on the volcano plot
interesting_genes <- result.df %>%
  # filter(gene %in% c("RUNX3", "ZBTB7B"))
  filter(abs(log2FoldChange) >= 3)
                  
## Volcano plot
ggplot(results.df.thresh, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(colour = threshold), alpha=0.5) +
  geom_text_repel(data = interesting_genes,
                   aes(label = gene),
                   force = 2, max.overlaps=30,
                   nudge_y = 1) +
  scale_color_manual(values=c("grey", "orange"))+
  labs(x="log2 fold change", y="-log10 adjusted p-value", title="Thymic CD4 vs CD8 cells") +
  theme_cowplot()+
  theme(legend.position = "none",
          plot.title = element_text(size = rel(1.5), hjust = 0.5),
          axis.title = element_text(size = rel(1.25))) 
# ggsave(file.path(path.plots, "volcano_CD4CD8.jpeg"), width=7, height=7, bg="white")
```
