---
title: "Human data - evaluate clusters based on k.param, # PCs and resolution"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(root.dir = "~/Projects/HumanThymusProject/scripts/human-thymus/")
```


## 1. Import

```{r import-librairies, message=FALSE}
# Libraries
library(ggplot2)
library(cowplot)
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
library(scclusteval)
```




## 2. Trying scClustEval on 5k PBMC data

Take a quick look at the PBMC 5k data
```{r pbmc1}
# Import data
pbmc <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/pbmc/pbmc_5k_v3_label_transfered_from_10k.rds")
subsample_idents<- readRDS("../../data/human-thymus/HumanData_05_scClustEval/pbmc/output_snakemake/gather_subsample.rds")
fullsample_idents<- readRDS("../../data/human-thymus/HumanData_05_scClustEval/pbmc/output_snakemake/gather_full_sample.rds")

# pbmc@meta.data
# DimPlot(pbmc)
```

Explore full dataset
```{r pbmc2}
## how many PCs to include
ElbowPlot(pbmc, ndims = 40)

# a tibble with a list column
fullsample_idents
# fullsample_idents[1,]$original_ident_full[[1]] # original_ident_full is the vector of cells with the cluster number they are assigned to.

## how many clusters for each different comibination of parameter set?
fullsample_idents %>%
  mutate(cluster_number = purrr::map_dbl(original_ident_full, ~length(unique(.x))))

###_________________________________________
# HEATMAPS
# what's the relationship of clusters between k_param 8, 20, 100, with same pc=20 and resolution = 0.6

# Get the row numbers
fullsample_idents %>% mutate(id = row_number()) %>%
  filter(pc == 20, resolution == 0.6, k_param %in% c(8, 20, 100))

## x-axis is k_param = 20, and y-axis is k_param = 8
PairWiseJaccardSetsHeatmap(fullsample_idents$original_ident_full[[2]],
                           fullsample_idents$original_ident_full[[11]],
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = T, cluster_column =T)

## x-axis is k_param = 100, and y-axis is k_param = 8
PairWiseJaccardSetsHeatmap(fullsample_idents$original_ident_full[[2]],
                           fullsample_idents$original_ident_full[[29]],
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = T, cluster_column =T)

# The k.param in the k-nearest neighbor algorithm after which a SNN graph is constructed. This parameter determines the resolution of the clustering where a bigger k yields a more interconnected graph and bigger clusters. We see if we increase the k param to 100, we get fewer number of total number of clusters.

###_________________________________________
# DIMENSION REDUCTION
# Let’s check how the clusters are splitting when we increase the k.param.
k8_ident <- fullsample_idents %>%
  filter(pc == 20, resolution == 0.6, k_param == 8)  %>%
  pull(original_ident_full) %>%
  `[[`(1)
pbmc <- AddMetaData(pbmc, metadata = k8_ident, col.name = "res0.6_k8")

k100_ident <- fullsample_idents %>%
  filter(pc == 20, resolution == 0.6, k_param == 100)  %>%
  pull(original_ident_full) %>%
  `[[`(1)
pbmc <- AddMetaData(pbmc, metadata = k100_ident, col.name = "res0.6_k100")

DimPlot(pbmc, reduction = "umap", label = TRUE, group.by = "res0.6_k8", repel = TRUE) + ggtitle("k_param 8") + NoLegend() |
  DimPlot(pbmc, reduction = "umap", label = TRUE, group.by = "res0.6_k100", repel = TRUE) + ggtitle("k_param 100") + NoLegend()

# Check the clusters vs the predicted cell types at k = 8
PairWiseJaccardSetsHeatmap(set_names(pbmc@meta.data$res0.6_k8, nm=colnames(pbmc)),
                           set_names(pbmc@meta.data$predicted.id, nm=colnames(pbmc)),
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
PairWiseJaccardSetsHeatmap(set_names(pbmc@meta.data$res0.6_k100, nm=colnames(pbmc)),
                           set_names(pbmc@meta.data$predicted.id, nm=colnames(pbmc)),
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
# when the k_param is big (100), many cell types are merged together.


###_________________________________________
# SILHOUETTE PLOTS
## change idents to cluster id when k is 8
Idents(pbmc)<- pbmc@meta.data$res0.6_k8
silhouette_scores<- CalculateSilhouette(pbmc, dims = 1:20)
sil_p1<- SilhouetteRainCloudPlot(silhouette_scores) + ggtitle("k_param 8")

## check silhouette score when k is 100
Idents(pbmc)<- pbmc@meta.data$res0.6_k100
silhouette_scores<- CalculateSilhouette(pbmc, dims = 1:20)
sil_p2<- SilhouetteRainCloudPlot(silhouette_scores) + ggtitle("k_param 100")

sil_p1 / sil_p2

# cluster 5 (k=100) split into clusters 7 and 12 (k=8)
# silhouette width for cluster 5 (k=100) is higher than clusters 7 and 12 (k=8), suggesting that k=100 is a better option?
```




## 3. Define functions

```{r sanity-check}
# **********************************************************************************************************

sanitycheck <- function(seuratobject, fullsample_idents, pc_nb=15, res=0.7, knn=20){
  cat("Please provide seurat object with already UMAP \n")
  params <- paste0("PC=", pc_nb[1], " | resolution=", res[1], " | k.param=", knn[1])
  cat("Parameters:", params, "\n")
  
  # ++++++
  # Perform clustering on computer
  # print(colnames(seuratobject@meta.data))
  seuratobject <- FindNeighbors(object = seuratobject, dims = 1:pc_nb, k.param=knn, nn.eps=0,
                                #assay = "SCT",
                                reduction = "harmony", verbose=F, force.recalc=T) %>%
    FindClusters(resolution = res, n.start=100, verbose=F) #random.seed=42)
  # cat("\nPost-clustering:\n")
  # print(colnames(seuratobject@meta.data))
  p1 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by="seurat_clusters") + ggtitle("Local computer")
  
  # ++++++
  # Get clusters from scClustEval
  ident <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb, resolution == res, k_param == knn)  %>%
    pull(original_ident_full) %>%
    `[[`(1)
  seuratobject <- AddMetaData(seuratobject, metadata = ident, col.name = "scclusteval_clusters")
  p2 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by = "scclusteval_clusters", repel = TRUE) +
    ggtitle("scClustEval")
  
  # ++++++
  # Combine plots
  title <- ggdraw() + 
    draw_label(params, fontface = 'bold', x = 0, hjust = 0) +
    theme(plot.margin = margin(0, 0, 0, 7)) # add margin on the left of the drawing canvas, so title is aligned with left edge of first plot
  plot_grid(
    title, plot_grid(p1, p2, ncol=2),
    ncol = 1,
    rel_heights = c(0.1, 1) # rel_heights values control vertical title margins
  )
}
```



```{r seurat-preprocessing}
#___________________________
# FUNCTION TO PREPROCESS DATA AND GET UMAP ON SEPARATE T CELL LINEAGES
Preprocess <- function(seuratobject, SCTsplit=T, res=0.5, harmony_vars= c("Method", "Batch"), printComputation=F){
  
  # Extract the counts data (from RNA assay) and metadata
  counts <- GetAssayData(object = seuratobject, slot = "counts", assay="RNA")
  metadata <- seuratobject@meta.data
  # Create new seurat object with only the counts and the "cleaned" metadata
  initseurat <- CreateSeuratObject(counts = counts, meta.data = metadata, min.cells=20) # keep only genes expressed in at least 20 cells
  
  cat("Initial seurat object:\n")
  print(initseurat) # this is for sanity check
  
  if(SCTsplit==T){
    cat("\n1-Perform SCTransform on split data\n")
    # Normalize with SCTransform by splitting object (per batch)
    seur.list <- SplitObject(initseurat, split.by = "orig.ident")
    for(i in 1:length(seur.list)){
      cat("...SCTransform on", names(seur.list)[i], "\n")
      seur.list[[i]] <-  SCTransform(seur.list[[i]], vars.to.regress = 'percent.mt', verbose=printComputation, seed.use = 1448145)
    }
    # seur.list <- lapply(X = seur.list, FUN = SCTransform, vars.to.regress="percent.mt", method="glmGamPois") # not working
    seur.SCT <- merge(seur.list[[1]], y = unlist(seur.list[-1], use.names=FALSE), merge.data = TRUE)
    sct.hvg  <- SelectIntegrationFeatures(object.list = seur.list, nfeatures = 3000)
    VariableFeatures(seur.SCT) <- sct.hvg
  }
  
  else if(SCTsplit==F){
    cat("\n1-Perform SCTransform on merged data")
    # Normalize with SCTransform (without splitting object)
    seur.SCT <- SCTransform(initseurat,
                            assay="RNA",
                            new.assay.name="SCT",
                            vars.to.regress = 'percent.mt',
                            seed.use = 1448145,
                            verbose=printComputation)
  }
  
  # Run PCA (don't scale data when using SCTransform)
  cat("\n2-Run PCA pre-integration")
  seur.SCT <- RunPCA(object = seur.SCT,
                     assay = "SCT",
                     npcs = 50,
                     seed.use=42,
                     verbose=printComputation)
  p1 <- DimPlot(seur.SCT, reduction = "pca", group.by = "orig.ident") + ggtitle("PCA pre-integration")
  ElbowPlot(seur.SCT)
  cat("\nSeurat object pre-integration:\n")
  print(seur.SCT)
  
  # Run Harmony for integration
  cat("\n3-Run Harmony\n")
  seur.harmony <- RunHarmony(object = seur.SCT,
                             assay.use = "SCT",
                             reduction = "pca",
                             dims.use = 1:50,
                             max.iter.harmony=30,
                             group.by.vars = harmony_vars,
                             plot_convergence = printComputation)
  cat("\nSeurat object post-integration:\n")
  print(seur.harmony)
  cat("\n4-Run UMAP\n")
  seur.harmony <- RunUMAP(object = seur.harmony, dims = 1:15, assay = "SCT", reduction = "harmony", verbose=printComputation, seed.use=42)
  # seur.harmony <- FindNeighbors(object = seur.harmony, dims = 1:15, assay = "SCT", reduction = "harmony", verbose=printComputation) %>%
  #   FindClusters(resolution = res, verbose=printComputation, random.seed=42)
  
  # Show PCA after integration
  p2 <- DimPlot(seur.harmony, reduction = "harmony", group.by = "orig.ident") + ggtitle("PCA post-integration")
  
  print(p1 | p2)
  
  return(seur.harmony)
}
```

```{r function-full-data}
#___________________________
# FUNCTION TO LOOK AT UMAP WITH 3 DIFFERENT PARAMETER SETS
paramUMAP <- function(objectoplot, fullsample_idents=NULL, objectwithclusters=NULL, pc_nb=15, res=0.7, knn=20){
  # If we have fullsample_idents provided (so coming from scClustEval)
  if(is.null(fullsample_idents)==FALSE){
    ident <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb, resolution == res, k_param == knn)  %>%
    pull(original_ident_full) %>%
    `[[`(1)
    objectoplot <- AddMetaData(objectoplot, metadata = ident, col.name = "set1")
    return(DimPlot(objectoplot, reduction = "umap", label = TRUE, group.by = "set1", repel = TRUE) + ggtitle("scClustEval (Elzar)"))
  }
  # If we have the object with clusters but with no UMAP
  else if(is.null(fullsample_idents)==TRUE & is.null(objectwithclusters)==FALSE){
    
  }
}

#___________________________
# FUNCTION TO LOOK AT UMAP WITH 3 DIFFERENT PARAMETER SETS
paramUMAP3 <- function(seuratobject, fullsample_idents, pc_nb=c(15,15,15), res=c(0.7, 0.7, 0.7), knn=c(20,20,20)){
  # Plot1
  cat("Set1:", pc_nb[1], res[1], knn[1], "\n")
  ident1 <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb[1], resolution == res[1], k_param == knn[1])  %>%
    pull(original_ident_full) %>%
    `[[`(1)
  seuratobject <- AddMetaData(seuratobject, metadata = ident1, col.name = "set1")

  # Plot2
  cat("Set2:", pc_nb[2], res[2], knn[2], "\n")
  ident2 <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb[2], resolution == res[2], k_param == knn[2])  %>%
    pull(original_ident_full) %>%
    `[[`(1)
  seuratobject <- AddMetaData(seuratobject, metadata = ident2, col.name = "set2")
  
  # Plot3
  cat("Set3:", pc_nb[3], res[3], knn[3], "\n")
  ident3 <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb[3], resolution == res[3], k_param == knn[3])  %>%
    pull(original_ident_full) %>%
    `[[`(1)
  seuratobject <- AddMetaData(seuratobject, metadata = ident3, col.name = "set3")
  
  # Combine plots
  p1 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by = "set1", repel = TRUE) +
    ggtitle(paste0("PC=", pc_nb[1], ", resolution=", res[1], ", k.param=", knn[1])) #+ NoLegend()
  p2 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by = "set2", repel = TRUE) +
    ggtitle(paste0("PC=", pc_nb[2], ", resolution=", res[2], ", k.param=", knn[2])) #+ NoLegend()
  p3 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by = "set3", repel = TRUE) +
    ggtitle(paste0("PC=", pc_nb[3], ", resolution=", res[3], ", k.param=", knn[3])) #+ NoLegend()
  print(p1 | p2 | p3)
}
#___________________________
```

```{r function-subsampled-data}
#___________________________
# FUNCTION TO LOOK AT JACCARD INDEX SUBSAMPLED DATA FOR ONE PARAMETER SET
subsampling <- function(seuratobject, fullsample_idents, subsample_idents_list, pc_nb=15, res=0.7, knn=20){
  plotitle <- paste0("cluster id w/ pc=",pc_nb,", res=", res, ", k=", knn)
  
  # Jaccard rainbow cloud
  row_nb <- subsample_idents_list %>%
    ungroup() %>%
    mutate(id = row_number()) %>%
    filter(pc == pc_nb, resolution == res, k_param == knn) %>%
    pull(id)
  p1 <- JaccardRainCloudPlot(subsample_idents_list$data[[row_nb]]$original_ident,
                             subsample_idents_list$data[[row_nb]]$recluster_ident) + 
    geom_hline(yintercept = c(0.6, 0.85), linetype = 2) +
    ylim(c(0,1))+
          #xlab(plotitle)
    ggtitle("Jaccard index btw subsample and total data")
  
  # UMAP
  ident <- fullsample_idents %>%
    dplyr::filter(pc == pc_nb, resolution == res, k_param == knn)  %>%
    pull(original_ident_full) %>%
    `[[`(1)
  seuratobject <- AddMetaData(seuratobject, metadata = ident, col.name = "set1")
  p2 <- DimPlot(seuratobject, reduction = "umap", label = TRUE, group.by = "set1", repel = TRUE) +
      ggtitle(plotitle)
  
  # Silhouette score
  Idents(seuratobject) <- seuratobject@meta.data$set1
  p3 <- SilhouetteRainCloudPlot(CalculateSilhouette(seuratobject, dims=1:pc_nb)) +
    geom_hline(yintercept = 0, linetype = 2) +
    ylim(c(-0.6,0.6))+
    ggtitle("Cluster separation (silhouette width)")
  
  print(p2 | p1 | p3)
}
```




## 4. scClustEval on thymic iNKT cells

```{r import-data-inkt}
# Import data
nkt.seur <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/input/nkt_object_22-11-10.rds")
nkt.subsampling  <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/gather_subsample.rds")
nkt.fullsample <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/gather_full_sample.rds")

# Take quick peek at the data
# nkt.seur
# nkt.seur@meta.data
DimPlot(nkt.seur)

# Preprocess data to have UMAP computed on nkt cells
# nkt.sct_split <- Preprocess(seuratobject = nkt.seur, SCTsplit = T, harmony_vars = c("Sex", "Method"), printComputation = F)
# DimPlot(nkt.sct_split, group.by="Donor") # sanity check
nkt.sct_merged <- Preprocess(seuratobject = nkt.seur, SCTsplit = F, harmony_vars = c("Sex", "Method"), printComputation = F)
DimPlot(nkt.sct_merged, group.by="Donor") # sanity check
ElbowPlot(nkt.sct_merged, ndims=50)
```


### 4.1. Explore different parameters on NKT dataset

```{r nkt-full-data, width=15}
## nb of clusters for each different combination of parameter set
nkt.fullsample %>%
  mutate(cluster_num = purrr::map_dbl(original_ident_full, ~length(unique(.x)))) %>% # 4 to 15 clusters
  arrange(cluster_num)


# Look at some different parameter sets
paramUMAP(nkt.sct_merged, fullsample_idents=nkt.fullsample, pc_nb=c(20,15,10), res=c(0.4, 0.7,1), knn=c(50, 20,10)) # lowest & highest number of clusters
paramUMAP(nkt.seur,       fullsample_idents=nkt.fullsample, pc_nb=c(20,15,10), res=c(0.4, 0.7,1), knn=c(50, 20,10)) # lowest & highest number of clusters
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/lowest_highest_cluster_nb2.jpeg", width=15, height=5)
```


### 4.2. Look at stable clusters with the subsampling

```{r nkt-subsampling}
# Remodel subsample_idents
nkt.subsampling
nkt.subsampling_list <- nkt.subsampling %>% 
  group_by(pc, resolution, k_param) %>% 
  nest()

# Identify NKT stable clusters
nkt.stableClust <- nkt.subsampling_list %>%
    mutate(stable_cluster = map(data, ~ AssignStableCluster(.x$original_ident,
                                                          .x$recluster_ident,
                                                          jaccard_cutoff = 0.6,
                                                          method = "jaccard_median"))) 
                                                          #percent_cutoff = 0.70)))

# Plot total number of clusters and number of stable clusters
ParameterSetScatterPlot(stable_clusters = nkt.stableClust,
                        fullsample_idents = nkt.fullsample,
                        x_var = "k_param",
                        y_var = "number",
                        facet_rows = "resolution",
                        facet_cols = "pc") + ggtitle("stable cluster if jaccard_median > 0.6")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/nkt_subsampling_stableclusters1.jpeg", width=7, height=5)

# Plot percentage of cells in stable clusters
ParameterSetScatterPlot(stable_clusters = nkt.stableClust,
                        fullsample_idents = nkt.fullsample,
                        x_var = "k_param",
                        y_var = "percentage",
                        facet_rows = "resolution",
                        facet_cols = "pc") +
  ggtitle("% of cells in stable clusters")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/nkt_subsampling_stableclusters2.jpeg", width=6, height=5)


# Look at parameter set of interest

# Call function
subsampling(nkt.sct_merged, fullsample_idents=nkt.fullsample, subsample_idents_list=nkt.subsampling_list, pc_nb=15, res=0.7, knn=50)
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/nkt_subsampling_pc10-res07-k20.jpeg", width=15, height=5)


# Verification scClustEval vs local
# test <- nkt.sct_merged
# ident <- nkt.fullsample %>%
#     dplyr::filter(pc == 15, resolution == 0.7, k_param == 50)  %>%
#     pull(original_ident_full) %>%
#     `[[`(1)
# test <- AddMetaData(test, metadata = ident, col.name = "scclusteval")
# b <- DimPlot(test, reduction = "umap", label = TRUE, group.by = "scclusteval", repel = TRUE) +
#       ggtitle("pc=15, res=0.7, k=50 | scClustEval")
# 
# test <- FindNeighbors(object = test, dims = 1:15, assay = "SCT", reduction = "harmony", verbose=F, k.param = 50) %>%
#     FindClusters(resolution = 0.7, verbose=F, random.seed=42)
# c <- DimPlot(test) + ggtitle ("pc=15, res=0.7, k=50 | local computer")
# 
# ggpubr::ggarrange(a,b,c, ncol=3, widths=c(1,0.8,0.8))
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_nkt/nkt_pc15-res07-k50.jpeg", width=17, height=5)

```




## 5. scClustEval on thymic MAIT cells

```{r import-data-mait}
# Import data
mait.seur <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/input/mait_object_22-11-10.rds")
mait.subsampling  <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_mait/gather_subsample.rds")
mait.fullsample <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_mait/gather_full_sample.rds")

# Take quick peek at the data
# mait.seur
# mait.seur@meta.data
DimPlot(mait.seur)

# Preprocess data to have UMAP computed on nkt cells
# mait.sct_split <- Preprocess(seuratobject = mait.seur, SCTsplit = T, harmony_vars = c("Sex", "Method"), printComputation = F)
# DimPlot(mait.sct_split, group.by="Donor") # sanity check
mait.sct_merged <- Preprocess(seuratobject = mait.seur, SCTsplit = F, harmony_vars = c("Sex", "Method"), printComputation = F)
DimPlot(mait.sct_merged, group.by="Donor") # sanity check
ElbowPlot(mait.sct_merged, ndims=50)
```

## 5.2. Look at stable clusters with the subsampling

```{r mait-subsampling}
# Remodel subsample_idents
mait.subsampling
mait.subsampling_list <- mait.subsampling %>% 
  group_by(pc, resolution, k_param) %>% 
  nest()

# Identify MAIT stable clusters
mait.stableClust <- mait.subsampling_list %>%
    mutate(stable_cluster = map(data, ~ AssignStableCluster(.x$original_ident,
                                                          .x$recluster_ident,
                                                          jaccard_cutoff = 0.85,
                                                          method = "jaccard_median"))) 
                                                          #percent_cutoff = 0.70)))

# Plot total number of clusters and number of stable clusters
ParameterSetScatterPlot(stable_clusters = mait.stableClust,
                        fullsample_idents = mait.fullsample,
                        x_var = "k_param",
                        y_var = "number",
                        facet_rows = "resolution",
                        facet_cols = "pc") + ggtitle("stable cluster if jaccard_median > 0.6")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_mait/mait_subsampling_stableclusters1_2.jpeg", width=7, height=7)

# Plot percentage of cells in stable clusters
ParameterSetScatterPlot(stable_clusters = mait.stableClust,
                        fullsample_idents = mait.fullsample,
                        x_var = "k_param",
                        y_var = "percentage",
                        facet_rows = "resolution",
                        facet_cols = "pc") +
  ggtitle("% of cells in stable clusters")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_mait/mait_subsampling_stableclusters2_2.jpeg", width=6, height=7)


# Look at parameter set of interest

# Call function
subsampling(mait.sct_merged, fullsample_idents=mait.fullsample, subsample_idents_list=mait.subsampling_list, pc_nb=10, res=0.2, knn=40)
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_mait/mait_subsampling_pc10-res02-k30.jpeg", width=15, height=5)

# ElbowPlot(test)
# test <- mait.sct_merged
# test <- FindNeighbors(object = test, dims = 1:10, k.param = 20, assay = "SCT", reduction = "harmony", verbose=F) %>%
#   FindClusters(resolution = 0.2, verbose=F, random.seed=42)
# DimPlot(test)
```




## 6. scClustEval on thymic gdT cells

```{r import-data-gdt}
# Import data
gdt.seur <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/input/gdt_object_22-11-10.rds")
gdt.subsampling  <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gather_subsample.rds")
gdt.fullsample <- readRDS("../../data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gather_full_sample.rds")

# Take quick peek at the data
# gdt.seur
# gdt.seur@meta.data
DimPlot(gdt.seur)

# Preprocess data to have UMAP computed on nkt cells
# gdt.sct_split <- Preprocess(seuratobject = gdt.seur, SCTsplit = T, harmony_vars = c("Sex", "Method"), printComputation = F)
# DimPlot(gdt.sct_split, group.by="Donor") # sanity check
gdt.sct_merged <- Preprocess(seuratobject = gdt.seur, SCTsplit = F, harmony_vars = c("Sex", "Method"), printComputation = F)
DimPlot(gdt.sct_merged, group.by="Donor") # sanity check
ElbowPlot(gdt.sct_merged, ndims=50)
```

## 5.2. Look at stable clusters with the subsampling

```{r gdt-subsampling}
# Remodel subsample_idents
gdt.subsampling
gdt.subsampling_list <- gdt.subsampling %>% 
  group_by(pc, resolution, k_param) %>% 
  nest()

# Identify gdT stable clusters
gdt.stableClust <- gdt.subsampling_list %>%
    mutate(stable_cluster = map(data, ~ AssignStableCluster(.x$original_ident,
                                                          .x$recluster_ident,
                                                          jaccard_cutoff = 0.6,
                                                          method = "jaccard_median")))
                                                          #percent_cutoff = 0.70)))

# Plot total number of clusters and number of stable clusters
ParameterSetScatterPlot(stable_clusters = gdt.stableClust,
                        fullsample_idents = gdt.fullsample,
                        x_var = "k_param",
                        y_var = "number",
                        facet_rows = "resolution",
                        facet_cols = "pc") + ggtitle("stable cluster if jaccard_median > 0.6")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gdt_subsampling_stableclusters1.jpeg", width=7, height=5)

# Plot percentage of cells in stable clusters
ParameterSetScatterPlot(stable_clusters = gdt.stableClust,
                        fullsample_idents = gdt.fullsample,
                        x_var = "k_param",
                        y_var = "percentage",
                        facet_rows = "resolution",
                        facet_cols = "pc") +
  ggtitle("% of cells in stable clusters")
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gdt_subsampling_stableclusters2.jpeg", width=6, height=5)


# Look at parameter set of interest

# Call function
subsampling(gdt.sct_merged, fullsample_idents=gdt.fullsample, subsample_idents_list=gdt.subsampling_list, pc_nb=15, res=0.4, knn=20)
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gdt_subsampling_pc15-res04-k20.jpeg", width=15, height=5)
subsampling(gdt.sct_merged, fullsample_idents=gdt.fullsample, subsample_idents_list=gdt.subsampling_list, pc_nb=20, res=0.4, knn=10)
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gdt_subsampling_pc20-res04-k10.jpeg", width=15, height=5)
subsampling(gdt.sct_merged, fullsample_idents=gdt.fullsample, subsample_idents_list=gdt.subsampling_list, pc_nb=20, res=1, knn=20)
# ggsave("~/Projects/HumanThymusProject/data/human-thymus/HumanData_05_scClustEval/output/thymic_gdt/gdt_subsampling_pc20-res1-k20.jpeg", width=15, height=5)

subsampling(gdt.sct_merged, fullsample_idents=gdt.fullsample, subsample_idents_list=gdt.subsampling_list, pc_nb=20, res=0.7, knn=20)

# ElbowPlot(test)
# test <- gdt.sct_merged
# test <- FindNeighbors(object = test, dims = 1:10, k.param = 20, assay = "SCT", reduction = "harmony", verbose=F) %>%
#   FindClusters(resolution = 0.2, verbose=F, random.seed=42)
# DimPlot(test)
```


