---
title: "Human unconventional T cell data - Compare SCTransform on split or merged data"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


# IMPORT

```{r import-librairies, message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(harmony)
library(Seurat)
library(RColorBrewer)
```

```{r import-data}
# Import seurat object
seur.human <- readRDS("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/raw_data/human_data/filtered_seurat_Harmony_07-22-22.RDS")

# Sanity check
print(seur.human) # 79,801 cells (it's the whole seurat object)

# Quick visualization
DimPlot(seur.human)
```

```{r export-seurat-human-scclusteval, include=FALSE}
# Modify seurat object to be fed to scClustEval
seur.human@meta.data[, c("nCount_SCT", "nFeature_SCT", "SCT_snn_res.0.7", "seurat_clusters", "SCT_snn_res.0.9", "SCT_snn_res.1")] <- NULL # remove some cluster columns to avoid any confusion
colnames(seur.human@meta.data)[14] <- "clusters_bigUMAP" # rename "new_clusters" column
# saveRDS(seur.human, "~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/07_scClustEval/input/tcells_full_object_22-11-10.rds")
```




# DEFINE FUNCTIONS

```{r functions}
#_______________
# PREPROCESSING #
#________________
Preprocess <- function(seuratobject, SCTsplit=T, res=0.5, harmony_vars= c("Method", "Batch"), printComputation=T){
  
  # Extract the counts data (from RNA assay) and metadata
  counts <- GetAssayData(object = seuratobject, slot = "counts", assay="RNA")
  metadata <- seuratobject@meta.data
  metadata[, c("nCount_SCT", "nFeature_SCT", "SCT_snn_res.0.7", "seurat_clusters", "SCT_snn_res.0.9", "SCT_snn_res.1")] <- NULL # remove some cluster columns to avoid any confusion
  colnames(metadata)[14] <- "clusters_bigUMAP" # rename "new_clusters" column
  # Create new seurat object with only the counts and the "cleaned" metadata
  initseurat <- CreateSeuratObject(counts = counts, meta.data = metadata, min.cells=20) # keep only genes expressed in at least 20 cells
  
  cat("Initial seurat object:\n")
  print(initseurat) # this is for sanity check
  
  if(SCTsplit==T){
    cat("\n1-Perform SCTransform on split data\n")
    # Normalize with SCTransform by splitting object (per batch)
    seur.list <- SplitObject(initseurat, split.by = "orig.ident")
    for(i in 1:length(seur.list)){
      cat("...SCTransform on", names(seur.list)[i], "\n")
      seur.list[[i]] <-  SCTransform(seur.list[[i]], vars.to.regress = 'percent.mt', verbose=printComputation, seed.use = 1448145)
    }
    # seur.list <- lapply(X = seur.list, FUN = SCTransform, vars.to.regress="percent.mt", method="glmGamPois") # not working
    seur.SCT <- merge(seur.list[[1]], y = unlist(seur.list[-1], use.names=FALSE), merge.data = TRUE)
    sct.hvg  <- SelectIntegrationFeatures(object.list = seur.list, nfeatures = 3000)
    VariableFeatures(seur.SCT) <- sct.hvg
  }
  
  else if(SCTsplit==F){
    cat("\n1-Perform SCTransform on merged data")
    # Normalize with SCTransform (without splitting object)
    seur.SCT <- SCTransform(initseurat,
                            assay="RNA",
                            new.assay.name="SCT",
                            vars.to.regress = 'percent.mt',
                            seed.use = 1448145,
                            verbose=printComputation)
  }
  
  # Run PCA (don't scale data when using SCTransform)
  cat("\n2-Run PCA pre-integration")
  seur.SCT <- RunPCA(object = seur.SCT, assay = "SCT", npcs = 50, seed.use=42, verbose=printComputation)
  p1 <- DimPlot(seur.SCT, reduction = "pca", group.by = "Batch") + ggtitle("PCA pre-integration")
  ElbowPlot(seur.SCT)
  cat("\nSeurat object pre-integration:\n")
  print(seur.SCT)
  
  # Run Harmony for integration
  cat("\n3-Run Harmony\n")
  seur.harmony <- RunHarmony(object = seur.SCT,
                             assay.use = "SCT",
                             reduction = "pca",
                             dims.use = 1:50,
                             max.iter.harmony=30,
                             group.by.vars = harmony_vars,
                             plot_convergence = printComputation)
  cat("\nSeurat object post-integration:\n")
  print(seur.harmony)
  cat("\n4-Run UMAP & clustering\n")
  seur.harmony <- RunUMAP(object = seur.harmony, dims = 1:15, assay = "SCT", reduction = "harmony", verbose=printComputation, seed.use=42)
  seur.harmony <- FindNeighbors(object = seur.harmony, dims = 1:15, assay = "SCT", reduction = "harmony", verbose=printComputation) %>%
    FindClusters(resolution = res, verbose=printComputation, random.seed=42)
  
  # Show PCA after integration
  p2 <- DimPlot(seur.harmony, reduction = "harmony", group.by = "Batch") + ggtitle("PCA post-integration")
  
  print(p1 | p2)
  
  return(seur.harmony)
}




#__________________
# GENE SIGNATURES #
#__________________

# Gene signatures from Park paper
parksignatures <- read.csv("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/park_genesignatures.csv", row.names=1)
Egress <- list(c("KLF2", "CORO1A", "CCR7", "CXCR4", "CXCR6", "FOXO1", "CXCR3", "S1PR1", "S1PR4",
                 "S100A4", "S100A6", "EMP3"))


# Function to look at these gene signatures
signatures <- function(seuratobject=NKT_SCTmerged, scorelist=DN_score, scorename="DN_score", celltype="NKT"){
  # Add signatures to new seurat object
  # NKT_SCTsplit_score <- AddModuleScore(object = NKT_SCTsplit, features = scorelist, assay = "SCT", name = scorename)
  seurat_scored <- AddModuleScore(object = seuratobject, features = scorelist, assay = "SCT", name = scorename)

  # SCpubr::do_FeaturePlot(sample = NKT_SCTsplit_score,
  #                        features = paste0(scorename, "1"),
  #                        plot.title = "NKT cells | SCTransform on split data",
  #                        reduction = "umap",
  #                        viridis_color_map = "inferno") |
  SCpubr::do_FeaturePlot(sample = seurat_scored, 
                         features = paste0(scorename, "1"),
                         plot.title = paste(celltype, "cells"),
                         reduction = "umap",
                         viridis_color_map = "inferno")
}
```





# THYMIC CONVENTIONAL T CELLS

## 1. Preprocessing

```{r cd4cd8-thym-preprocessing, fig.width=10}
# Isolate CD4 and CD8 T cells
CD4CD8_Thymus <- subset(seur.human, ident = c("CD4_Thymus", "CD8_Thymus")) # 27351 cells

# Quick QC
FeatureScatter(CD4CD8_Thymus, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by="Batch")
FeatureScatter(CD4CD8_Thymus, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by="Batch")

# PREPROCESS
CD4CD8_SCTsplit <- Preprocess(seuratobject = CD4CD8_Thymus, SCTsplit = T, res = 0.3, harmony_vars = c("Sex", "Method"), printComputation = F)
CD4CD8_SCTmerged <- Preprocess(seuratobject = CD4CD8_Thymus, SCTsplit = F, res = 0.3, harmony_vars = c("Sex", "Method"), printComputation = F)
```


## 2. Visualize

### 2.1. CD4/CD8 UMAP with SCTransform on split vs merged data

```{r cd4cd8-thym-umap, fig.width=10}
# Sanity checks
# DimPlot(CD4CD8_SCTsplit, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on split data") |
#   DimPlot(CD4CD8_SCTmerged, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on merged data")

# Look at clusters
DimPlot(CD4CD8_SCTsplit, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("SCT on split data") |
  DimPlot(CD4CD8_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("SCT on merged data")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_SCTransform_split-vs-merged.jpeg", width = 12, height=6)
# Look at CD4/CD8 separation
DimPlot(CD4CD8_SCTsplit, dims = 1:2, reduction = "umap", group.by = "cell.ident", label = T) + ggtitle("SCT on split data") |
  DimPlot(CD4CD8_SCTmerged, dims = 1:2, reduction = "umap", group.by = "cell.ident", label = T) + ggtitle("SCT on merged data")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_SCTransform_split-vs-merged2.jpeg", width = 12, height=6)

# Look at some features
SCpubr::do_FeaturePlot(sample = CD4CD8_SCTsplit, 
                       features = c("CD4", "CD8A", "CD8B", "FOXP3"),
                       # plot.title = paste(celltype, "cells"),
                       order=T,
                       ncol=4,
                       reduction = "umap",
                       viridis_color_map = "inferno")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_featureexpression.jpeg", width = 18, height=6)
```

### 2.2. CD4/CD8 T gene signatures

```{r mait-signatures}
# Look at gene signatures
Cd8aa1_score <- list(c("NUCB2", "CD27", "LEF1", "PTPN7", "TRAC", "PRKCH", "CD28", "LCP2", "CD2", "IKZF1", "AQP3",
                      "GNG4", "SH3BGRL3", "LIMD2", "ELOVL5", "TOX", "SH2D1A", "CAPZA1", "XCL1"))
Cd8aa2_score <- list(c("CTSW", "TRGC2", "SMC4", "CD7", "IKZF2", "ZNF683", "ACTN1", "SMIM24", "HCST", "MME",
                       "TRG-AS1", "RGS2", "RTKN2", "PPP2R5C", "ID3", "TRGC1", "CLEC2D", "DCXR", "ELF1", "CPNE7"))

signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$gdT), scorename="gdT_score") # negative control
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_1.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$DPq),  scorename="DPq_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_2.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$DPp),  scorename="DPp_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_3.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8aa1),  scorename="CD8aa1_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_4.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8aa2),  scorename="CD8aa2_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_5.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$NKT),     scorename="NKT_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_6.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$abT_entry), scorename="SPentry_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_7.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$T_agonist),    scorename="Tagonist_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_8.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=Egress,    scorename="Egress_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_9.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD4T),    scorename="CD4T_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_10.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$Th17),    scorename="Th17_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_11.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$Treg),    scorename="Treg_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_12.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$Treg_diff),  scorename="Treg_diff_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_13.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8T),  scorename="CD8T_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_14.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8T_mem),  scorename="CD8Tmem_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_15.jpeg", width=9, height=7)
signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$DN_early),    scorename="DNearly_score") |
  signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$DNp),    scorename="DNp_score") |
  signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$DNq),    scorename="DNq_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_16.jpeg", width=18, height=7)

signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8aa1[-3]),  scorename="CD8aa1_score") |
  signatures(seuratobject=CD4CD8_SCTsplit, celltype = "CD4 & CD8", scorelist=list(parksignatures$CD8aa2),  scorename="CD8aa2_score") |
  SCpubr::do_DimPlot(sample = CD4CD8_SCTsplit, cells.highlight = WhichCells(seur.human, idents="CD4_Thymus"), reduction = "umap", plot.title="CD4 T cells", legend.position="none")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4cd8_parksignatures_17.jpeg", width=18, height=7)
```

```{r}
# Isolate CD4 T cells
CD4_Thymus <- subset(seur.human, ident = "CD4_Thymus") # 16590 cells

# PREPROCESS
CD4_SCTmerged <- Preprocess(seuratobject = CD4_Thymus, SCTsplit = F, res = 0.3, harmony_vars = c("Sex", "Method"), printComputation = F)

# sanity check
# DimPlot(CD4_SCTmerged, dims = 1:2, reduction = "umap", split.by = "Donor", label = T) + ggtitle("SCT on merged data")

# Clusters
DimPlot(CD4_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("SCT on merged data")

# Signature genes
signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$DN_early),    scorename="DNearly_score") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$DNp),    scorename="DNp_score") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$DNq),    scorename="DNq_score")

signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$DPq),  scorename="DPq_score") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$DPp),  scorename="DPp_score")

DimPlot(CD4_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("Thymic CD4 T cells") |
 SCpubr::do_FeaturePlot(sample = CD4_SCTmerged, features = "CD4", plot.title="CD4 cells", order=T, reduction = "umap",viridis_color_map = "inferno") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$CD8aa1[-3]),  scorename="CD8aa1_score") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$CD8aa2),  scorename="CD8aa2_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4_parksignatures_1.jpeg", width=23, height=7)

signatures(seuratobject=CD4_SCTmerged, celltype = "With CD8A gene | CD4", scorelist=list(parksignatures$CD8aa1),  scorename="CD8aa1_score") |
  signatures(seuratobject=CD4_SCTmerged, celltype = "Without CD8A gene | CD4", scorelist=list(parksignatures$CD8aa1[-3]),  scorename="CD8aa1_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/10_HumanCD4CD8data/cd4_parksignatures_2.jpeg", width=14, height=7)

signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$abT_entry), scorename="SPentry_score")

signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$T_agonist),    scorename="Tagonist_score")
signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$Treg_diff),  scorename="Treg_diff_score")
signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$Treg),    scorename="Treg_score")

signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$CD4T),    scorename="CD4T_score")
signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=list(parksignatures$Th17),    scorename="Th17_score")

signatures(seuratobject=CD4_SCTmerged, celltype = "CD4", scorelist=Egress,    scorename="Egress_score")
```





# THYMIC MAIT CELLS

## 1. Preprocessing

```{r mait-thym-preprocessing, fig.width=10}
# Isolate MAIT cells
MAIT_Thymus <- subset(seur.human, ident = "MAIT_Thymus") # 4689 cells

# Quick QC
FeatureScatter(MAIT_Thymus, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by="Batch")
FeatureScatter(MAIT_Thymus, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by="Batch")

# PREPROCESS
MAIT_SCTsplit <- Preprocess(seuratobject = MAIT_Thymus, SCTsplit = T, res = 0.3, harmony_vars = c("Sex", "Method"), printComputation = F)
MAIT_SCTmerged <- Preprocess(seuratobject = MAIT_Thymus, SCTsplit = F, res = 0.3, harmony_vars = c("Sex", "Method"), printComputation = F)

# SAVE OBJECT (for scClustEval)
# MAIT_Thymus@meta.data[, c("nCount_SCT", "nFeature_SCT", "SCT_snn_res.0.7", "seurat_clusters", "SCT_snn_res.0.9", "SCT_snn_res.1")] <- NULL # remove some cluster columns to avoid any confusion
# colnames(MAIT_Thymus@meta.data)[14] <- "clusters_bigUMAP" # rename "new_clusters" column
# saveRDS(MAIT_Thymus, "~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/07_scClustEval/input/mait_object_22-11-10.rds")
```


## 2. Visualize

### 2.1. MAIT UMAP with SCTransform on split vs merged data

```{r mait-thym-umap, fig.width=10}
# Sanity checks
# DimPlot(MAIT_SCTsplit, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on split data") |
#   DimPlot(MAIT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on merged data")

# Look at clusters
DimPlot(MAIT_SCTsplit, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("SCT on split data") |
  DimPlot(MAIT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("SCT on merged data")
```

### 2.2. MAIT gene signatures

```{r mait-signatures}
# Look at gene signatures
# signatures(scorelist=DN_score, scorename="DN_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=gdT_score, scorename="gdT_score") # negative control
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=DPq_score,  scorename="DPq_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=DPp_score,  scorename="DPp_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=Cd8aa1_score,  scorename="CD8aa1_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=Cd8aa2_score,  scorename="CD8aa2_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=NKT_score,     scorename="NKT_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=SPentry_score, scorename="SPentry_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=Tago_score,    scorename="Tagonist_score")
signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=Egress,    scorename="Egress_score")
# signatures(seuratobject=MAIT_SCTmerged, celltype = "MAIT", scorelist=NK_score,      scorename="NK_score") # several genes are not found

# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/08_HumanMAITdata/mait_parksignatures_9.jpeg", width=9, height=7)
```


### 2.3. MAIT clustering resolution

```{r mait-cluster-resolutions}
# Try different resolutions
MAIT_SCT_06 <- Preprocess(seuratobject = MAIT_Thymus, SCTsplit = F, res = 0.6, harmony_vars = c("Sex", "Method"), printComputation = F)
MAIT_SCT_08 <- Preprocess(seuratobject = MAIT_Thymus, SCTsplit = F, res = 0.8, harmony_vars = c("Sex", "Method"), printComputation = F)
MAIT_SCT_1 <- Preprocess(seuratobject = MAIT_Thymus, SCTsplit = F, res = 1, harmony_vars = c("Sex", "Method"), printComputation = F)

DimPlot(MAIT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.3", label = T) + ggtitle("res=0.3") |
  DimPlot(MAIT_SCT_06, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.6", label = T) + ggtitle("res=0.6") |
  DimPlot(MAIT_SCT_08, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.8", label = T) + ggtitle("res=0.8") |
  DimPlot(MAIT_SCT_1, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.1", label = T) + ggtitle("res=1")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/mait_sctransform-merged_resolutions.jpeg", width=20, height=5)
```





# THYMIC NKT CELLS

## 1. Preprocessing

```{r nkt-thym-preprocessing, fig.width=10}
# Isolate NKT cells
NKT_Thymus <- subset(seur.human, ident = "NKT_Thymus") # 2551 cells

# Quick QC
FeatureScatter(NKT_Thymus, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by="Batch")
FeatureScatter(NKT_Thymus, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by="Batch")

# PREPROCESS
NKT_SCTsplit <- Preprocess(seuratobject = NKT_Thymus, SCTsplit = T, res = 0.4, harmony_vars = c("Sex", "Method"), printComputation = F)
NKT_SCTmerged <- Preprocess(seuratobject = NKT_Thymus, SCTsplit = F, res = 0.4, harmony_vars = c("Sex", "Method"), printComputation = F)

# SAVE OBJECT (for scClustEval)
# NKT_Thymus@meta.data[, c("nCount_SCT", "nFeature_SCT", "SCT_snn_res.0.7", "seurat_clusters", "SCT_snn_res.0.9", "SCT_snn_res.1")] <- NULL # remove some cluster columns to avoid any confusion
# colnames(NKT_Thymus@meta.data)[14] <- "clusters_bigUMAP" # rename "new_clusters" column
# saveRDS(NKT_Thymus, "~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/07_scClustEval/input/nkt_object_22-11-10.rds")
```


## 2. Visualize

### 2.1. NKT UMAP with SCTransform on split vs merged data

```{r nkt-thym-umap, fig.width=10}
# Sanity checks
# DimPlot(NKT_SCTsplit, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on split data") |
#   DimPlot(NKT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "Donor", label = T) + ggtitle("SCT on merged data")

# Look at clusters
DimPlot(NKT_SCTsplit, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("SCT on split data") |
  DimPlot(NKT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("SCT on merged data")
```

### 2.2. NKT gene signatures

```{r nkt-signatures}
# Look at gene signatures
# signatures(scorelist=DN_score, scorename="DN_score")  # several genes are not found
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=gdT_score, scorename="gdT_score") # negative control
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_1.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=DPq_score,  scorename="DPq_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_2.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=DPp_score,  scorename="DPp_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_3.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=Cd8aa1_score,  scorename="CD8aa1_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_4.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=Cd8aa2_score,  scorename="CD8aa2_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_5.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=NKT_score,     scorename="NKT_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_6.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=SPentry_score, scorename="SPentry_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_7.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=Tago_score,    scorename="Tagonist_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_8.jpeg", width=9, height=7)
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=Egress,    scorename="Egress_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_parksignatures_9.jpeg", width=9, height=7)
# signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=NK_score,      scorename="NK_score") # several genes are not found

# Signatures of c1 vs c2 cells
c1_score <- list(c("CD4", "COTL1", "LEF1", "MAP3K1", "PLP2", "XCL1", "ICOS"))
c2_score <- list(c("GNLY", "GZMB", "GZMH", "KLRD1", "KLRF1", "KLRK1", "NKG7", "PRF1", "TYROBP", "DBN1", "CD161", "CD94", "CCR5"))
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=c1_score,    scorename="c1_score")
signatures(seuratobject=NKT_SCTmerged, celltype = "NKT", scorelist=c2_score,    scorename="c2_score")
```


### 2.3. NKT clustering resolution

```{r nkt-cluster-resolutions}
# Try different resolutions
NKT_SCT_06 <- Preprocess(seuratobject = NKT_Thymus, SCTsplit = F, res = 0.6, harmony_vars = c("Sex", "Method"), printComputation = F)
NKT_SCT_08 <- Preprocess(seuratobject = NKT_Thymus, SCTsplit = F, res = 0.8, harmony_vars = c("Sex", "Method"), printComputation = F)
NKT_SCT_1 <- Preprocess(seuratobject = NKT_Thymus, SCTsplit = F, res = 1, harmony_vars = c("Sex", "Method"), printComputation = F)

DimPlot(NKT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("res=0.4") |
  DimPlot(NKT_SCT_06, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.6", label = T) + ggtitle("res=0.6") |
  DimPlot(NKT_SCT_08, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.8", label = T) + ggtitle("res=0.8") |
  DimPlot(NKT_SCT_1, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.1", label = T) + ggtitle("res=1")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/nkt_sctransform-merged_resolutions.jpeg", width=20, height=5)
```





# THYMIC GD T CELLS

## 1. Preprocessing

```{r gdt-thym-preprocessing, fig.width=10}
# Isolate gdT cells
gdT_Thymus <- subset(seur.human, ident = "GD_Thymus") # 2981 cells

# Quick QC
FeatureScatter(gdT_Thymus, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by="Batch")
FeatureScatter(gdT_Thymus, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by="Batch")

# PREPROCESS
gdT_SCTsplit <- Preprocess(seuratobject = gdT_Thymus, SCTsplit = T, res = 0.4, harmony_vars = c("Sex", "Method"), printComputation = F)
gdT_SCTmerged <- Preprocess(seuratobject = gdT_Thymus, SCTsplit = F, res = 0.4, harmony_vars = c("Sex", "Method"), printComputation = F)

# SAVE OBJECT (for scClustEval)
# gdT_Thymus@meta.data[, c("nCount_SCT", "nFeature_SCT", "SCT_snn_res.0.7", "seurat_clusters", "SCT_snn_res.0.9", "SCT_snn_res.1")] <- NULL # remove some cluster columns to avoid any confusion
# colnames(gdT_Thymus@meta.data)[14] <- "clusters_bigUMAP" # rename "new_clusters" column
# saveRDS(gdT_Thymus, "~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/07_scClustEval/input/gdt_object_22-11-10.rds")
```


## 2. Visualize


### 2.1. gdT UMAP with SCTransform on split vs merged data

```{r gdt-thym-umap, fig.width=10}
# Sanity checks
# DimPlot(seur.harmony, dims = 1:2, group.by = "orig.ident")
# DimPlot(seur.harmony, dims = 1:2, group.by = "Sex")
# DimPlot(seur.harmony, dims = 1:2, group.by = "Donor")
# DimPlot(seur.harmony, dims = 1:2, group.by = "Batch")
# DimPlot(seur.harmony, dims = 1:2, group.by = "Method")

# Look at clusters
DimPlot(gdT_SCTsplit, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("SCT on split data") |
  DimPlot(gdT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("SCT on merged data")

```

### 2.2. gdT gene signatures

```{r gdT-signatures}
# Look at gene signatures
# signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=DN_score, scorename="DN_score")  # several genes are not found
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=gdT_score, scorename="gdT_score") # negative control
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_1.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=DPq_score,  scorename="DPq_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_2.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=DPp_score,  scorename="DPp_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_3.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=Cd8aa1_score,  scorename="CD8aa1_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_4.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=Cd8aa2_score,  scorename="CD8aa2_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_5.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=NKT_score,     scorename="NKT_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_6.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=SPentry_score, scorename="SPentry_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_7.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=Tago_score,    scorename="Tagonist_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_8.jpeg", width=9, height=7)
signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=Egress,    scorename="Egress_score")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/09_HumanGDTdata/gdt_parksignatures_9.jpeg", width=9, height=7)
# signatures(seuratobject=gdT_SCTmerged, celltype = "gdT", scorelist=NK_score,      scorename="NK_score") # several genes are not found

# rownames(seur.human)[grep("CD1", rownames(seur.human))]
```


### 2.3. gdT clustering resolution

```{r gdt-cluster-resolutions}
# Try different resolutions
GDT_SCT_06 <- Preprocess(seuratobject = gdT_Thymus, SCTsplit = F, res = 0.6, harmony_vars = c("Sex", "Method"), printComputation = F)
GDT_SCT_08 <- Preprocess(seuratobject = gdT_Thymus, SCTsplit = F, res = 0.8, harmony_vars = c("Sex", "Method"), printComputation = F)
GDT_SCT_1 <- Preprocess(seuratobject = gdT_Thymus, SCTsplit = F, res = 1, harmony_vars = c("Sex", "Method"), printComputation = F)

DimPlot(gdT_SCTmerged, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.4", label = T) + ggtitle("res=0.4") |
  DimPlot(GDT_SCT_06, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.6", label = T) + ggtitle("res=0.6") |
  DimPlot(GDT_SCT_08, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.0.8", label = T) + ggtitle("res=0.8") |
  DimPlot(GDT_SCT_1, dims = 1:2, reduction = "umap", group.by = "SCT_snn_res.1", label = T) + ggtitle("res=1")
# ggsave("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/04_HumanNKTdata/gdt_sctransform-merged_resolutions.jpeg", width=20, height=5)
```







