---
title: "Download and prepare Park Thymic Atlas"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```




# IMPORT LIBRAIRIES
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```




# HUMAN THYMUS ATLAS

## Import data
The Human Thymus Atlas was downloaded as a seurat object from [cellxgene](https://cellxgene.cziscience.com/collections/de13e3e2-23b6-40ed-a413-e9e12d7d3910), the one containing 255,901 cells.
```{r import-data-human, fig.height=10, fig.width=15}
# import rds object downloaded
seur_hu <- readRDS("./data_github/park_dataset/c6e08ab6-ab3b-41dc-8058-8e6442e081ec.rds")
seur_hu2 <- readRDS("./data/human-thymus/HumanData_16_ParkIntegration/park_thymus_seurat.rds")
print(seur_hu)

# check gene names
print(rownames(seur_hu)[1:5]) # as they are ENSEMBL IDs, we'll need to translate them to gene symbols

# check metadata columns
SCpubr::do_DimPlot(
  seur_hu, group.by="cell_type", reduction="umap", raster=T, legend.position="right")
```

The cell annotation doesn't correspond to Figure 1 of the [Park et al.](https://www.science.org/doi/10.1126/science.aay3224) paper. We are looking for different annotation levels such as the ones on the [cellatlas.io](https://app.cellatlas.io/thymus-development/dataset/8/scatterplot) website. We downloaded the AnnData object from cellatlas.io, and then exported the metadata as a `.csv` file using the [download_park_thymic_atlas_human_metadata.ipynb](./download_park_thymic_atlas_human_metadata.ipynb) script. Let's import this metadata file.
```{r import-metadata-human}

```








# MURINE THYMUS ATLAS

## Analysis 1
```{r analysis-1}

```


## Analysis 2
```{r analysis-2}

```

## Analysis 1
```{r analysis-1}

```


## Analysis 2
```{r analysis-2}

```

## Analysis 3
```{r analysis-3}

```


## Analysis 4
```{r analysis-4}

```




# SESSION INFO
```{r}
sessionInfo()
```


