---
title: "Generate an orthologs dataframe for cross-species comparisons"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```




# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(tidyverse)
library(biomaRt)
library(Seurat)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```


## Import human & mouse biomarts
```{r import-biomarts} 
# Define marts to map mouse to human and ensembl to entrez IDs
# listDatasets(useMart('ensembl'))
mart.hu <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                   dataset="hsapiens_gene_ensembl")
mart.ms <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                   dataset="mmusculus_gene_ensembl")
```


## Import data
This will be to check that we find a majority of genes of interest (aka genes from our data) in the orthologs table.
```{r import-data}
seur_human <- readRDS("./data_geo/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
seur_mouse <- readRDS("data_geo/seurat_objects/seurat_mouse_integrated_object_23_12_02.rds")

# genes in list
genes_data_list <- list(
  "hu"=unique(rownames(seur_human)),
  "ms"=unique(rownames(seur_mouse))
)

# check nb of genes
lapply(genes_data_list, function(x) length(x))
```




# GET ORTHOLOG TABLES

Try different ways of getting the mouse-human orthologs table (get human orthologs of mouse genes, or vice-versa).

## Human orthologs of mouse genes
```{r biomart-ms-reference}
# Check available attributes in the mouse mart
# listAttributes(mart.ms)[stringr::str_detect(listAttributes(mart.ms)$name, "hsapiens"),]

# Get human orthologs of mouse genes
ms.ortholgs <-
  getBM(
    attributes = c(
      'ensembl_gene_id',
      'external_gene_name',
      'hsapiens_homolog_ensembl_gene',
      'hsapiens_homolog_associated_gene_name',
      'hsapiens_homolog_orthology_type',
      'hsapiens_homolog_orthology_confidence'
    ),
    filters = 'external_gene_name',
    values = genes_data_list$ms,
    mart = mart.ms
  ) #%>%
  # dplyr::rename(
  #   mmusculus_ensembl_gene_id=ensembl_gene_id,
  #   mmusculus_gene_name=external_gene_name
  # ) %>%
  # filter(hsapiens_homolog_associated_gene_name!="")
table(genes_data_list$ms %in% ms.ortholgs$external_gene_name, useNA="ifany") # 16,069 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% ms.ortholgs$hsapiens_homolog_associated_gene_name, useNA="ifany") # 4,993 hu gene symbols in our data not in the biomart
```


## Mouse orthologs of human genes
```{r biomart-hu-reference}
# Get mouse orthologs of human genes
hu.ortholgs <-
  getBM(
    attributes = c(
      'ensembl_gene_id',
      'external_gene_name',
      'mmusculus_homolog_ensembl_gene',
      'mmusculus_homolog_associated_gene_name',
      'mmusculus_homolog_orthology_type',
      'mmusculus_homolog_orthology_confidence'
    ),
    filters = 'external_gene_name',
    values = genes_data_list$hu,
    mart = mart.hu
  )
table(genes_data_list$ms %in% hu.ortholgs$mmusculus_homolog_associated_gene_name, useNA="ifany") # 18,633 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% hu.ortholgs$external_gene_name, useNA="ifany") # 3,217 hu gene symbols in our data not in the biomart
```


## Imported tables from online biomart
I imported these tables from [online biomart](https://useast.ensembl.org/biomart/martview/8ea546479bd07c2257297b0b8e0ddb3a) on Feb 21st 2024. I tried two methods:
- in the Mouse genes dataset (GRCm39), I chose attributes "Gene stable ID", "Gene stable ID version", "Human gene stable ID", "Human gene name", "Human homology type" and "Human orthology confidence";
- in the Human genes dataset (GRCh38.p14), I chose attributes "Gene stable ID", "Gene stable ID version", "Mouse gene stable ID", "Mouse gene name", "Mouse homology type" and "Mouse orthology confidence".

Then I downloaded the results tables in .csv format:
```{r biomart-online}
# Import biomart (mouse as reference)
biomart.ms <- read.csv("./data_github/cross_species/biomart_export_ms.txt", header=T)
colnames(biomart.ms) <-
  c(
    "ms_ensemblID",
    "ms_ensemblID_version",
    "ms_symbol",
    "hu_ensemblID",
    "hu_symbol",
    "hu_homology_type",
    "hu_orthology_confidence"
  )
# remove genes with no human ortholog, keep only 1:1 orthologs
biomart.ms <- biomart.ms %>%
  filter(hu_homology_type=="ortholog_one2one") %>%
  filter(hu_symbol!="")
table(genes_data_list$ms %in% biomart.ms$ms_symbol, useNA="ifany") # 16,069 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% biomart.ms$hu_symbol, useNA="ifany") # 4,954 hu gene symbols in our data not in the biomart


# Import biomart (human as reference)
biomart.hu <- read.csv("./data/cross_species/biomart_export_hu.txt", header=T)
colnames(biomart.hu) <-
  c(
    "hu_ensemblID",
    "hu_ensemblID_version",
    "hu_symbol",
    "ms_ensemblID",
    "ms_symbol",
    "ms_homology_type",
    "ms_orthology_confidence"
  )
biomart.hu <- biomart.hu %>%
  filter(ms_homology_type=="ortholog_one2one") %>%
  filter(ms_symbol!="")
table(genes_data_list$ms %in% biomart.hu$ms_symbol, useNA="ifany") # 16,062 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% biomart.hu$hu_symbol, useNA="ifany") # 4,954 hu gene symbols in our data not in the biomart


# Merge the 2 tables
biomart.df <- biomart.ms %>%
  left_join(biomart.hu, by=  c(
    "hu_ensemblID",
    "hu_symbol",
    "ms_ensemblID",
    "ms_symbol"
  )) %>%
  dplyr::select(-c(hu_homology_type, ms_homology_type)) %>%
  distinct()
table(genes_data_list$ms %in% biomart.df$ms_symbol, useNA="ifany") # 16,069 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% biomart.df$hu_symbol, useNA="ifany") # 4,954 hu gene symbols in our data not in the biomart
```

The best compromise is the biomart table imported from online (compared to using previously biomaRt package).




# SAVE ORTHOLOG TABLE

## Clean-up ortholog table (keep only 1:1 orthologs and orthology confidence of 1)
```{r clean-up-df}
ortholog.df <- biomart.df %>%
  filter(hu_orthology_confidence==1 & ms_orthology_confidence==1) %>%
  dplyr::select(-c(hu_orthology_confidence, ms_orthology_confidence)) %>%
  mutate(date=Sys.Date())

# checkup
table(genes_data_list$ms %in% ortholog.df$ms_symbol) # 16,606 ms gene symbols from data not in the biomart (out of ~31,000)
table(genes_data_list$hu %in% ortholog.df$hu_symbol) # 5,238 hu gene symbols in our data not in the biomart (out of ~17,000)
```


## Bigger ortholog table (with 1:many orthologs and orthology confidence of 0 or 1)
I noticed some important genes were 1:many orthologs or had orthology confidence of 0 (e.g. _Cd8_, ....). So let's create and save a bigger orthologs table that we can use depending on the analysis.
```{r}
# Import biomart (mouse as reference)
biomart.ms <- read.csv("./data_github/cross_species/biomart_export_ms.txt", header=T)
colnames(biomart.ms) <-
  c(
    "ms_ensemblID",
    "ms_ensemblID_version",
    "ms_symbol",
    "hu_ensemblID",
    "hu_symbol",
    "hu_homology_type",
    "hu_orthology_confidence"
  )
# remove genes with no human ortholog
biomart.ms <- biomart.ms %>%
  filter(hu_symbol!="")
# table(genes_data_list$ms %in% biomart.ms$ms_symbol, useNA="ifany") # 14,227 ms gene symbols from data not in the biomart
# table(genes_data_list$hu %in% biomart.ms$hu_symbol, useNA="ifany") # 4,435 hu gene symbols in our data not in the biomart


# Import biomart (human as reference)
biomart.hu <- read.csv("./data_github/cross_species/biomart_export_hu.txt", header=T)
colnames(biomart.hu) <-
  c(
    "hu_ensemblID",
    "hu_ensemblID_version",
    "hu_symbol",
    "ms_ensemblID",
    "ms_symbol",
    "ms_homology_type",
    "ms_orthology_confidence"
  )
biomart.hu <- biomart.hu %>%
  filter(ms_symbol!="")
# table(genes_data_list$ms %in% biomart.hu$ms_symbol, useNA="ifany") # 14,217 ms gene symbols from data not in the biomart
# table(genes_data_list$hu %in% biomart.hu$hu_symbol, useNA="ifany") # 4,437 hu gene symbols in our data not in the biomart


# Merge the 2 tables
ortholog.df.full <- biomart.ms %>%
  left_join(biomart.hu, by=  c(
    "hu_ensemblID",
    "hu_symbol",
    "ms_ensemblID",
    "ms_symbol"
  )) %>%
  # remove rows that don't have ms symbols
  filter(ms_symbol!="") %>%
  dplyr::select(-c(hu_orthology_confidence, hu_homology_type)) %>%
  distinct() %>%
  dplyr::rename(
    orthology_confidence=ms_orthology_confidence,
    homology_type=ms_homology_type
    )
# sanity checks
table(biomart.df$homology_type, useNA="ifany")
table(biomart.df$orthology_confidence, useNA="ifany")

table(genes_data_list$ms %in% biomart.df$ms_symbol, useNA="ifany") # 14,227 ms gene symbols from data not in the biomart
table(genes_data_list$hu %in% biomart.df$hu_symbol, useNA="ifany") # 4,437 hu gene symbols in our data not in the biomart
```




## Save
```{r save-df, echo=T, eval=F}
# Save
write.csv(ortholog.df, "./data_github/cross_species/ortholog_table_ms_hu_one2one.csv", row.names = F)
write.csv(ortholog.df.full, "./data_github/cross_species/ortholog_table_ms_hu_full.csv", row.names = F)
```




# SESSION INFO
```{r}
sessionInfo()
```
