---
title: "Plot pyMetaNeighbor across species (Figure 6B)"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(RColorBrewer)
library(reshape2)
library(ggrepel)
library(patchwork)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```


## Import data
```{r import-data, fig.width=7}
# human
seur.hu <- list(
  "nkt"=readRDS("./data_github/seurat_objects_thymus/seurat_thymus_inkt.rds"),
  "mait"=readRDS("./data_github/seurat_objects_thymus/seurat_thymus_mait.rds"),
  "gdt"=readRDS("./data_github/seurat_objects_thymus/seurat_thymus_gdt.rds")
)
DimPlot(seur.hu$nkt,  group.by="clusters_per_lineage", cols=cols_thym_nkt)
DimPlot(seur.hu$mait, group.by="clusters_per_lineage", cols=cols_thym_mait)
DimPlot(seur.hu$gdt,  group.by="clusters_per_lineage", cols=cols_thym_gdt)

# mouse
seur.ms <- readRDS("./data_geo/seurat_objects/seurat_mouse_integrated_object_23_12_02.rds")
DimPlot(seur.ms,  group.by="clusters_annotation", label=T)
```

Import output from metaneighbor
```{r import-data-pymtn}
mtn <- read.csv("./data_github/cross_species/innateT_ms_hu_pymtn_output_2024_08_27.csv", row.names=1)
mtn <- as.matrix(mtn)
mtn[1:5,1:5]
```




# PLOT METANEIGHBOR

## Quick look at heatmap
```{r mtn-heatmap, fig.height=10, fig.width=10}
gplots::heatmap.2(
  mtn,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human innate T (4284 HVGs)",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(9, 9)
)
```


## Define plotting function
```{r define-function}
bubble_plot <- function(df, order_x, order_y, label_x, label_y, auroc_min){
  
  # print(head(df))
  bp.x <- NULL
  bp.y <- NULL
  hm.clean <- NULL
  p <- NULL
  
  # x barplot
  print("barplot x")
  bp.x <- ggplot(
    data = df %>% select(var_x, ncells_x) %>% distinct(),
    aes(x = factor(var_x, levels = order_x), y = ncells_x)
  ) +
    geom_bar(
      stat = "identity",
      fill = "#bdbdbd"
    ) +
    theme_cowplot() +
    scale_x_discrete(position = "top") +
    labs(y = "#cells") +
    theme(
      axis.text = element_text(size = 15),
      axis.text.x = element_text(angle = 45, hjust = 0),
      # axis.ticks.y = element_blank(),
      axis.title.y = element_text(size = 20),
      axis.title.x = element_blank(),
      axis.line.x = element_blank(),
      legend.position = "none",
      plot.margin = margin(10, 0, 0, 0)
    )
  
  # PROPORTION OF MOUSE GDT CELLS IN EACH CLUSTER
  print("barplot y")
  bp.y <- ggplot(
    data = df %>% select(var_y, ncells_y) %>% distinct(),
    aes(x = factor(var_y, levels = order_y), y = ncells_y)
  ) +
    geom_bar(
      stat = "identity",
      fill = "#bdbdbd"
    ) +
    theme_cowplot() +
    scale_x_discrete(position = "top") +
    labs(y = "#cells") +
    coord_flip() +
    theme(
      axis.title.y = element_blank(),
      axis.text = element_text(size = 15),
      axis.text.x = element_text(angle = 45, hjust = 1),
      # axis.ticks.x = element_blank(),
      axis.title.x = element_text(size = 20),
      axis.line.y = element_blank(),
      legend.position = "none"
    )
  
  # BUBBLE PLOT
  print("bubble plot")
  hm.clean <- ggplot(df, aes(
    x = factor(var_x, levels = order_x),
    y = factor(var_y, levels = order_y)
  )) +
    # geom_point(aes(size = auroc, color= auroc))+
    geom_point(aes(size = abs(auroc - 0.5), color = auroc)) +
    geom_text(
      data = df %>% filter(auroc > auroc_min) %>% mutate(across("auroc", \(x) round(x, 2))),
      aes(label = auroc),
      color = "white"
    ) +
    scale_size_continuous(
      limits = c(0, 0.5),
      breaks = seq(0, 0.5, by = 0.1),
      range = c(1, 15)
    ) +
    scale_color_gradient2(
      low = "#2166ac",
      mid = "white",
      high = "#a50f15",
      midpoint = 0.5,
      limits = c(0, 1),
      name = "AUROC",
      breaks = seq(0, 1, by = 0.2)
    ) +
    labs(x = label_x, y = label_y, size = "AUROC") +
    theme_cowplot() +
    theme(
      legend.position = "bottom",
      legend.key.width = unit(0.8, 'cm'),
      axis.text = element_text(size = 15),
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 20)
    )
  
  
  # COMBINE
  print("combine")
  p <- (bp.x + plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
  return(p)
}
```


## Prepare data for plotting

### Prepare metadata
This is the same code chunk as in the script [pymtn_crosspecies_02_prepare_data.Rmd](./pymtn_crosspecies_02_prepare_data.Rmd).
```{r prep-metadata}
# Mouse
ms.metadata <- seur.ms@meta.data %>%
  dplyr::select(c(tcell_lineage, clusters_integrated_data)) %>%
  mutate(
    clusters_to_compare = case_when(
      clusters_integrated_data == "0" ~ "post_selection",
      clusters_integrated_data == "1" ~ "immature_CD24_GD",
      clusters_integrated_data == "2" ~ "immature_CD24_GD",
      clusters_integrated_data == "3" ~ "signaling",
      clusters_integrated_data == "4" ~ "signaling",
      clusters_integrated_data == "5" ~ "cycling",
      clusters_integrated_data == "6" ~ "transition",
      clusters_integrated_data == "7" ~ "typeII",
      clusters_integrated_data == "8" ~ "typeI",
      clusters_integrated_data == "9" ~ "typeI",
      clusters_integrated_data == "10" ~ "typeIII",
      clusters_integrated_data == "11" ~ "typeIII",
      clusters_integrated_data == "12" ~ "12"
    ),
    study = "mouse"
  )
# table(ms.metadata[,c("clusters_integrated_data", "clusters_to_compare")], useNA="ifany") # sanity check

# Human
hu.metadata <- rbind(
  seur.hu$nkt@meta.data[,c("tcell_lineage", "clusters_per_lineage")],
  seur.hu$mait@meta.data[,c("tcell_lineage", "clusters_per_lineage")],
  seur.hu$gdt@meta.data[,c("tcell_lineage", "clusters_per_lineage")]
) %>%
  mutate(
    study="human",
    clusters_to_compare=clusters_per_lineage
    )

head(ms.metadata)
head(hu.metadata)

# keep only columns which will be necessary for metaneighbor
# ms.metadata <- ms.metadata %>% dplyr::select(study, clusters_to_compare)
# hu.metadata <- hu.metadata %>% dplyr::select(study, clusters_to_compare)
```


### Get number of cells per cluster
First, we will identify clusters that represent more than 1% of number of cells in each species.
```{r prep-nb-cells-per-cluster}
# Human
hu_clusters_to_keep <- hu.metadata %>%
  summarise(ncells=n(), .by=clusters_to_compare) %>%
  mutate(
    totalcells=sum(ncells),
    percentcells=ncells*100/totalcells
  ) %>%
  filter(percentcells>1)
head(hu_clusters_to_keep)

# Mouse
ms_clusters_to_keep <- ms.metadata %>%
  summarise(ncells=n(), .by=clusters_to_compare) %>%
  mutate(
    totalcells=sum(ncells),
    percentcells=ncells*100/totalcells
  ) %>%
  filter(percentcells>1)
head(ms_clusters_to_keep)
```

Then, we will generate a dataframe to feed into the `bubble_plot` function, which will generate Fig 6B.
```{r prep-df-for-plotting}
mtn.df_mshu <- melt(mtn) %>%
  # keep human clusters in Var1 and murine clusters in Var2
  filter(str_detect(Var1,"human")) %>%
  mutate(Var1 = gsub("human\\|", "", Var1)) %>%
  filter(str_detect(Var2, "mouse|Mouse")) %>%
  mutate(Var2 = gsub("mouse|Mouse\\.", "", Var2)) %>%
  # add nb of human cells to dataframe
  left_join(
    as.data.frame(table(hu.metadata$clusters_to_compare)),
    by="Var1"
    ) %>%
  # rename a few columns
  rename(
    var_x = Var1,
    Var1 = Var2,
    auroc = value,
    ncells_x = Freq
  ) %>%
  # add nb of mouse cells to dataframe
  left_join(
    ms_clusters_to_keep %>% select(clusters_to_compare, ncells) %>% rename(Freq=ncells),
    by=c("Var1"="clusters_to_compare")
    ) %>%
  # rename a few columns again
  rename(
    var_y=Var1,
    ncells_y=Freq
    ) %>%
  # keep min groups of cells that represent at least 1% of total number of cells in species
  filter(
    var_x %in% hu_clusters_to_keep$clusters_to_compare & var_y %in% ms_clusters_to_keep$clusters_to_compare
    ) %>%
  # rename a few clusters
  mutate(
    var_x=str_replace(var_x, "_thymus", ""),
    var_y=case_when(
      var_y=="immature_CD24_GD" ~ "immature CD24+ GD",
      var_y=="post_selection" ~ "post-selection/naive",
      .default=var_y
    )
  )

# sanity check
mtn.df_mshu
```


## Plot Figure 6B
```{r plot-mtn, fig.width=15, fig.height=10}
order_mouse <- rev(c(
    "signaling",
    "post-selection/naive",
    "immature CD24+ GD",
    "cycling",
    "transition",
    "typeI",
    "typeII",
    "typeIII",
    "12"
))
order_human <- c(
  paste0("iNKT_c", 0:6),
  paste0("MAIT_c", 0:6),
  paste0("GDT_c", 0:7)
)

bubble_plot(
  df=mtn.df_mshu,
  order_x = order_human[order_human %in% unique(mtn.df_mshu$var_x)],
  order_y = order_mouse[order_mouse %in% unique(mtn.df_mshu$var_y)],
  label_x = "Human clusters",
  label_y = "Mouse clusters",
  auroc_min= 0.65
)
```

To note, some of the AUROC scores are slightly different from Fig6B in manuscript (e.g. `iNKT_c5 x typeIII` has an AUROC score of 0.68 here, instead of 0.69 in Fig6B) because when cleaning up our scripts for this analysis, we decided to download a more recent orthologs table from the biomart website (downloaded in Feb 2024, compared to Sept 2022 to generate Fig6B). However, as you can observe, the differences are very minor and don't impact the conclusions made in our manuscript.




# SESSION INFO
```{r}
sessionInfo()
```