---
title: "Chapter X - R Notebook"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# Import human data
seur_human <- readRDS("./data/clean_data/seurat_human_integrated_object_23_12_01.rds")

# Import mouse data
seur_tinn_mouse <- readRDS("./data/cross-species/08_innateT_cross_species/Analysis_all_mouse-Tinn_filtered_seurat_MNN.rds")
DimPlot(seur_tinn_mouse,  group.by="New_clusters", label=T)

# Import GEP genes
genes.gep <- read.csv("./data/human-PBMC/HumanData_27_ComparisonGEPsGarner/limited_nonimputed_genes_per_gep_post_rank_threshold_k12.csv", row.names=1)
genes.gep <- gather(genes.gep, key=gep, value=gene, colnames(genes.gep)) %>%
  filter(!is.na(gene)) %>%
  mutate(gep=sub("_", "", gep))
# head(genes.gep)
# table(genes.gep$gep, useNA="ifany")
# table(is.na(genes.gep$gene))
```




# FUNCTIONS
```{r define-functions}
lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  min_gene_score <- min(seuratobj@meta.data[,genesignature])
  min_cutoff <- 0
  if(min_gene_score>0){min_cutoff <- min_gene_score}
  
  p <- do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=min_cutoff,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  )
  # p <- ggrastr::rasterise(
  #   p,
  #   layers="Point",
  #   dpi=300
  # )
  print(p)
  # if(file_name != "no"){
  #   ggsave(paste0("~/Projects/phd/data/figures/chapter_04/figs_unfinished/section1/", file_name),
  #          plot=p,
  #          width=9, height=6)
  # }
}

plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(axis.text.y=element_text(face="italic"),
          axis.text.x=element_text(angle=45, hjust=1))+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}
```




# SCORE GEPs ON MOUSE OBJECT

## Get orthologous genes
```{r gep-orthologous}
# import ortholog table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_full.csv")

# check how many GEP genes are found in ortholog table
table(genes.gep$gene %in% ortholog.df$hu_symbol, useNA="ifany")

# subset ortholog table to GEP genes
ortholog.df <- ortholog.df %>% filter(hu_symbol %in% genes.gep$gene)

# translate GEP genes to ms and keep only mouse genes that are 
genes.gep_translated <- genes.gep %>%
  rename(hu_symbol=gene) %>%
  left_join(
    ortholog.df %>% select(ms_symbol, hu_symbol),
    by="hu_symbol"
    ) %>%
  filter(ms_symbol %in% rownames(seur_tinn_mouse))
```


## Make into list
```{r gep-gene-lists}
# human
gep_list_hu <- lapply(split(genes.gep_translated, genes.gep_translated$gep), function(x) x %>% pull(hu_symbol) %>% unique())
lapply(gep_list_hu, function(x) length(x))

# mouse
gep_list_ms <- lapply(split(genes.gep_translated, genes.gep_translated$gep), function(x) x %>% pull(ms_symbol) %>% unique())
lapply(gep_list_ms, function(x) length(x))
```


## Score on human and mouse datasets
```{r score-gene-list}
seur_human@meta.data[32:43] <- NULL
seur_tinn_mouse@meta.data[31:42] <- NULL

# score
seur_human <- AddModuleScore(seur_human, name = names(gep_list_hu), features=gep_list_hu, seed=1)
seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = names(gep_list_ms), features=gep_list_ms, seed=1, ctrl=500)


# Remove the annoying numbers that are being added
colnames(seur_human@meta.data)[32:43] <- names(gep_list_hu)
colnames(seur_tinn_mouse@meta.data)[31:42] <- names(gep_list_ms)
```


## Plot scores
```{r plot-human-geps, fig.width=6, fig.height=5}
for(i in paste0("GEP", 1:12)){
  lineage_genesignature(seur_human, i)
}
```


```{r plot-mouse-geps, fig.width=6, fig.height=5}
for(i in paste0("GEP", 1:12)){
  lineage_genesignature(seur_tinn_mouse, i)
}
```



## Try scoring GEP5 with different number of control genes
```{r score-gep5-mouse-with-diff-nb-ctrl-genes}
# seur_tinn_mouse@meta.data[43:44] <- NULL
seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = "GEP5_ctrl100", features=gep_list_ms[8], seed=1, ctrl=100)
seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = "GEP5_ctrl500", features=gep_list_ms[8], seed=1, ctrl=500)
# seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = "GEP5_ctrl1000", features=gep_list_ms[8], seed=1, ctrl=1000)
# seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = "GEP5_ctrl2000", features=gep_list_ms[8], seed=1, ctrl=2000)
colnames(seur_tinn_mouse@meta.data)[43:44] <- c("GEP5_ctrl100", "GEP5_ctrl500")

ggplot(
  as.data.frame(seur_tinn_mouse@meta.data)[,c("GEP5_ctrl100", "GEP5_ctrl500")],
  aes(x=GEP5_ctrl100, y=GEP5_ctrl500)
)+
  geom_point()
```

Surprisingly, it seems like GEP5 is not very highly expressed in mouse Tinn integrated dataset. This is not due to number of control genes chosen. Let's check out by bubble plot the expression level of the GEP5 genes in the mouse Tinn dataset.

Let's check out the distribution of average expression of all genes vs GEP5 genes.
```{r}
# Human
hu_avg_exp <- data.frame("avgexp"=rowMeans(seur_human@assays$RNA@data))
hu_avg_exp$gep5_gene <- ifelse(rownames(hu_avg_exp) %in% gep_list_hu$GEP5, T, F)
# table(rownames(hu_avg_exp[hu_avg_exp$gep5_gene==T,]) %in% gep_list_hu$GEP5, useNA="ifany")
hist(hu_avg_exp$avgexp, breaks=100)
hist(hu_avg_exp$avgexp, breaks=100)

# Mouse
ms_avg_exp <- data.frame("avgexp"=rowMeans(seur_tinn_mouse@assays$RNA@data))
ms_avg_exp$gep5_gene <- ifelse(rownames(ms_avg_exp) %in% gep_list_ms$GEP5, T, F)
# table(rownames(ms_avg_exp[ms_avg_exp$gep5_gene==T,]) %in% gep_list_ms$GEP5, useNA="ifany")

# Plot
ggplot(hu_avg_exp, aes(x=gep5_gene, y=avgexp))+
  geom_boxplot()+
  labs(title="Human")+
  ylim(0,2)
ggplot(ms_avg_exp, aes(x=gep5_gene, y=avgexp))+
  geom_boxplot()+
  labs(title="Mouse")+
  ylim(0,2)
```


## Score GEP5 only on 1:1 orthologs with orthology confidence of 1
```{r score-gep5-one2one, fig.width=6, fig.height=5}
genes.gep_translated.one2one <- genes.gep %>%
  rename(hu_symbol=gene) %>%
  left_join(
    ortholog.df %>% filter(homology_type=="ortholog_one2one" & orthology_confidence==1) %>% select(ms_symbol, hu_symbol),
    by="hu_symbol"
    ) %>%
  filter(ms_symbol %in% rownames(seur_tinn_mouse))

# Make into list
gep5_one2one_hu <- genes.gep_translated.one2one %>% filter(gep=="GEP5") %>% pull(hu_symbol) %>% unique()
gep5_one2one_ms <- genes.gep_translated.one2one %>% filter(gep=="GEP5") %>% pull(ms_symbol) %>% unique()
length(gep5_one2one_hu)
length(gep5_one2one_ms)

# Score on mouse object
seur_tinn_mouse <- AddModuleScore(seur_tinn_mouse, name = "GEP5_one2one", features=list("gep5"=gep5_one2one_ms), seed=1, ctrl=100)
colnames(seur_tinn_mouse@meta.data)[45] <- "GEP5_one2one"
ggplot(
  as.data.frame(seur_tinn_mouse@meta.data)[,c("GEP5_ctrl100", "GEP5_one2one")],
  aes(x=GEP5_ctrl100, y=GEP5_one2one)
 )+
  geom_point()
lineage_genesignature(seur_tinn_mouse, "GEP5_one2one")
```




## BUBBLE PLOT
```{r bubbleplot-gep5-human, fig.height=7, fig.width=5}
plot_DotPlot(seur_human, group="clusters_integrated_data", features=sort(gep5_one2one_hu)[1:50], scaling=T)
plot_DotPlot(seur_human, group="clusters_integrated_data", features=sort(gep5_one2one_hu)[51:100], scaling=T)
plot_DotPlot(seur_human, group="clusters_integrated_data", features=sort(gep5_one2one_hu)[101:150], scaling=T)
plot_DotPlot(seur_human, group="clusters_integrated_data", features=sort(gep5_one2one_hu)[151:179], scaling=T)
```

```{r bubbleplot-gep5-mouse, fig.height=7, fig.width=5}
plot_DotPlot(seur_tinn_mouse, group="New_clusters", features=sort(gep5_one2one_ms)[1:50], scaling=T)
plot_DotPlot(seur_tinn_mouse, group="New_clusters", features=sort(gep5_one2one_ms)[51:100], scaling=T)
plot_DotPlot(seur_tinn_mouse, group="New_clusters", features=sort(gep5_one2one_ms)[101:150], scaling=T)
plot_DotPlot(seur_tinn_mouse, group="New_clusters", features=sort(gep5_one2one_ms)[151:179], scaling=T)
```



Let's compare the distribution of %cells expressing GEP5 genes.
```{r percent-cells-expressing-GEP5-genes}
# human
hu_percent_exp <- data.frame("hu_ncells_exp"=rowSums(seur_human@assays$RNA@data>0)) %>%
  rownames_to_column("hu_symbol") %>%
  filter(hu_symbol %in% gep_list_hu$GEP5) %>%
  mutate("hu_percent_exp"=hu_ncells_exp*100/ncol(seur_human))

# mouse
ms_percent_exp <- data.frame("ms_ncells_exp"=rowSums(seur_tinn_mouse@assays$RNA@data>0)) %>%
  rownames_to_column("ms_symbol") %>%
  filter(ms_symbol %in% gep_list_ms$GEP5) %>%
  mutate("ms_percent_exp"=ms_ncells_exp*100/ncol(seur_tinn_mouse))

# combine by ortholog genes
percent_exp.df <- hu_percent_exp %>%
  left_join(
    genes.gep_translated %>% select(hu_symbol, ms_symbol),
    by="hu_symbol"
  ) %>%
  left_join(
    ms_percent_exp,
    by="ms_symbol"
  )

# plot
hist(hu_percent_exp$hu_percent_exp, breaks=100)
hist(ms_percent_exp$ms_percent_exp, breaks=100)
ggplot(percent_exp.df, aes(x=hu_percent_exp, y=ms_percent_exp))+
  geom_point()+
  geom_abline()+
  scale_x_log10()+
  scale_y_log10()

# how many genes are expressed in less than 1% of cells?
ggplot(
  rbind(
    data.frame(
      "species"="human",
      "percent_exp"=hu_percent_exp$hu_percent_exp
    ),
    data.frame(
      "species"="mouse",
      "percent_exp"=ms_percent_exp$ms_percent_exp
    )
  ),
  aes(x=percent_exp, colour=species)
)+
  geom_histogram(bins=100)
```




# SESSION INFO
```{r}
sessionInfo()
```


