---
title: "Chapter X - R Notebook"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/clean_data/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", group.by="clusters_integrated_data", cols=cols_integrated)

seur_thym <- list(
  "nkt"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.nkt.RDS"),
  "mait"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.mait.RDS"),
  "gdt"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.gd.RDS")
)
```




# FUNCTIONS
```{r define-functions}
# Set color matrix
BlendMatrix <- function(
    n = 10,
    col.threshold = 0.5,
    two.colors = c("#ff0000", "#00ff00"),
    negative.color = "black"
) {
  if (0 > col.threshold || col.threshold > 1) {
    stop("col.threshold must be between 0 and 1")
  }
  C0 <- as.vector(col2rgb(negative.color, alpha = TRUE))
  C1 <- as.vector(col2rgb(two.colors[1], alpha = TRUE))
  C2 <- as.vector(col2rgb(two.colors[2], alpha = TRUE))
  blend_alpha <- (C1[4] + C2[4])/2
  C0 <- C0[-4]
  C1 <- C1[-4]
  C2 <- C2[-4]
  merge.weight <- min(255 / (C1 + C2 +  C0 + 0.01))
  sigmoid <- function(x) {
    return(1 / (1 + exp(-x)))
  }
  blend_color <- function(
    i,
    j,
    col.threshold,
    n,
    C0,
    C1,
    C2,
    alpha,
    merge.weight
  ) {
    c.min <- sigmoid(5 * (1 / n - col.threshold)) # 5*
    c.max <- sigmoid(5 * (1 - col.threshold)) # 5*
    c1_weight <- sigmoid(5 * (i / n - col.threshold)) # 5*
    c2_weight <- sigmoid(5 * (j / n - col.threshold)) # 5*
    c0_weight <-  sigmoid(5 * ((i + j) / (2 * n) - col.threshold)) # 5*
    c1_weight <- (c1_weight - c.min) / (c.max - c.min)
    c2_weight <- (c2_weight - c.min) / (c.max - c.min)
    c0_weight <- (c0_weight - c.min) / (c.max - c.min)
    C1_length <- sqrt(sum((C1 - C0) ** 2))
    C2_length <- sqrt(sum((C2 - C0) ** 2))
    C1_unit <- (C1 - C0) / C1_length
    C2_unit <- (C2 - C0) / C2_length
    C1_weight <- C1_unit * c1_weight
    C2_weight <- C2_unit * c2_weight
    C_blend <- C1_weight * (i - 1) * C1_length / (n - 1) + C2_weight * (j - 1) * C2_length / (n - 1) + (i - 1) * (j - 1) * c0_weight * C0 / (n - 1) ** 2 + C0
    C_blend[C_blend > 255] <- 255
    C_blend[C_blend < 0] <- 0
    return(rgb(
      red = C_blend[1],
      green = C_blend[2],
      blue = C_blend[3],
      alpha = alpha,
      maxColorValue = 255
    ))
  }
  blend_matrix <- matrix(nrow = n, ncol = n)
  for (i in 1:n) {
    for (j in 1:n) {
      blend_matrix[i, j] <- blend_color(
        i = i,
        j = j,
        col.threshold = col.threshold,
        n = n,
        C0 = C0,
        C1 = C1,
        C2 = C2,
        alpha = blend_alpha,
        merge.weight = merge.weight
      )
    }
  }
  return(blend_matrix)
}

# Plot color matrix
Melt <- function(x) {
  if (!is.data.frame(x = x)) {
    x <- as.data.frame(x = x)
  }
  return(data.frame(
    rows = rep.int(x = rownames(x = x), times = ncol(x = x)),
    cols = unlist(x = lapply(X = colnames(x = x), FUN = rep.int, times = nrow(x = x))),
    vals = unlist(x = x, use.names = FALSE)
  ))
}

BlendMap <- function(color.matrix, step=2, xtext='rows', ytext="cols") {
  color.heat <- matrix(
    data = 1:prod(dim(x = color.matrix)) - 1,
    nrow = nrow(x = color.matrix),
    ncol = ncol(x = color.matrix),
    dimnames = list(
      1:nrow(x = color.matrix),
      1:ncol(x = color.matrix)
    )
  )
  
  # xbreaks <- seq.int(from = 0, to = nrow(x = color.matrix), by = step)
  # ybreaks <- seq.int(from = 0, to = ncol(x = color.matrix), by = step)
  color.heat <- Melt(x = color.heat)
  color.heat$rows <- as.numeric(x = as.character(x = color.heat$rows))
  color.heat$cols <- as.numeric(x = as.character(x = color.heat$cols))
  color.heat$vals <- factor(x = color.heat$vals)
  plot <- ggplot(
    data = color.heat,
    mapping = aes_string(x = "rows", y = "cols", fill = 'vals')
  ) +
    geom_raster(show.legend = FALSE) +
    theme(plot.margin = unit(x = rep.int(x = 0, times = 4), units = 'cm')) +
    # scale_x_continuous(breaks = xbreaks, expand = c(0, 0), labels = xbreaks) +
    # scale_y_continuous(breaks = ybreaks, expand = c(0, 0), labels = ybreaks) +
    scale_fill_manual(values = as.vector(x = color.matrix)) +
    labs(x=xtext, y=ytext)+
    theme_cowplot()
  
  if(step!=0){
    xbreaks <- seq.int(from = 0, to = nrow(x = color.matrix), by = step)
    ybreaks <- seq.int(from = 0, to = ncol(x = color.matrix), by = step)
    plot <- plot +
      scale_x_continuous(breaks = xbreaks, expand = c(0, 0), labels = xbreaks) +
      scale_y_continuous(breaks = ybreaks, expand = c(0, 0), labels = ybreaks)
  } else if(step==0){
    plot <- plot+theme(axis.text=element_blank(), axis.ticks=element_blank())
  }
  
  return(plot)
}

# Normalize expression level of 2 features & return also a "blend" expression (corresponding to color matrix)
BlendExpression <- function(data, nlevels=100) {
  if (ncol(x = data) != 2) {
    stop("'BlendExpression' only blends two features")
  }
  features <- colnames(x = data)
  data <- as.data.frame(x = apply(
    X = data,
    MARGIN = 2,
    FUN = function(x) {
      return(round(x = (nlevels-1) * (x - min(x)) / (max(x) - min(x))))
    }
  ))
  data[, 3] <- data[, 1] + data[, 2] * nlevels
  # colnames(x = data) <- c(features, paste(features, collapse = '_'))
  colnames(x = data) <- c(features, "blend")
  for (i in 1:ncol(x = data)) {
    data[, i] <- factor(x = data[, i])
  }
  return(data)
}

PlotCoexpression <- function(seuratobj,
                             features,
                             reduction_name="umap",
                             plotting="blend",
                             pt_size=1,
                             set_negative_scores_to_0=F,
                             pwithmatrix=T,
                             matrix_coordinates=c(0.05, 0.06),
                             rasterdpi=300,
                             nlevels=100,
                             cols.neg="#969696", cols.pos=c("#74c476", "#fd8d3c"), col.threshold=0.5, colmatrix_stepsize=10,
                             order=T){
  # GET COLOR MATRIX
  # cat("\n-> Getting color matrix\n")
  color.matrix <- BlendMatrix(
    two.colors = cols.pos,
    col.threshold = col.threshold,
    negative.color = cols.neg,
    n=nlevels
  )
  
  # DEFINE COLOR LIST FOR PLOTTING
  # cat("\n-> Defining colors for plotting\n")
  # colors <- list(
  #   color.matrix[, 1], # red
  #   color.matrix[1, ], # green
  #   as.vector(x = color.matrix)
  # )
  

  # BLEND EXPRESSION
  # cat("\n-> Blending features expression\n")
  df <- seuratobj@meta.data[, features]
  if(set_negative_scores_to_0==T){
    df[df<0] <- 0
  }
  df <- BlendExpression(df, nlevels=nlevels) # 3 columns
  # head(df)
  # GET PLOTTING DATAFRAME
  # cat("\n-> Defining plotting DF\n")
  dims <- as.data.frame(seuratobj@reductions[[reduction_name]]@cell.embeddings)
  # head(dims)
  df_final <- merge(df, dims, by="row.names")
  rownames(df_final) <- df_final$Row.names
  df_final$Row.names <- NULL
  # print(tail(df_final))
  
  # PLOT
  if(plotting=="feature1"){
    # cat("\n-> Plotting feature 1\n")
    if(order==T){df_final <- df_final[order(df_final[,1]),]}
    df_final[,1] <- as.numeric(as.character(df_final[,1])) # transform factors to numbers for plotting
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[1]))+
      geom_point(size=pt_size)+
      scale_color_gradient(low=color.matrix[1, 1], high=color.matrix[nrow(color.matrix), 1])
  }
  else if(plotting=="feature2"){
    # cat("\n-> Plotting feature 2\n")
    if(order==T){df_final <- df_final[order(df_final[,2]),]}
    df_final[,2] <- as.numeric(as.character(df_final[,2])) # transform factors to numbers for plotting
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[2]))+
      geom_point(size=pt_size)+
      scale_color_gradient(low=color.matrix[1, 1], high=color.matrix[1, ncol(color.matrix)])
  }
  else if(plotting=="blend"){
    # cat("\n-> Plotting blended features\n")
    if(order==T){
      # order points by increasing value of score 1 + score 2 (coexpression)
      df_final <- df_final[order(as.numeric(as.character(df_final[,1]))+as.numeric(as.character(df_final[,2]))),] 
      # print(tail(df_final))
      }
    # Colors
    cols.use <- as.vector(color.matrix)
    names(cols.use) <- as.character(0:(length(cols.use)-1))
    # Plot
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[3]))+
      geom_point(size=pt_size)+
      scale_color_manual(values=cols.use)+
      labs(x="UMAP1", y="UMAP2")+
      theme_cowplot()+
      theme(legend.position="none",
            axis.text=element_blank(),
            axis.title=element_blank(),
            axis.ticks = element_blank(),
            axis.line = element_blank(),
            panel.border=element_rect(color="white", fill=NA, size=1))
    # rasterise to avoid computer crashing
    p <- ggrastr::rasterise(p, layers="Point", dpi=rasterdpi)
    # add color matrix if specified
    if(pwithmatrix==T){
      # cat("\n-> Adding color matrix on plot\n")
      p <- ggdraw(p)+
        draw_plot(
          BlendMap(color.matrix, step=colmatrix_stepsize, xtext=features[1], ytext=features[2]),
          x=matrix_coordinates[1],y=matrix_coordinates[2],.25,.25
          )
    }
  }
  
  return(p)
}
```




# ANALYSIS

## Score gene signatures
```{r score-gene-signatures}
gene_signatures <- list(
  "type1"=c("EOMES", "GZMK", "CCL5", "TBX21", "NKG7", "GZMA"),
  "type17"=c("RORC", "RORA", "CCR6", "IL23R", "BLK", "SCART1")
)

# score
seur_integrated <- AddModuleScore(seur_integrated, features=gene_signatures, name=names(gene_signatures), seed=1)
colnames(seur_integrated@meta.data)[32:33] <- names(gene_signatures)
seur_thym <- lapply(seur_thym, function(x) AddModuleScore(x, features=gene_signatures, name=names(gene_signatures), seed=1))
colnames(seur_thym$nkt@meta.data)[78:79] <- names(gene_signatures)
colnames(seur_thym$mait@meta.data)[78:79] <- names(gene_signatures)
colnames(seur_thym$gdt@meta.data)[78:79] <- names(gene_signatures)

# add the type 1 and type 17 scored on integrated object in the individual thymic objects
seur_thym$nkt@meta.data[,c("type1_integrated", "type17_integrated")] <- seur_integrated@meta.data[rownames(seur_integrated@meta.data) %in% rownames(seur_thym$nkt@meta.data),c("type1", "type17")]
seur_thym$mait@meta.data[,c("type1_integrated", "type17_integrated")] <- seur_integrated@meta.data[rownames(seur_integrated@meta.data) %in% rownames(seur_thym$mait@meta.data),c("type1", "type17")]
seur_thym$gdt@meta.data[,c("type1_integrated", "type17_integrated")] <- seur_integrated@meta.data[rownames(seur_integrated@meta.data) %in% rownames(seur_thym$gdt@meta.data),c("type1", "type17")]
```

Check how the type1 and type17 scores correlate when they're scored on individual thymic seurat objects versus on the integrated seurat object.
```{r check-gene-signatures-correlation}
# make dataframe
df <- rbind(
  as.data.frame(seur_thym$nkt@meta.data) %>%
    select(type1, type1_integrated, type17, type17_integrated, cell_annot) %>%
    mutate(tcell_lineage="iNKT"),
  as.data.frame(seur_thym$mait@meta.data) %>%
    select(type1, type1_integrated, type17, type17_integrated, cell_annot) %>%
    mutate(tcell_lineage="MAIT"),
  as.data.frame(seur_thym$gdt@meta.data) %>%
    select(type1, type1_integrated, type17, type17_integrated, cell_annot) %>%
    mutate(tcell_lineage="GD")
) %>%
  dplyr::rename(cluster_per_lineage=cell_annot)

ggplot(df, aes(x=type1, y=type1_integrated))+
  geom_point()+
  facet_wrap(~tcell_lineage)+
  labs(x="Signature scored on thymic object", y="Signature scored on integrated object", title="Type 1 signature")

ggplot(df, aes(x=type17, y=type17_integrated))+
  geom_point()+
  facet_wrap(~tcell_lineage)+
  labs(x="Signature scored on thymic object", y="Signature scored on integrated object", title="Type 17 signature")
```

Overall the scores correlate pretty well, the only difference is that scores from integrated object are lower than scores from separate thymic objects (especially for type 1 signature). Theoretically, scoring the gene signatures on the integrated object would be better as it contains a more diverse set of cells, and so it would provide better "background" genes (for control). Let's keep the scores from integrated object.

```{r}
# remove scores computed on each seurat object individually
seur_thym$nkt@meta.data[,c("type1", "type17")] <- NULL
seur_thym$mait@meta.data[,c("type1", "type17")] <- NULL
seur_thym$gdt@meta.data[,c("type1", "type17")] <- NULL

# rename scores computed on integrated seurat object
colnames(seur_thym$nkt@meta.data)[78:79] <- c("type1", "type17")
colnames(seur_thym$mait@meta.data)[78:79] <- c("type1", "type17")
colnames(seur_thym$gdt@meta.data)[78:79] <- c("type1", "type17")
```



## Plot coexpression of type 1/type 17 on scatter plot
```{r coexpression-scatterplot, fig.width=10, fig.height=5}
# update color vector names
names(cols_thym_nkt) <- gsub('iNKT_thymus_', 'NKT_', names(cols_thym_nkt))
names(cols_thym_mait) <- gsub('thymus_', '', names(cols_thym_mait))
names(cols_thym_gdt) <- gsub('thymus_', '', names(cols_thym_gdt))

ggplot(
  df %>% filter(cluster_per_lineage %in% c("NKT_c5", "NKT_c6", "MAIT_c6", "GDT_c7")),
  # df,
  aes(x=type1_integrated, y=type17_integrated, color=cluster_per_lineage)
)+
  geom_point()+
  facet_wrap(~factor(tcell_lineage, levels=c("iNKT", "MAIT", "GD")), ncol=1)+
  scale_color_manual(values=c(cols_thym_nkt, cols_thym_mait, cols_thym_gdt))+
  labs(x="Type 1 signature", y="Type 17 signature")+
  theme_cowplot()
ggsave("./scripts-in-progress/human-thymus/HumanThymus_26_coexpression_type1_type17/plots/scatterplot_coexpression_vertical.pdf", width=6, height=10)
```

Check what proportion of MAIT_c6 cells have a type 1 and type 17 signature above 0
```{r}
df %>%
  filter(cluster_per_lineage == "GDT_c7") %>%
  select(type1_integrated, type17_integrated) %>%
  mutate(type1_positivescore=type1_integrated>0,
         type17_positivescore=type17_integrated>0) %>%
  group_by(type1_positivescore, type17_positivescore) %>%
  count() %>%
  ungroup() %>%
  mutate(sum=sum(n),
         freq=n*100/sum)
```

So about 35% of MAIT_c6 cells have a positive type 1 and type 17 scores.



## Plot gene signatures
```{r plot-coexpression, fig.height=5, fig.width=5}
# iNKT
PlotCoexpression(
  seuratobj = seur_thym$nkt,
  features = c("type1", "type17"),
  set_negative_scores_to_0 = T, # whether to have negative signature scores set to 0 (negative scores mean that signature genes are expressed less than control genes)
  plotting = "blend",
  pt_size=2,
  matrix_coordinates = c(0, 0),
  cols.neg="#bdbdbd", # negative color
  cols.pos=c("red", "blue"),
  col.threshold=0.1, # at which threshold to switch from grey to color
  colmatrix_stepsize=100, # how many steps in the color matrix
  order=T # whether to order cells
)
# ggsave("./scripts-in-progress/human-thymus/HumanThymus_26_coexpression_type1_type17/plots/coexpression_nkt.pdf", width=5, height=5)

# MAIT
PlotCoexpression(
  seuratobj = seur_thym$mait,
  features = c("type1", "type17"),
  set_negative_scores_to_0 = T,
  plotting = "blend",
  matrix_coordinates = c(0.7, 0.06),
  cols.neg="#bdbdbd",
  cols.pos=c("red", "blue"),
  pt_size=2,
  col.threshold=0.01,
  colmatrix_stepsize=100,
  order=T
)
# ggsave("./scripts-in-progress/human-thymus/HumanThymus_26_coexpression_type1_type17/plots/coexpression_mait.pdf", width=5, height=5)


# GD
PlotCoexpression(
  seuratobj = seur_thym$gdt,
  features = c("type1", "type17"),
  set_negative_scores_to_0 = T,
  plotting = "blend",
  matrix_coordinates = c(0, 0),
  cols.neg="#bdbdbd",
  cols.pos=c("red", "blue"),
  pt_size=2,
  col.threshold=0.1,
  colmatrix_stepsize=100,
  order=T
)
# ggsave("./scripts-in-progress/human-thymus/HumanThymus_26_coexpression_type1_type17/plots/coexpression_gdt.pdf", width=5, height=5)
```


```{r, fig.height=5, fig.width=5}
PlotCoexpression(
  seuratobj = seur_thym$gdt,
  features = c("type1", "type17"),
  set_negative_scores_to_0 = T,
  plotting = "blend",
  matrix_coordinates = c(0.7, 0.06),
  pt_size = 1,
  cols.neg="#d9d9d9",
  cols.pos=c("red", "blue"),
  col.threshold=0.1,
  colmatrix_stepsize=100,
  order=T
)
```



Investigate how to reorder cells
```{r test_cell_ordering, fig.width=5, fig.height=5, include=FALSE}
# color matrix
test_color.matrix <- BlendMatrix(
  two.colors = c("red", "blue"),
  col.threshold = 0.1,
  negative.color = "#969696",
  n=100
)
test_cols.use <- as.vector(test_color.matrix)
names(test_cols.use) <- as.character(0:(length(test_cols.use)-1))


test <- seur_thym$mait@meta.data[, c("type1", "type17")]
test[test<0] <- 0
test <- BlendExpression(test, nlevels=100) # 3 columns
test_dims <- as.data.frame(seur_thym$mait@reductions[["umap"]]@cell.embeddings)
test_final <- merge(test, test_dims, by="row.names")
rownames(test_final) <- test_final$Row.names
test_final$Row.names <- NULL
print(tail(test_final))


# try different ordering
# test_final <- test_final[order(test_final[,1], test_final[,2]),] # order feature 1 then feature 2
# test_final <- test_final[order(test_final[,2], test_final[,1]),] # order feature 2 then 1
# test_final <- test_final[order(test_final[,3]),] # order by blend
test_final <- test_final[order(as.numeric(as.character(test_final[,1])) + as.numeric(as.character(test_final[,2]))),] 
tail(test_final)

# Plot
p <- ggplot(test_final,
            aes_string(x=colnames(test_dims)[1], y=colnames(test_dims)[2], color=colnames(test_final)[3]))+
  geom_point(size=1)+
  scale_color_manual(values=test_cols.use)+
  labs(x="UMAP1", y="UMAP2")+
  theme_cowplot()+
  theme(legend.position="none",
        axis.text=element_blank(),
        axis.title=element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.border=element_rect(color="white", fill=NA, size=1))+
  xlim(-8,20)+
  ylim(-7,6)
p <- ggdraw(p)+
  draw_plot(
    BlendMap(test_color.matrix, step=100, xtext="type1", ytext="type17"),
    x=0.5,y=0.06,.25,.25
    )
p
```


## Analysis 2
```{r analysis-2}

```

## Analysis 3
```{r analysis-3}

```


## Analysis 4
```{r analysis-4}

```




# SESSION INFO
```{r}
sessionInfo()
```


