---
title: "Chapter 2 - Figure 7"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(reshape2)
library(ggrepel)
library(ComplexHeatmap)

library(tidyverse)
library(dplyr)
library(Seurat)
# library(harmony)
# library(SCpubr)
library(MetaNeighbor)
library(SummarizedExperiment)
library(gplots)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)
```




# FUNCTIONS
```{r define-functions}
plot_mtn_heatmap <- function(
    mtn_matrix,
    vector_groups_with_min_cells,
    df_annotation,
    df_nbcells,
    grouping_labels=c("a", "b", "c"),
    heatmapclus=F
    ){
  
  # keep only groups that have at least 100 cells
  cat("Metaneighbor matrix dimensions:", dim(mtn_matrix), "\n") # 88 rows & columns
  mtn_sub <- mtn_matrix[rownames(mtn_matrix) %in% vector_groups_with_min_cells, colnames(mtn_matrix) %in% vector_groups_with_min_cells]
  cat("Metaneighbor matrix without groups < 100 cells:", dim(mtn_sub), "\n") # 43 rows and 43 columns
  
  # reorder rows & columns by cluster
  groups_order_rows <- df_annotation[rownames(mtn_sub),] %>% arrange(grouping1, grouping2)
  groups_order_cols <- df_annotation[colnames(mtn_sub),] %>% arrange(grouping1, grouping2)
  mtn_sub <- mtn_sub[rownames(groups_order_rows), rownames(groups_order_cols)]
  cat("\nOrder matrix rows:", rownames(mtn_sub), "\n\n")
  cat("Order matrix columns:", colnames(mtn_sub), "\n\n")
  
  # row annotation
  row_annot <-
    rowAnnotation(
      # annotations
      nbcells=anno_barplot(
        df_nbcells %>% filter(group %in% rownames(mtn_sub)) %>% arrange(match(group, rownames(mtn_sub))) %>% pull(n),
        axis_param = list(direction = "reverse")
        ),
      grouping1=df_annotation[rownames(mtn_sub),"grouping1"],
      grouping2=df_annotation[rownames(mtn_sub),"grouping2"],
      # annotation labels
      annotation_label=c(grouping_labels[1], grouping_labels[2], grouping_labels[3]),
      # annotation colors
      col=list(
        grouping1=deframe(df_annotation[rownames(mtn_sub),c("grouping1", "col_grouping1")]),
        grouping2=deframe(df_annotation[rownames(mtn_sub),c("grouping2", "col_grouping2")])
        )
      )
  
  # column annotation
  col_annot <-
    HeatmapAnnotation(
      # annotations
      nbcells=anno_barplot(
        df_nbcells %>% filter(group %in% colnames(mtn_sub)) %>% arrange(match(group, colnames(mtn_sub))) %>% pull(n)
        ),
      grouping1=df_annotation[colnames(mtn_sub),"grouping1"],
      grouping2=df_annotation[colnames(mtn_sub),"grouping2"],
      # annotation labels
      annotation_label=c(grouping_labels[1], grouping_labels[2], grouping_labels[3]),
      # annotation colors
      col=list(
        grouping1=deframe(df_annotation[colnames(mtn_sub),c("grouping1", "col_grouping1")]),
        grouping2=deframe(df_annotation[colnames(mtn_sub),c("grouping2", "col_grouping2")])
        )
      )
  
  # heatmap
  Heatmap(
    mtn_sub,
    name="auroc",
    # cluster
    cluster_rows = heatmapclus,
    cluster_columns = heatmapclus,
    # annotations
    top_annotation=col_annot,
    left_annotation=row_annot,
    # add values
    layer_fun = function(j, i, x, y, width, height, fill) {
          v = pindex(mtn_sub, i, j)
          l = v > 0.8 | v < 0.1
          grid.text(sprintf("%.2f", v[l]), x[l], y[l], gp = gpar(fontsize = 10))
          }
    )
}
```




# METANEIGHBOR BY LINEAGE (across donors)

## Prepare data
```{r mtn_branches_prepare_data}
# subset to thymocytes only
seur_thym <- subset(seur_integrated, subset=tissue=="Thymus")
print(seur_thym)

# create a new metadata column for "branches"
seur_thym@meta.data
seur_thym@meta.data$tcell_lineage_split <- dplyr::case_when(
  seur_thym@meta.data$tcell_lineage=="iNKT" & seur_thym@meta.data$clusters_integrated_data %in% 0:11  ~ "iNKT non-memory",
  seur_thym@meta.data$tcell_lineage=="iNKT" & seur_thym@meta.data$clusters_integrated_data %in% 12:17 ~ "iNKT memory",
  seur_thym@meta.data$tcell_lineage=="MAIT" & seur_thym@meta.data$clusters_integrated_data %in% 0:11  ~ "MAIT non-memory",
  seur_thym@meta.data$tcell_lineage=="MAIT" & seur_thym@meta.data$clusters_integrated_data %in% 12:17 ~ "MAIT memory",
  seur_thym@meta.data$tcell_lineage=="GD"   & seur_thym@meta.data$clusters_integrated_data %in% 0:11  ~ "GD non-memory",
  seur_thym@meta.data$tcell_lineage=="GD"   & seur_thym@meta.data$clusters_integrated_data %in% 12:17 ~ "GD memory",
  seur_thym@meta.data$tcell_lineage=="CD4" & seur_thym@meta.data$clusters_integrated_data %in% 0:11   ~ "CD4 non-memory",
  seur_thym@meta.data$tcell_lineage=="CD4" & seur_thym@meta.data$clusters_integrated_data %in% 12:17  ~ "CD4 memory",
  seur_thym@meta.data$tcell_lineage=="CD8" & seur_thym@meta.data$clusters_integrated_data %in% 0:11   ~ "CD8 non-memory",
  seur_thym@meta.data$tcell_lineage=="CD8" & seur_thym@meta.data$clusters_integrated_data %in% 12:17  ~ "CD8 memory",
  .default = NA
)
table(seur_thym@meta.data[,c("clusters_integrated_data", "tcell_lineage_split")], useNA="ifany") # sanity check
table(seur_thym@meta.data[,c("tcell_lineage_split", "tcell_lineage")], useNA="ifany") # sanity check


# get count matrix
se_thym <- SummarizedExperiment(
  assays=seur_thym@assays[["RNA"]]@counts,
  colData=seur_thym@meta.data[,c("tcell_lineage_split", "donor_id")]
  )
```


## Run metaneighbor
```{r run_mtn_branches}
mtn <- MetaNeighborUS(
  var_genes=VariableFeatures(seur_thym),
  dat=se_thym,
  study_id=se_thym$donor_id,
  cell_type=se_thym$tcell_lineage_split,
  fast_version=T
  )
```

```{r plot_mtn_branches_fullheatmap, fig.width=14, fig.height=14}
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```

## Plot metaneighbor full heatmap
First, let's obtain the number of cells per group (we are not interested in groups that contain less than 100 cells, it's less reliable).
```{r prepare_data_heatmaps_branches}
# Get number of cells per lineage and cluster
nbcells <- seur_thym@meta.data %>%
  as_tibble() %>%
  group_by(tcell_lineage_split, donor_id) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(group=paste(donor_id, tcell_lineage_split, sep="|")) %>%
  # get percent of cells per lineage (e.g. %MAIT in MAIT_effector_donor1)
  separate(tcell_lineage_split, c("tcell_lineage", "memoryB"), sep=" ", remove=F) %>%
  select(-memoryB) %>%
  group_by(tcell_lineage) %>%
  mutate(nbcells_per_lineage=sum(n)) %>%
  ungroup() %>%
  select(-tcell_lineage_split) %>%
  mutate(percentcells_per_lineage=n*100/nbcells_per_lineage) %>%
  # clean up
  select(-nbcells_per_lineage, -donor_id) %>%
  relocate(group)
groups_with_min_cells <- nbcells %>% filter(n>=100) %>% pull(group)

# general dataframe for annotation
df_annot <- data.frame("names"=nbcells$group) %>%
  separate(names, c("donor_id", "tcell_lineage_split"), sep="\\|", remove=F) %>%
  separate(tcell_lineage_split, c("tcell_lineage", "memoryB"), sep=" ", remove=F) %>%
  select(-memoryB) %>%
  mutate(col_donor=case_when(
    donor_id=="1" ~"#f7f7f7",
    donor_id=="2" ~"#cccccc",
    donor_id=="3" ~"#969696",
    donor_id=="4" ~"#525252"
  )) %>%
  left_join(enframe(cols_lineages) %>% dplyr::rename(col_lineages=value), by=join_by(tcell_lineage==name)) %>%
  column_to_rownames("names") %>%
  # rename columns for heatmap function
  dplyr::rename(grouping2=donor_id,
                grouping1=tcell_lineage,
                col_grouping2=col_donor,
                col_grouping1=col_lineages)

```

Then, we can plot the heatmap (full heatmap).
```{r heatmap_full, fig.width=12, fig.height=12}
pdf("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section1/ch2_fig7_mtn_lineage.pdf", width=12, height=10)
plot_mtn_heatmap(
  mtn_matrix=mtn,
  vector_groups_with_min_cells = groups_with_min_cells,
  df_annotation = df_annot,
  # df_nbcells=nbcells3,
  df_nbcells=nbcells %>% select(-n) %>% dplyr::rename(n=percentcells_per_lineage),
  grouping_labels=c("%cells", "lineage", "donor"),
  heatmapclus=T
  )
dev.off()
```

Look at distribution of AUROC score between CD4-iNKT and CD8-MAIT (non-effector)
```{r}
median(
  reshape2::melt(mtn) %>%
    filter(Var1 %in% grep("CD4 non-memory", Var1, value=T) & Var2 %in% grep("iNKT non-memory", Var2, value=T)) %>%
    pull(value)
  )

median(
  reshape2::melt(mtn) %>%
    filter(Var1 %in% grep("CD8 non-memory", Var1, value=T) & Var2 %in% grep("MAIT non-memory", Var2, value=T)) %>%
    pull(value)
  )
```






# SESSION INFO
```{r}
sessionInfo()
```