---
title: "Plot pyMetaNeighbor with Park et al. dataset (Supp Fig 4B)"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```




# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(RColorBrewer)
library(reshape2)
library(ggrepel)
library(patchwork)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```


## Import data
```{r import-data, fig.width=7}
# our integrated dataset
seur_integrated <- readRDS("./data_geo/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
seur_thymus <- subset(seur_integrated, subset=tissue=="Thymus") # keep only thymocytes
do_DimPlot(
  seur_thymus,
  group.by = "clusters_integrated_data",
  reduction = "umap_integrated",
  colors.use = cols_integrated,
  legend.position="right"
) # supp Fig 4A (top)


# Park et al. thymocyte data (corresponding to Fig 2C)
seur_park <- readRDS("./data_github/park_dataset/park_seurat_human_thymocytes.rds")
do_DimPlot(
  seur_park,
  group.by="cell_types",
  legend.position="right",
  colors.use = cols_park,
  label=T
) # supp Fig 4A (bottom)
```

Import output from metaneighbor. To note, to reproduce Supplementary Fig 4B, you will want to import the output from the slow version of metaneighbor (which needs to be run on a high-performing cluster with the script [pymtn_park_02_run_metaneighbor.py](./pymtn_park_02_run_metaneighbor.py), unless you have a super powerful computer unlike mine). As an alternative, you can also run on your local computer a fast version of metaneighbor (with the script [pymtn_park_02_run_metaneighbor.ipynb](./pymtn_park_02_run_metaneighbor.ipynb)) and import that `.csv` file as input (commented out below).
```{r import-data-pymtn}
mtn <- read.csv(
  # if run on HPC (reproduces Supp Fig 4B)
  "./data_github/park_dataset/pymtn/human_thym_merged_with_park_pymtn_slow_version_output_240902.csv",
  # if run on local computer
  # "./data_github/park_dataset/pymtn/human_thym_merged_with_park_pymtn_fast_version_output_240902.csv",
  row.names=1
  )
mtn <- as.matrix(mtn)
mtn[1:5,1:5]
```




# PLOT METANEIGHBOR

## Quick look at heatmap
```{r mtn-heatmap, fig.height=10, fig.width=10}
gplots::heatmap.2(
  mtn,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human innate T (3104 HVGs)",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(9, 9)
)
```


## Define plotting function
```{r define-function}
bubble_plot <- function(df, order_x, order_y, label_x, label_y, auroc_min=0.9){
  
  # print(head(df))
  bp.x <- NULL
  bp.y <- NULL
  hm.clean <- NULL
  p <- NULL
  
  # x barplot
  print("barplot x")
  bp.x <- ggplot(
    data = df %>% select(var_x, ncells_x) %>% distinct(),
    aes(x = factor(var_x, levels = order_x), y = ncells_x)
  ) +
    geom_bar(
      stat = "identity",
      fill = "#bdbdbd"
    ) +
    theme_cowplot() +
    scale_x_discrete(position = "top") +
    labs(y = "#cells") +
    theme(
      axis.text = element_text(size = 15),
      axis.title.y = element_text(size = 20),
      axis.title.x = element_blank(),
      axis.line.x = element_blank(),
      legend.position = "none",
      plot.margin = margin(10, 0, 0, 0)
    )
  
  # PROPORTION OF MOUSE GDT CELLS IN EACH CLUSTER
  print("barplot y")
  bp.y <- ggplot(
    data = df %>% select(var_y, ncells_y) %>% distinct(),
    aes(x = factor(var_y, levels = order_y), y = ncells_y)
  ) +
    geom_bar(
      stat = "identity",
      fill = "#bdbdbd"
    ) +
    theme_cowplot() +
    scale_x_discrete(position = "top") +
    labs(y = "#cells") +
    coord_flip() +
    theme(
      axis.title.y = element_blank(),
      axis.text = element_text(size = 15),
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title.x = element_text(size = 20),
      axis.line.y = element_blank(),
      legend.position = "none"
    )
  
  # BUBBLE PLOT
  print("bubble plot")
  hm.clean <- ggplot(df, aes(
    x = factor(var_x, levels = order_x),
    y = factor(var_y, levels = order_y)
  )) +
    geom_point(aes(size = auroc, color= auroc))+
    geom_text(
      data = df %>% filter(auroc > auroc_min) %>% mutate(across("auroc", \(x) round(x, 2))),
      aes(label = auroc),
      color = "white"
    ) +
    scale_size_continuous(
      limits = c(0, 1),
      breaks = seq(0.2, 0.8, by = 0.2),
      range = c(1, 15)
    ) +
    scale_color_gradient2(
      low="#d9d9d9",
      mid = "white",
      high = "#a50f15",
      midpoint = 0.5,
      limits = c(0, 1),
      name = "AUROC",
      breaks = seq(0, 1, by = 0.2)
    ) +
    labs(x = label_x, y = label_y, size = "AUROC") +
    theme_cowplot() +
    theme(
      legend.position = "bottom",
      legend.key.width = unit(0.8, 'cm'),
      axis.text = element_text(size = 15),
      axis.title = element_text(size = 20)
    )
  
  
  # COMBINE
  print("combine")
  p <- (bp.x + plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
  return(p)
}
```


## Prepare data for plotting

### Prepare metadata
This is the same code chunk as in the script [pymtn_park_01_prepare_data.Rmd](./pymtn_park_01_prepare_data.Rmd).
```{r}
# our data: keep only columns of interest
seur_thymus@meta.data <- seur_thymus@meta.data[,c(
  "nCount_RNA",
  "nFeature_RNA",
  "clusters_integrated_data"
)]
colnames(seur_thymus@meta.data) <- c("nCount", "nFeature", "clusters_to_compare")
seur_thymus@meta.data$clusters_to_compare <- as.character(seur_thymus@meta.data$clusters_to_compare)
seur_thymus@meta.data$clusters_to_compare <- paste0("c", seur_thymus@meta.data$clusters_to_compare)
seur_thymus@meta.data$study <- "gapin_data"
head(seur_thymus@meta.data)

# park data: keep only columns of interest
seur_park@meta.data <- seur_park@meta.data[,c(
  "nCount_originalexp",
  "nFeature_originalexp",
  "cell_types"
)]
colnames(seur_park@meta.data) <- c("nCount", "nFeature", "clusters_to_compare")
seur_park@meta.data$clusters_to_compare <- as.character(seur_park@meta.data$clusters_to_compare)
seur_park@meta.data$study <- "park_data"
head(seur_park@meta.data)
```


### Get number of cells per cluster
We will generate a dataframe to feed into the `bubble_plot` function, which will generate Supplementary Fig 4B.
```{r prep-nb-cells-per-cluster}
mtn.df <- melt(mtn) %>%
  filter(str_detect(Var1,"gapin_data")) %>%
  mutate(Var1 = gsub("gapin_data\\|", "", Var1)) %>%
  filter(str_detect(Var2, "park_data")) %>%
  mutate(Var2 = gsub("park_data\\.", "", Var2)) %>%
  # rename Park clusters (some symbols got removed)
  mutate(Var2=case_when(
    Var2=="CD4.T" ~ "CD4+T",
    Var2=="CD8.T" ~ "CD8+T",
    Var2=="CD8αα.I." ~ "CD8αα(I)",
    Var2=="CD8αα.II." ~ "CD8αα(II)",
    Var2=="DN.P." ~ "DN(P)",
    Var2=="DN.Q." ~ "DN(Q)",
    Var2=="DN.early." ~ "DN(early)",
    Var2=="DP.P." ~ "DP(P)",
    Var2=="DP.Q." ~ "DP(Q)",
    Var2=="T.agonist." ~ "T(agonist)",
    Var2=="Treg.diff." ~ "Treg(diff)",
    Var2=="αβT.entry." ~ "αβT(entry)",
    .default=Var2
  )) %>%
  # add nb of cells per cluster (from our dataset) to dataframe
  left_join(
    as.data.frame(table(seur_thymus$clusters_to_compare)),
    by="Var1"
    ) %>%
  # rename a few columns
  rename(
    var_x = Var1,
    Var1 = Var2,
    auroc = value,
    ncells_x = Freq
  ) %>%
  # add nb of cells per cluster (from park dataset) to dataframe
  left_join(
    as.data.frame(table(seur_park$clusters_to_compare)),
    by="Var1"
    ) %>%
  # rename a few columns again
  rename(
    var_y=Var1,
    ncells_y=Freq
    )

# sanity check
mtn.df
```


## Plot Supplementary Figure 4B
```{r plot-supp-fig-4b, fig.width=15, fig.height=15}
order_gapin <- paste0("c", 0:17)
order_park <- rev(c(
  "DN(early)",
  "DN(P)",
  "DN(Q)",
  "DP(P)",
  "DP(Q)",
  "αβT(entry)",
  "T(agonist)",
  "CD8αα(I)",
  "CD8αα(II)",
  "γδT",
  "Treg(diff)",
  "Treg",
  "CD4+T",
  "CD8+T",
  "Th17",
  "NKT"
))

bubble_plot(
  df=mtn.df,
  order_x = order_gapin,
  order_y = order_park,
  label_x = "",
  label_y = "Park clusters",
  auroc_min= 0.9
)
```




# SESSION INFO
```{r}
sessionInfo()
```