---
title: "Supplementary Fig 10 - Compare GEPs with literature gene signatures"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```




# IMPORT

## Import librairies
```{r import-librairies}
library(Seurat)
library(cowplot)
library(ggplot2)
library(ggh4x) # used for ordering facets in ggplot
library(tidyverse)
library(tidytext)
library(pheatmap)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```


## Import data
The list of genes characterizing each GEP was determined based on the gene ranks associated to each GEP (see Methods details & Supplementary Fig 17).
```{r import-data}
# genes characterizing each GEP
genes.gep <- read.csv("./data_github/cNMF/limited_nonimputed_genes_per_gep_post_rank_threshold_k12.csv", row.names=1)

# literature-derived gene signatures
genes.cano    <- readxl::read_excel("./data_github/literature_gene_signatures/canogamez_supptable.xlsx", skip=3, sheet=1)
genesCD8.rose <- readxl::read_excel("./data_github/literature_gene_signatures/rose_supp_clusmodulesCD8.xlsx", sheet=1)
genesCD4.rose <- readxl::read_excel("./data_github/literature_gene_signatures/rose_supp_clusmodulesCD4.xlsx", sheet=1)
genes.poon    <- readxl::read_excel("./data_github/literature_gene_signatures/poon_supptable.xlsx", sheet=6)[,-1]
genes.terekhova_general <- readxl::read_excel("./data_github/literature_gene_signatures/terekhova_supptable.xlsx", sheet=1)
genes.terekhova_cd4 <- readxl::read_excel("./data_github/literature_gene_signatures/terekhova_supptable.xlsx", sheet=3)
genes.terekhova_cd8 <- readxl::read_excel("./data_github/literature_gene_signatures/terekhova_supptable.xlsx", sheet=5)
genes.terekhova_gdt <- readxl::read_excel("./data_github/literature_gene_signatures/terekhova_supptable.xlsx", sheet=7)

# seurat object
seur_integrated <- readRDS("./data_geo/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
seur.pbmc <- subset(seur_integrated, tissue=="PBMC")
DimPlot(seur.pbmc, reduction="umap_integrated", group.by="clusters_integrated_data", cols=cols_integrated)
```




# PREPARE GENE LISTS

We will store all gene lists dataframes (genes for each GEP, genes for each literature-derived gene signature) in a list, which we will eventually combine into a single long-format dataframe.
```{r set-genesig-list}
genesig_longdf_list <- list()
```


## GEPs
```{r prep-geps}
head(genes.gep)
geps_df <- gather(genes.gep, key=geneprogram, value=gene, colnames(genes.gep)) %>%
  filter(!is.na(gene)) %>%
  mutate(geneprogram=sub("_", "", geneprogram))
head(geps_df)

# check nb of genes per GEP
table(geps_df$geneprogram, useNA="ifany")

# add df to genesig_longdf_list
genesig_longdf_list[["GEPs"]] <- geps_df

# sanity check
# table(is.na(genesig_longdf_list$GEPs$gene)) # all FALSE
```


## Cano-Gamez _et al._ gene signatures
```{r prep-canogamez}
head(genes.cano)

# take a look at the different clusters/gene programs available
table(genes.cano$cluster, useNA="ifany")

# get a long format df with gene signatures
cano_df <- genes.cano %>%
  mutate(
    cluster=gsub(" ", "", cluster),
    cluster=paste0("CanoGamez | CD4_", cluster)
  ) %>%
  # keep only significant genes & with minimum log2FC
  filter(p_val_adj < 0.05 & LFC > 0.25) %>% # lowest log2FC is 0.25
  # keep only columns of interest
  select(cluster, gene) %>%
  rename(geneprogram=cluster) %>%
  # add general categories
  mutate(geneprogram_cat=case_when(
    geneprogram=="CanoGamez | CD4_Tnaive" ~"Tnaive",
    geneprogram=="CanoGamez | CD4_TCM"    ~"Tcm",
    geneprogram=="CanoGamez | CD4_TEM"    ~"Tem",
    geneprogram=="CanoGamez | CD4_TEMRA"  ~"Temra",
    geneprogram=="CanoGamez | CD4_nTreg"  ~"Treg"
  ))
head(cano_df)

# add df to genesig_longdf_list
genesig_longdf_list[["canogamez"]] <- cano_df

# table(is.na(genesig_longdf_list$canogamez$gene)) # sanity check
```


## Rose _et al._ gene signatures
Let's first do a little sanity check that we are able to reproduce the heatmaps in [Figures 2e and 2h](https://www.nature.com/articles/s42003-023-04747-9/figures/2) of the Rose _et al._ paper.
```{r rose-reproduce-htmps, fig.height=6, fig.width=4}
htmp <- function(rosedf, lineage){
  # Get rpkm
  if(lineage=="CD4"){ counts <- rosedf[,9:20] }
  else if(lineage=="CD8"){ counts <- rosedf[,9:24] }
  counts <- as.data.frame(counts)
  rownames(counts) <- rosedf$ENTREZID
  # Get metadata
  meta <- rosedf[,c(1,3)] %>% column_to_rownames("ENTREZID")
  cols <- RColorBrewer::brewer.pal(5, "Greys")
  names(cols) <- unique(meta$k.5.cluster)
  # Plot
  pheatmap(counts,
           scale="row",
           cluster_rows=F,
           cluster_cols=F,
           color=rev(colorRampPalette(RColorBrewer::brewer.pal(10, "RdBu"))(100)),
           annotation_row=meta,
           annotation_colors=list(k.5.cluster=cols),
           show_rownames=F)
}

htmp(genesCD4.rose, lineage="CD4")
htmp(genesCD8.rose, lineage="CD8")
```

These plots indeed resemble Figures 2e & h from the original paper. Let's obtain the gene signatures then.
```{r prep-rose}
head(genesCD4.rose)
head(genesCD8.rose)

roseCD4 <- genesCD4.rose %>%
  select(SYMBOL, k.5.cluster) %>%
  filter(!is.na(SYMBOL)) %>% # the few NA symbols are miRNA or uncharacterized loci
  rename(gene=SYMBOL, geneprogram=k.5.cluster) %>%
  mutate(
    geneprogram=ifelse(geneprogram==1, "Rose | CD4_modul1_Tem/cm",
                       ifelse(geneprogram==2, "Rose | CD4_modul2_Tcm/em",
                              ifelse(geneprogram==3, "Rose | CD4_modul3_Tem/cm",
                                     ifelse(geneprogram==4, "Rose | CD4_modul4_Tem",
                                            ifelse(geneprogram==5, "Rose | CD4_modul5_Tnaive", "?")))))
  )

roseCD8 <- genesCD8.rose %>%
  select(SYMBOL, k.5.cluster) %>%
  filter(!is.na(SYMBOL)) %>% # the few NA symbols are miRNA or uncharacterized loci
  rename(gene=SYMBOL, geneprogram=k.5.cluster) %>%
  mutate(
    geneprogram=ifelse(geneprogram==1, "Rose | CD8_modul1_Temra",
                       ifelse(geneprogram==2, "Rose | CD8_modul2_Tcm/em",
                              ifelse(geneprogram==3, "Rose | CD8_modul3_Tem/emra",
                                     ifelse(geneprogram==4, "Rose | CD8_modul4_Tnaive",
                                            ifelse(geneprogram==5, "Rose | CD8_modul5_Tnaive/cm", "?")))))
  )

rose_df <- rbind(roseCD4, roseCD8) %>%
  mutate(geneprogram_cat=case_when(
    geneprogram=="Rose | CD4_modul1_Tem/cm"    ~ "Tem",
    geneprogram=="Rose | CD4_modul2_Tcm/em"    ~ "Tcm",
    geneprogram=="Rose | CD4_modul3_Tem/cm"    ~ "Tem",
    geneprogram=="Rose | CD4_modul4_Tem"       ~ "Tem",
    geneprogram=="Rose | CD4_modul5_Tnaive"    ~ "Tnaive",
    geneprogram=="Rose | CD8_modul1_Temra"     ~ "Temra",
    geneprogram=="Rose | CD8_modul2_Tcm/em"    ~ "Tcm",
    geneprogram=="Rose | CD8_modul3_Tem/emra"  ~ "Tem",
    geneprogram=="Rose | CD8_modul4_Tnaive"    ~ "Tnaive",
    geneprogram=="Rose | CD8_modul5_Tnaive/cm" ~ "Tnaive"
  ))
head(rose_df)

# add df to genesig_longdf_list
genesig_longdf_list[["rose"]] <- rose_df

# table(is.na(genesig_longdf_list$rose$gene)) # sanity check
```


## Poon _et al._ gene signatures
```{r prep-poon}
head(genes.poon)
colnames(genes.poon) <- gsub(" ", "", colnames(genes.poon))

# Transform to long format gene program by gene program (each gene program has 3 columns: genes symbols, padj, logFC)
poon_df <- data.frame()
for (i in seq(1,ncol(genes.poon), 3)){
  subdf <- genes.poon[,i:(i+2)]
  geneprogram <- sub("_.*", "",colnames(subdf)[1])
  colnames(subdf) <- c("gene", "padj", "logFC")
  subdf <- subdf %>%
    filter(padj<0.001 & logFC>0.25)
  cat(geneprogram, "\n")
  cat("# genes:", nrow(subdf), "\n")
  cat("min log2FC:", min(subdf$logFC), "\n\n")
  subdf$geneprogram <- paste0("Poon | ", geneprogram) # add column with the name of the geneprogram
  poon_df <- rbind(poon_df, subdf)
}
head(poon_df)
table(poon_df$geneprogram, useNA="ifany")

poon_df <- poon_df %>%
  select(geneprogram, gene) %>%
  mutate(geneprogram_cat=case_when(
    geneprogram=="Poon | CyclingTRM"  ~"Trm",
    geneprogram=="Poon | CD4Naive"    ~"Tnaive",
    geneprogram=="Poon | CD4TCM/TFH"  ~"Tcm",
    geneprogram=="Poon | CD4TRM"      ~"Trm",
    geneprogram=="Poon | CD4Treg"     ~"Treg",
    geneprogram=="Poon | CD8MAIT"     ~"MAIT",
    geneprogram=="Poon | CD8Naive"    ~"Tnaive",
    geneprogram=="Poon | CD8TEM/TEMRA"~"Temra",
    geneprogram=="Poon | CD8TRM"      ~"Trm"
  ))

# add df to genesig_longdf_list
genesig_longdf_list[["poon"]] <- poon_df

# table(is.na(genesig_longdf_list$poon$gene)) # sanity check
```

## Terekhova _et al._ gene signatures
In this dataset, the annotation "Tnaive/cm/em/emra" was evaluated based on the heatmaps provided in the paper (looking at CCR7, CD62L, CD45RA, CD45RO).
```{r prep-terekhova}
head(genes.terekhova_general)
# head(genes.terekhova_cd4)
# head(genes.terekhova_cd8)
# head(genes.terekhova_gdt)

terekhova_df <- cbind(
  genes.terekhova_general[,"MAITcells"],
  genes.terekhova_cd4,
  genes.terekhova_cd8,
  genes.terekhova_gdt
  ) %>%
  pivot_longer(cols=everything(), names_to="geneprogram", values_to = "gene") %>%
  mutate(geneprogram=paste0("Terekhova | ", geneprogram),
         geneprogram_cat=case_when(
           geneprogram=="Terekhova | MAITcells"                ~ "MAIT",
           geneprogram=="Terekhova | CD4_Tnaive"               ~ "Tnaive",
           geneprogram=="Terekhova | CD4_Tnaive_IFN"           ~ "Tnaive",
           geneprogram=="Terekhova | CD4_Tfh"                  ~ "Tcm",
           geneprogram=="Terekhova | CD4_Th1"                  ~ "Tcm",
           geneprogram=="Terekhova | CD4_Th1/Th17"             ~ "Tem",
           geneprogram=="Terekhova | CD4_Th17"                 ~ "Tem",
           geneprogram=="Terekhova | CD4_Th22"                 ~ "Tem",
           geneprogram=="Terekhova | CD4_Th2"                  ~ "Tcm",
           geneprogram=="Terekhova | CD4_HLADRpos_memory"      ~ "Tem",
           geneprogram=="Terekhova | CD4_Exhausted-like_memory"~ "Tcm",
           geneprogram=="Terekhova | CD4_Terminal_effector"    ~ "Temra",
           geneprogram=="Terekhova | CD4_Temra"                ~ "Temra",
           geneprogram=="Terekhova | CD4_Treg_cytotoxic"       ~ "Treg",
           geneprogram=="Terekhova | CD4_Treg_naive"           ~ "Treg",
           geneprogram=="Terekhova | CD4_Treg_memory"          ~ "Treg",
           geneprogram=="Terekhova | CD4_Treg_KLRB1+_RORC+"    ~ "Treg",
           geneprogram=="Terekhova | CD8_Tnaive"        ~ "Tnaive",
           geneprogram=="Terekhova | CD8_Tnaive_IFN"    ~ "Tnaive",
           geneprogram=="Terekhova | CD8_Tcm_CCR4pos"   ~ "Tcm",
           geneprogram=="Terekhova | CD8_Tcm_CCR4neg"   ~ "Tcm",
           geneprogram=="Terekhova | CD8_Trm"           ~ "Trm",
           geneprogram=="Terekhova | CD8_Tmem_KLRC2pos" ~ "other",
           geneprogram=="Terekhova | CD8_Tem_GZMKpos"   ~ "Tem",
           geneprogram=="Terekhova | CD8_HLADRpos"      ~ "Tem",
           geneprogram=="Terekhova | CD8_proliferative" ~ "Tprolif",
           geneprogram=="Terekhova | CD8_Tem_GZMBpos"   ~ "Temra",
           geneprogram=="Terekhova | CD8_Temra"         ~ "Temra",
           geneprogram=="Terekhova | CD8_NKTlike"       ~ "Temra",
           geneprogram=="Terekhova | GD_naive"     ~ "Tnaive",
           geneprogram=="Terekhova | GD_Vd2_GZMK+" ~ "Tem",
           geneprogram=="Terekhova | GD_Vd2_GZMB+" ~ "Temra",
           geneprogram=="Terekhova | GD_Vd1_GZMK+" ~ "Tem",
           geneprogram=="Terekhova | GD_Vd1_GZMB+" ~ "Temra"
         ))

# add df to genesig_longdf_list
genesig_longdf_list[["terekhova"]] <- terekhova_df

# table(is.na(genesig_longdf_list$terekhova$gene)) # sanity check
```


## Combine gene lists
```{r prep-gene-lists-into-one-df}
longdf <- bind_rows(genesig_longdf_list, .id="dataset")
table(longdf$dataset, useNA="ifany")
# canogamez      GEPs      poon      rose   terekhova
#   387         4,068    26,463     1,672       3,400

# remove any genes that we don't have in seur.pbmc (i.e. we didn't detect in our dataset)
longdf <- longdf %>%
  filter(gene %in% rownames(seur.pbmc)) %>%
  filter(geneprogram != "GEP12") # remove batch-driven GEP
table(longdf$dataset, useNA="ifany")
# canogamez      GEPs      poon      rose  terekhova
#   341         3,787    19,959     1,542      3,145

# check out the number of genes per signature/program
table(longdf$geneprogram)

# show the longdf
print(longdf)
```

For later use, define colors for each category of gene programs (Tnaive, cm, etc.).
```{r prep-cols-vector}
# Create a colors vector & dataframe
cols_df <- longdf %>%
  select(geneprogram, geneprogram_cat) %>%
  distinct() %>%
  filter(!geneprogram %in% grep("GEP", unique(longdf$geneprogram), value=T)) %>%
  mutate(color= case_when(
    geneprogram_cat=="Tnaive"   ~ "#b3e2cd",
    geneprogram_cat=="Tcm"      ~ "#f4cae4",
    geneprogram_cat=="Tem"      ~ "#cbd5e8",
    geneprogram_cat=="Temra"    ~ "#fdcdac",
    geneprogram_cat=="Treg"     ~ "#fbb4ae",
    geneprogram_cat=="Trm"      ~ "#4292c6",
    geneprogram_cat=="MAIT" ~ "#8c6bb1",
    geneprogram_cat=="Tprolif"  ~ "#e5d8bd",
    geneprogram_cat=="other"     ~ "#f2f2f2",
  ))
cols_geneprogcat <- unique(cols_df$color)
names(cols_geneprogcat) <- unique(cols_df$geneprogram_cat)
```




# COMPUTE OVERLAP BETWEEN GENE LISTS (Sup Fig 10A)

## Define Jaccard function
```{r function-jaccard}
overlapcoef <- function(a, b, coef="jaccardweight") {
  # Jaccard Index
  if(coef=="jaccard"){
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
  }
  # Weighted Jaccard Index (to take into account differences in lengths between A and B)
  else if(coef=="jaccardweight"){
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    maxj = min(length(a),length(b))/max(length(a), length(b))
    return (intersection/(union*maxj))
  }
}
```


## Define null hypothesis function
```{r function-null-hypothesis}
NullOverlap_double <- function(
    seuratobj=seur.pbmc,
    df_geneprograms=longdf,
    geneprograms_to_test=unique(longdf$geneprogram),
    coefficient_to_compute="jaccardweight",
    nbins=25,
    nrandom=1000
){
  
  # ___________________________________________________________
  # -- 1. Get DF with all genes & binned by expression level --
  cat("\n-- 1. GET ALL GENES & BIN THEM BY EXPRESSION LEVEL --\n")
  allgenes_binned_DF <- data.frame("gene"=rownames(seuratobj),
                                   "totalexpression"=rowSums(seuratobj@assays$RNA@data))
  # nrow(allgenes_binned_DF) == nrow(seuratobj) # sanity check
  # table(rownames(allgenes_binned_DF)==allgenes_binned_DF$gene) # sanity check
  # hist(allgenes_binned_DF$totalexpression, breaks=100) # visualize distribution
  
  # Get the bins by percentiles
  allgenes_binned_DF <- allgenes_binned_DF[order(allgenes_binned_DF$totalexpression),]
  allgenes_binned_DF$bin <- ggplot2::cut_number(x = allgenes_binned_DF$totalexpression, n = nbins, labels = FALSE, right = FALSE)
  cat("Number of genes in each bin, from bin 1 to bin", nbins, ":\n")
  print(table(allgenes_binned_DF$bin))
  # ggplot(allgenes_binned_DF)+
  #   geom_density(aes(x=totalexpression, group=bin, color=factor(bin, levels=1:10))) # sanity check
  
  # _____________________________________________________
  # -- 2. Draw random ctrl genes for each gene program --
  cat("\n-- 2. DRAW RANDOM CTRL GENES FOR EACH GENE SET --\n")
  randomgenesets_list <- list()
  
  for(refgeneprog in geneprograms_to_test){
    cat("\n++ Drawing random ctrl genes for", refgeneprog, "\n")
    # Find in which expression bins are our genes from the gene set
    refgeneset <- df_geneprograms[df_geneprograms$geneprogram==refgeneprog, "gene"]
    refgeneset_binned_DF <- allgenes_binned_DF[allgenes_binned_DF$gene %in% refgeneset,]
    # table(nrow(refgeneset_binned_DF)==length(refgeneset))
    
    # ++++++++++++++++++++++++++++++++++
    # Do 1000 random draws of gene lists
    ctrlgenematrix <- matrix(nrow=length(refgeneset), ncol=nrandom)
    for(i in 1:nrandom){
      if(i%%100==0){cat(paste0("\nrandom draw #", i))} # progress bar
      # Loop by expression bin & sample equal number of genes from each bin
      ctrlgeneset <- c()
      for(bin in unique(refgeneset_binned_DF$bin)){
        # print(bin)
        ngenes_to_sample <- sum(refgeneset_binned_DF$bin==bin)
        # print(ngenes_to_sample)
        ctrlgenes_in_same_bin <- sample(x=allgenes_binned_DF[allgenes_binned_DF$bin==bin,"gene"], size=ngenes_to_sample, replace=FALSE)
        # print(length(ctrlgenes_in_same_bin))
        ctrlgeneset <- c(ctrlgeneset, ctrlgenes_in_same_bin)
      }
      # Sanity check
      if(length(ctrlgeneset)!=length(refgeneset)){cat("\nPROBLEM: control gene set of size", length(ctrlgeneset), "while gene program length is", length(refgeneset))}
      
      # Add ctrlgeneset to matrix
      ctrlgenematrix[,i] <- ctrlgeneset
    }
    # Sanity check
    cat("\n\n++ Selection of random ctrl genes over! ++\n")
    cat("Length of reference gene set is", length(unique(refgeneset)), "and length of the random gene sets:", nrow(ctrlgenematrix), "\n")
    # ++++++++++++++++++++++++++++++++++
    
    # Add to the list
    randomgenesets_list[[refgeneprog]] <- ctrlgenematrix
  }
  
  # _____________________________________________________________________________
  # -- 3. Compute random & observed overlap between each pairwise gene program --
  cat("\n-- 3. COMPUTE RANDOM AND OBSERVED OVERLAP --\n")
  
  combinations <- t(combn(geneprograms_to_test,2))
  colnames(combinations) <- c("geneprogram1", "geneprogram2")
  
  # add self combinations
  combinations <- rbind(combinations,
                        t(matrix(rep(geneprograms_to_test, 2), nrow=2, byrow=T)))
  combinations <- as.data.frame(combinations)
  ncombinations <- nrow(combinations)
  
  for(i in 1:nrow(combinations)){
    # cat("\n", combinations[i,1], "vs", combinations[i,2])
    if(i%%25==0){cat(paste0("\ncomparison ", i, "/", ncombinations))} # progress bar
    
    # Get observed overlap
    genesetA <- df_geneprograms %>% filter(geneprogram==combinations[i,"geneprogram1"]) %>% pull(gene)
    genesetB <- df_geneprograms %>% filter(geneprogram==combinations[i,"geneprogram2"]) %>% pull(gene)
    observedpercent <- overlapcoef(genesetA, genesetB, coef=coefficient_to_compute)
    
    # Get vector of random overlaps
    randomgenesetA <- randomgenesets_list[[combinations[i,"geneprogram1"]]]
    randomgenesetB <- randomgenesets_list[[combinations[i,"geneprogram2"]]]
    ctrlpercent <- sapply(1:nrandom, function(x) overlapcoef(randomgenesetA[,x], randomgenesetB[,x], coef=coefficient_to_compute))
    if(length(ctrlpercent) != nrandom){cat("\nPROBLEM: vector of random overlaps of length", length(ctrlpercent), "when there should be", nrandom, "random sets")}
    
    # save info
    combinations[i,"observedoverlap"] <- observedpercent
    combinations[i,"randomoverlap_mean"] <- mean(ctrlpercent)
    combinations[i,"randomoverlap_min"] <- min(ctrlpercent)
    combinations[i,"randomoverlap_max"] <- max(ctrlpercent)
    combinations[i,"pval"] <- sum(ctrlpercent>observedpercent)/nrandom
  }
  
  return(combinations)
}
```


## Compute wJI between gene signatures
```{r compute-JI-btw-gene-signatures}
genesets_overlap_wjaccard <- NullOverlap_double(
  seuratobj=seur.pbmc,
  df_geneprograms=longdf,
  geneprograms_to_test=unique(longdf$geneprogram),
  coefficient_to_compute="jaccardweight",
  nbins=25, nrandom=1000
  )
head(genesets_overlap_wjaccard)
```

Let's have a first look at the top 10 literature-derived gene signatures with highest weighted JI for each GEP.
```{r plot-wji-top10, fig.width=15, fig.height=8}
genesets_overlap_wjaccard %>%
  # keep geps of interest
  filter(geneprogram1 %in% c("GEP3", "GEP4", "GEP5", "GEP6")) %>%
  filter(!geneprogram2 %in% grep("GEP", geneprogram2, value=T)) %>%
  # keep only top 10 wJI and add padj
  group_by(geneprogram1) %>%
  mutate(padj=pval*n_distinct(geneprogram2)) %>%
  top_n(10, observedoverlap) %>%
  ungroup() %>%
  distinct() %>%
  # last details (color, padj to plot)
  left_join(cols_df, by=join_by("geneprogram2"=="geneprogram")) %>%
  mutate(padj_toplot=ifelse(padj==0, "< 0.001",
                            ifelse(padj<0.05, as.character(round(padj, 2)), ""))) %>%
  # PLOT
  ggplot(aes(x=reorder_within(geneprogram2, -observedoverlap, geneprogram1), y=observedoverlap))+
  geom_bar(stat="identity", aes(fill=geneprogram_cat))+
  facet_wrap(~factor(geneprogram1, levels=paste0("GEP", 1:11)), nrow=1, scales="free_x")+
  geom_point(aes(y=randomoverlap_mean), shape="_", size=3)+
  tidytext::scale_x_reordered() +
  geom_text(aes(label=padj_toplot), size=4, angle=90, hjust=-0.2)+
  ylim(c(0,1))+
  scale_fill_manual(values=cols_geneprogcat, name="")+
  theme_cowplot()+
  theme(
    axis.text.x=element_text(angle=45, hjust=1, size=12),
    panel.grid.major.y=element_line(colour="lightgrey", linetype=2)#,
    # plot.margin=margin(10,10,10,70)
  )+
  labs(x="", y="Weighted Jaccard index")
```


## Plot wJI (supp fig 10A)
```{r plot-supp-fig10A, fig.height=20, fig.width=20}
# define order of literature-derived gene signatures (i.e. order of facets)
order_programs <- c(
  cols_df %>% filter(geneprogram_cat=="Tnaive") %>% arrange(geneprogram) %>% pull(geneprogram),
  cols_df %>% filter(geneprogram_cat=="Tcm") %>% arrange(geneprogram) %>% pull(geneprogram),
  cols_df %>% filter(geneprogram_cat %in% c("MAIT")) %>% arrange(geneprogram) %>% pull(geneprogram),
  cols_df %>% filter(geneprogram_cat %in% c("Tem")) %>% arrange(geneprogram) %>% pull(geneprogram),
  cols_df %>% filter(geneprogram_cat=="Temra") %>% arrange(geneprogram) %>% pull(geneprogram)
)


genesets_overlap_wjaccard %>%
  # keep geps of interest
  filter(geneprogram1 %in% c("GEP3", "GEP4", "GEP5", "GEP6")) %>%
  filter(!geneprogram2 %in% c(grep("GEP", geneprogram2, value=T))) %>%
  # add padj
  group_by(geneprogram1) %>%
  mutate(padj=pval*n_distinct(geneprogram2)) %>%
  ungroup() %>%
  distinct() %>%
  # last details (color, padj to plot)
  left_join(cols_df, by=join_by("geneprogram2"=="geneprogram")) %>%
  mutate(padj_toplot=ifelse(padj==0, "< 0.001",
                            ifelse(padj<0.05, as.character(round(padj, 2)), ""))) %>%
  # remove some geneprograms that we're not interested in (they don't show very high wJI with GEP3-6)
  filter(!geneprogram_cat %in% c("Treg", "Trm", "Tprolif", "other")) %>%
  filter(!geneprogram2 %in% c(
    grep("Tnaive_IFN", geneprogram2, value=T),
    "Terekhova | CD4_Exhausted-like_memory",
    grep("Terekhova \\| GD_", geneprogram2, value=T)
    )) %>%
  # PLOT SUPP FIG 10A
  ggplot(aes(x=factor(geneprogram1, levels=rev(paste0("GEP", 1:11))), y=observedoverlap))+
    geom_bar(stat="identity", aes(fill=geneprogram_cat))+
    geom_point(aes(y=randomoverlap_mean), shape="|", size=3)+
    tidytext::scale_x_reordered() +
    geom_text(aes(label=padj_toplot), size=4, angle=0, hjust=-0.2)+
    facet_manual(~factor(geneprogram2, levels=order_programs),
                 design=rbind(
                   c(1:6), c(7:8,NA,NA,NA,NA), # naive
                   c(9:14), c(15:17,NA,NA,NA), # Tcm
                   c(18:19,NA,NA,NA,NA), # MAIT
                   c(20:25), c(26:30, NA), # Tem
                   c(31:36), c(37:38,NA,NA,NA,NA) # Temra
                   ),
                 scales="free_x")+
    ylim(c(0,1))+
    coord_flip()+
    theme_cowplot()+
    scale_fill_manual(values=cols_geneprogcat, name="")+
    labs(y="Weighted Jaccard Index", x="", title="Supplementary Figure 10A")
```




# COEXPRESSION GENE LISTS (Sup Fig 10B)

## Compute gene signature scores
We will first put all gene signatures into a list.
```{r longdf-into-list}
# a few sanity checks
table(is.na(longdf$geneprogram)) # should all be FALSE
table(is.na(longdf$gene)) # should all be FALSE

# get gene lists into a list (lol)
geneprograms.list <- list()
for(gp in unique(longdf$geneprogram)){
  genes_in_program <- longdf %>%
    filter(geneprogram==gp & gene %in% rownames(seur_integrated)) %>%
    pull(gene)
  print(paste0(gp, ": ", length(genes_in_program), " genes"))
  geneprograms.list[[gp]] <- genes_in_program
}
# print(lengths(geneprograms.list))

# rename some gene signatures
names(geneprograms.list) <- gsub(" \\| ", "_", names(geneprograms.list))
names(geneprograms.list) <- gsub("\\/", "_", names(geneprograms.list))
names(geneprograms.list) <- gsub("\\+", "", names(geneprograms.list))
names(geneprograms.list) <- paste0("score_", names(geneprograms.list))

# sanity check
names(geneprograms.list)
```

Now let's compute the gene signature scores.
```{r compute-genesig-scores, fig.height=20, fig.width=10}
# compute score
seur_integrated <- AddModuleScore(seur_integrated, name = names(geneprograms.list), features=geneprograms.list, seed=1)

# Remove the annoying numbers that are being added
colnames(seur_integrated@meta.data)[32:100] <- names(geneprograms.list)

# Sanity check GEPs (should look similar to Fig 1K, even though it's not GEP usage)
SCpubr::do_FeaturePlot(
  seur_integrated,
  features=colnames(seur_integrated@meta.data)[32:42],
  reduction="umap_integrated",
  viridis.palette = "B",
  use_viridis = T,
  order=T,
  ncol=3
  )
```


## Define functions to plot coexpression
These functions were taken from Seurat [`visualization.R`](https://github.com/satijalab/seurat/blob/HEAD/R/visualization.R) script. These functions are used by `Seurat::FeaturePlot(..., blend=T)`. We used them and wrote our own in-house plotting function `PlotCoexpression()`.
```{r functions-coexpression}
# SET COLOR GRADIENT
BlendMatrix <- function(
    n = 10,
    col.threshold = 0.5,
    two.colors = c("#ff0000", "#00ff00"),
    negative.color = "black"
) {
  if (0 > col.threshold || col.threshold > 1) {
    stop("col.threshold must be between 0 and 1")
  }
  C0 <- as.vector(col2rgb(negative.color, alpha = TRUE))
  C1 <- as.vector(col2rgb(two.colors[1], alpha = TRUE))
  C2 <- as.vector(col2rgb(two.colors[2], alpha = TRUE))
  blend_alpha <- (C1[4] + C2[4])/2
  C0 <- C0[-4]
  C1 <- C1[-4]
  C2 <- C2[-4]
  merge.weight <- min(255 / (C1 + C2 +  C0 + 0.01))
  sigmoid <- function(x) {
    return(1 / (1 + exp(-x)))
  }
  blend_color <- function(
    i,
    j,
    col.threshold,
    n,
    C0,
    C1,
    C2,
    alpha,
    merge.weight
  ) {
    c.min <- sigmoid(5 * (1 / n - col.threshold))
    c.max <- sigmoid(5 * (1 - col.threshold))
    c1_weight <- sigmoid(5 * (i / n - col.threshold))
    c2_weight <- sigmoid(5 * (j / n - col.threshold))
    c0_weight <-  sigmoid(5 * ((i + j) / (2 * n) - col.threshold))
    c1_weight <- (c1_weight - c.min) / (c.max - c.min)
    c2_weight <- (c2_weight - c.min) / (c.max - c.min)
    c0_weight <- (c0_weight - c.min) / (c.max - c.min)
    C1_length <- sqrt(sum((C1 - C0) ** 2))
    C2_length <- sqrt(sum((C2 - C0) ** 2))
    C1_unit <- (C1 - C0) / C1_length
    C2_unit <- (C2 - C0) / C2_length
    C1_weight <- C1_unit * c1_weight
    C2_weight <- C2_unit * c2_weight
    C_blend <- C1_weight * (i - 1) * C1_length / (n - 1) + C2_weight * (j - 1) * C2_length / (n - 1) + (i - 1) * (j - 1) * c0_weight * C0 / (n - 1) ** 2 + C0
    C_blend[C_blend > 255] <- 255
    C_blend[C_blend < 0] <- 0
    return(rgb(
      red = C_blend[1],
      green = C_blend[2],
      blue = C_blend[3],
      alpha = alpha,
      maxColorValue = 255
    ))
  }
  blend_matrix <- matrix(nrow = n, ncol = n)
  for (i in 1:n) {
    for (j in 1:n) {
      blend_matrix[i, j] <- blend_color(
        i = i,
        j = j,
        col.threshold = col.threshold,
        n = n,
        C0 = C0,
        C1 = C1,
        C2 = C2,
        alpha = blend_alpha,
        merge.weight = merge.weight
      )
    }
  }
  return(blend_matrix)
}


# SET COLOR MATRIX
Melt <- function(x) {
  if (!is.data.frame(x = x)) {
    x <- as.data.frame(x = x)
  }
  return(data.frame(
    rows = rep.int(x = rownames(x = x), times = ncol(x = x)),
    cols = unlist(x = lapply(X = colnames(x = x), FUN = rep.int, times = nrow(x = x))),
    vals = unlist(x = x, use.names = FALSE)
  ))
}


# PLOT COLOR MATRIX
BlendMap <- function(color.matrix, step=2, xtext='rows', ytext="cols") {
  color.heat <- matrix(
    data = 1:prod(dim(x = color.matrix)) - 1,
    nrow = nrow(x = color.matrix),
    ncol = ncol(x = color.matrix),
    dimnames = list(
      1:nrow(x = color.matrix),
      1:ncol(x = color.matrix)
    )
  )
  
  # xbreaks <- seq.int(from = 0, to = nrow(x = color.matrix), by = step)
  # ybreaks <- seq.int(from = 0, to = ncol(x = color.matrix), by = step)
  color.heat <- Melt(x = color.heat)
  color.heat$rows <- as.numeric(x = as.character(x = color.heat$rows))
  color.heat$cols <- as.numeric(x = as.character(x = color.heat$cols))
  color.heat$vals <- factor(x = color.heat$vals)
  plot <- ggplot(
    data = color.heat,
    mapping = aes_string(x = "rows", y = "cols", fill = 'vals')
  ) +
    geom_raster(show.legend = FALSE) +
    theme(plot.margin = unit(x = rep.int(x = 0, times = 4), units = 'cm')) +
    # scale_x_continuous(breaks = xbreaks, expand = c(0, 0), labels = xbreaks) +
    # scale_y_continuous(breaks = ybreaks, expand = c(0, 0), labels = ybreaks) +
    scale_fill_manual(values = as.vector(x = color.matrix)) +
    labs(x=xtext, y=ytext)+
    theme_cowplot()
  
  if(step!=0){
    xbreaks <- seq.int(from = 0, to = nrow(x = color.matrix), by = step)
    ybreaks <- seq.int(from = 0, to = ncol(x = color.matrix), by = step)
    plot <- plot +
      scale_x_continuous(breaks = xbreaks, expand = c(0, 0), labels = xbreaks) +
      scale_y_continuous(breaks = ybreaks, expand = c(0, 0), labels = ybreaks)
  } else if(step==0){
    plot <- plot+theme(axis.text=element_blank(), axis.ticks=element_blank())
  }
  
  return(plot)
}


# Normalize expression level of 2 features
BlendExpression <- function(data, nlevels=100) {
  if (ncol(x = data) != 2) {
    stop("'BlendExpression' only blends two features")
  }
  features <- colnames(x = data)
  data <- as.data.frame(x = apply(
    X = data,
    MARGIN = 2,
    FUN = function(x) {
      return(round(x = (nlevels-1) * (x - min(x)) / (max(x) - min(x))))
    }
  ))
  data[, 3] <- data[, 1] + data[, 2] * nlevels
  # colnames(x = data) <- c(features, paste(features, collapse = '_'))
  colnames(x = data) <- c(features, "blend")
  for (i in 1:ncol(x = data)) {
    data[, i] <- factor(x = data[, i])
  }
  return(data)
}


# PLOT UMAP WITH COEXPRESSION
PlotCoexpression <- function(
    seuratobj=seur_integrated,
    features,
    plotting="blend",
    pwithmatrix=T,
    rasterdpi=300,
    nlevels=100,
    cols.neg="#969696",
    cols.pos=c("#74c476", "#fd8d3c"),
    col.threshold=0.5,
    colmatrix_stepsize=0,
    order=T
){
  # GET COLOR MATRIX
  # cat("\n-> Getting color matrix\n")
  color.matrix <- BlendMatrix(
    two.colors = cols.pos,
    col.threshold = col.threshold,
    negative.color = cols.neg,
    n=nlevels
  )
  
  # DEFINE COLOR LIST FOR PLOTTING
  # cat("\n-> Defining colors for plotting\n")
  colors <- list(
    color.matrix[, 1], # red
    color.matrix[1, ], # green
    as.vector(x = color.matrix)
  )
  
  # BLEND EXPRESSION
  # cat("\n-> Blending features expression\n")
  df <- seuratobj@meta.data[, features]
  df <- BlendExpression(df, nlevels=nlevels) # 3 columns
  # head(df)
  # GET PLOTTING DATAFRAME
  # cat("\n-> Defining plotting DF\n")
  dims <- seuratobj@reductions$umap_integrated@cell.embeddings
  # head(dims)
  df_final <- cbind(df, dims)
  # head(df_final)
  
  # PLOT
  if(plotting=="feature1"){
    # cat("\n-> Plotting feature 1\n")
    if(order==T){df_final <- df_final[order(df_final[,1]),]}
    df_final[,1] <- as.numeric(as.character(df_final[,1])) # transform factors to numbers for plotting
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[1]))+
      geom_point(size=0.1)+
      scale_color_gradient(low=color.matrix[1, 1], high=color.matrix[nrow(color.matrix), 1])
  }
  else if(plotting=="feature2"){
    # cat("\n-> Plotting feature 2\n")
    if(order==T){df_final <- df_final[order(df_final[,2]),]}
    df_final[,2] <- as.numeric(as.character(df_final[,2])) # transform factors to numbers for plotting
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[2]))+
      geom_point(size=0.1)+
      scale_color_gradient(low=color.matrix[1, 1], high=color.matrix[1, ncol(color.matrix)])
  }
  else if(plotting=="blend"){
    # cat("\n-> Plotting blended features\n")
    if(order==T){df_final <- df_final[order(df_final[,1], df_final[,2]),]} # order points by increasing value of score 1 and score 2
    # Colors
    cols.use <- as.vector(color.matrix)
    names(cols.use) <- as.character(0:(length(cols.use)-1))
    # Plot
    p <- ggplot(df_final, aes_string(x=colnames(dims)[1], y=colnames(dims)[2], color=colnames(df_final)[3]))+
      geom_point(size=0.3)+
      scale_color_manual(values=cols.use)+
      labs(x="UMAP1", y="UMAP2")+
      theme_cowplot()+
      theme(legend.position="none",
            axis.text=element_blank(),
            axis.title=element_blank(),
            axis.ticks = element_blank(),
            axis.line = element_blank(),
            panel.border=element_rect(color="white", fill=NA, size=1))
    # rasterise to avoid computer crashing
    p <- ggrastr::rasterise(p, layers="Point", dpi=rasterdpi)
    # add color matrix if specified
    if(pwithmatrix==T){
      # cat("\n-> Adding color matrix on plot\n")
      p <- ggdraw(p)+
        draw_plot(BlendMap(color.matrix, step=colmatrix_stepsize, xtext=features[1], ytext=features[2]),
                  0.05,0.06,.25,.25)
    }
  }
  
  return(p)
}
```


## Plot coexpression on integrated umap (Sup Fig 10B)
First plot all UMAPs and save them in variables.
```{r plot-coexpression}
# GEP4
p11 <- PlotCoexpression(features=c("score_GEP4", "score_CanoGamez_CD4_TCM")) +
  draw_label("CD4 Tcm", hjust=0.5, y=0.98, size=30)
p12 <- PlotCoexpression(features=c("score_GEP4", "score_Rose_CD8_modul2_Tcm_em")) +
  draw_label("CD8 Tcm/em (module 2)", hjust=0.5, y=0.98, size=30)
p13 <- PlotCoexpression(features=c("score_GEP4", "score_Terekhova_CD8_Tcm_CCR4pos")) +
  draw_label("CD8 Tcm CCR4+", hjust=0.5, y=0.98, size=30)
p14 <- PlotCoexpression(features=c("score_GEP4", "score_Poon_CD4TCM_TFH")) +
  draw_label("CD4 Tcm/fh", hjust=0.5, y=0.98, size=30)


# GEP5
p21 <- PlotCoexpression(features=c("score_GEP5", "score_CanoGamez_CD4_TEM")) +
  draw_label("CD4 Tem", hjust=0.5, y=0.98, size=30)
p22 <- PlotCoexpression(features=c("score_GEP5", "score_Rose_CD4_modul4_Tem")) +
  draw_label("CD4 Tem (module 4)", hjust=0.5, y=0.98, size=30)
p23 <- PlotCoexpression(features=c("score_GEP5", "score_Terekhova_MAITcells")) +
  draw_label("MAIT", hjust=0.5, y=0.98, size=30)
p24 <- PlotCoexpression(features=c("score_GEP5", "score_Poon_CD8MAIT")) +
  draw_label("CD8 MAIT", hjust=0.5, y=0.98, size=30)


# GEP6
p31 <- PlotCoexpression(features=c("score_GEP6", "score_CanoGamez_CD4_TEMRA")) +
  draw_label("CD4 Temra", hjust=0.5, y=0.98, size=30)
p32 <- PlotCoexpression(features=c("score_GEP6", "score_Rose_CD8_modul1_Temra")) +
  draw_label("CD8 Temra (module 1)", hjust=0.5, y=0.98, size=30)
p33 <- PlotCoexpression(features=c("score_GEP6", "score_Terekhova_CD8_Temra")) +
  draw_label("CD8 Temra", hjust=0.5, y=0.98, size=30)
p34 <- PlotCoexpression(features=c("score_GEP6", "score_Poon_CD8TEM_TEMRA")) +
  draw_label("CD8 Tem/emra", hjust=0.5, y=0.98, size=30)
```

Now combine all the plots.
```{r plot-coexpression-combine, fig.height=20, fig.width=25}
plot_grid(
  plot_grid(p11,p12,p13,p14, nrow = 1, scale=0.95),
  plot_grid(p21,p22,p23,p24, nrow = 1, scale=0.95),
  plot_grid(p31,p32,p33,p34, nrow = 1, scale=0.95),
  nrow=3, scale=0.95
)
```




# SESSION INFO
```{r}
sessionInfo()
```