---
title: "04_CorrelationComparison"
output: html_notebook
---

# IMPORT

```{r import-librairies}
library(ggplot2)
library(tidyverse)
library(corrplot)
library(magrittr)
library(dplyr)
```

```{r import-data}
path <- "~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/02_CorrelationComparion"
# Mouse data
DEgenes_ms.df <- readRDS(file.path(path, "ms_DEG.rds"))
ExpressionTable_ms <- readRDS(file.path(path, "ms_avgexp.rds"))
# Pig data
DEgenes_pig.df <- readRDS(file.path(path, "pig_DEG.rds"))
ExpressionTable_pig <- readRDS(file.path(path, "pig_avgexp.rds"))
# Pig data
DEgenes_hu.df <- readRDS(file.path(path, "hu_DEG.rds"))
ExpressionTable_hu <- readRDS(file.path(path, "hu_avgexp.rds"))

# Import orthologs table
ortholog.df <- read.csv("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/03_BiomartTable/big_ass_ortholog_table.csv")
```




# COMPARATIVE ANALYSIS

Step 1: Take intersect, union, species1 or species2 of DEgenes for analysis

```{r select-DEgenes}
# First, check whether DEgenes can all be found in the ortholog table (many can't be found because I removed genes with orthology confidence=0)
table(unique(DEgenes_ms.df$gene) %in% ortholog.df$ms_symbol_data) # 500 not
table(unique(DEgenes_hu.df$gene) %in% ortholog.df$hu_symbol) # 353 not
table(unique(DEgenes_pig.df$gene) %in% ortholog.df$pig_symbol_data) # 444 not

# Get the nb of human and pig DEgenes
length(unique(DEgenes_ms.df$gene)) # 2448
length(unique(DEgenes_hu.df$gene)) # 1481
length(unique(DEgenes_pig.df$gene)) # 2085


# Second, subset the ortholog table to only genes of interest (3 species)
DEgenes <- ortholog.df %>%
  # Intersection
  # filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) & hu_symbol %in% unique(DEgenes_hu.df$gene) & pig_symbol_data %in% unique(DEgenes_pig.df$gene)) %>%
  # Union
  filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) | hu_symbol %in% unique(DEgenes_hu.df$gene) | pig_symbol_data %in% unique(DEgenes_pig.df$gene)) %>%
  filter(!is.na(ms_symbol_data)) %>%
  # Keep only genes that can be found in the expression tables
  filter(ms_symbol_data %in% rownames(ExpressionTable_ms)) %>%
  filter(hu_symbol %in% rownames(ExpressionTable_hu)) %>%
  filter(pig_symbol_data %in% rownames(ExpressionTable_pig))

# intersect ms-hu-pig: 395 rows
# union ms-hu-pig: 3,319 rows


# Subset the ortholog table to only genes of interest (2 species)
DEgenes <- ortholog.df %>%
  # Intersection
  # filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) & hu_symbol %in% unique(DEgenes_hu.df$gene)) %>%
  filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) & pig_symbol_data %in% unique(DEgenes_pig.df$gene)) %>%
  # Union
  # filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) | hu_symbol %in% unique(DEgenes_hu.df$gene)) %>%
  # filter(ms_symbol_data %in% unique(DEgenes_ms.df$gene) | pig_symbol_data %in% unique(DEgenes_pig.df$gene)) %>%
  filter(!is.na(ms_symbol_data)) %>%
  # filter(hu_orthology_confidence == 1) %>%
  filter(pig_orthology_confidence == 1) %>%
  # Keep only genes that can be found in the expression tables
  filter(ms_symbol_data %in% rownames(ExpressionTable_ms)) %>%
  # filter(hu_symbol %in% rownames(ExpressionTable_hu))
  filter(pig_symbol_data %in% rownames(ExpressionTable_pig))
```


Step 2: Prune Expression Tables & Remove rows with no expression

```{r prune-expression-tables}
# Sanity check
# table(DEgenes$ms_symbol_data %in% rownames(ExpressionTable_ms))
# table(DEgenes$pig_symbol_data %in% rownames(ExpressionTable_pig))
# table(DEgenes$hu_symbol %in% rownames(ExpressionTable_hu))

# Prune expression tables
AvgExp.ms = ExpressionTable_ms[rownames(ExpressionTable_ms) %in% DEgenes$ms_symbol_data,]
AvgExp.ms = AvgExp.ms[rowSums (AvgExp.ms)!=0,]
AvgExp.hu = ExpressionTable_hu[rownames(ExpressionTable_hu) %in% DEgenes$hu_symbol,]
AvgExp.hu = AvgExp.hu[rowSums (AvgExp.hu)!=0,]
AvgExp.pig = ExpressionTable_pig[rownames(ExpressionTable_pig) %in% DEgenes$pig_symbol_data,]
AvgExp.pig = AvgExp.pig[rowSums (AvgExp.pig)!=0,]
```



Step 3: Replace human and pig gene names with mouse gene names

```{r translate-gene-names}
## HUMAN ##
# Add new column with human gene names
AvgExp.hu[,ncol(AvgExp.hu)+1] = rownames(AvgExp.hu)
colnames(AvgExp.hu)[ncol(AvgExp.hu)] = "hu_symbol"
# Add mouse ortholog genes, keeping only the ones from the DEgenes list
AvgExp.hu <- AvgExp.hu %>%
  merge(DEgenes[,c("hu_symbol", "ms_symbol_data")], by="hu_symbol", all.y = F) %>%
  distinct() %>%
  # for any ms gene that may have 2 human orthologs, average expression
  group_by(ms_symbol_data) %>%
  dplyr::select(-hu_symbol) %>%
  summarize_all("mean")
AvgExp.hu <- as.data.frame(AvgExp.hu)
rownames(AvgExp.hu) = AvgExp.hu$ms_symbol_data
AvgExp.hu$ms_symbol_data <- NULL


## PIG ##
# Add new column with human gene names
AvgExp.pig[,ncol(AvgExp.pig)+1] = rownames(AvgExp.pig)
colnames(AvgExp.pig)[ncol(AvgExp.pig)] = "pig_symbol_data"
# Add mouse ortholog genes, keeping only the ones from the DEgenes list
AvgExp.pig <- AvgExp.pig %>%
  merge(DEgenes[,c("pig_symbol_data", "ms_symbol_data")], by="pig_symbol_data", all.y = F) %>%
  distinct() %>%
  # for any ms gene that may have 2 human orthologs, average expression
  group_by(ms_symbol_data) %>%
  dplyr::select(-pig_symbol_data) %>%
  summarize_all("mean")
AvgExp.pig <- as.data.frame(AvgExp.pig)
rownames(AvgExp.pig) = AvgExp.pig$ms_symbol_data
AvgExp.pig$ms_symbol_data <- NULL
```


Step 4: Scale Expression Tables by gene average

```{r scale-expression-tables}
# Ms
avg = rowMeans(AvgExp.ms)
AvgExp.ms = sweep(AvgExp.ms,1,avg,"/")
rm(avg)

# Hu
avg = rowMeans(AvgExp.hu)
AvgExp.hu = sweep(AvgExp.hu,1,avg,"/")
rm(avg)

# Pig
avg = rowMeans(AvgExp.pig)
AvgExp.pig = sweep(AvgExp.pig,1,avg,"/")
rm(avg)
```



Step 5: Merge Expression tables
```{r merge-expression-tables}
# Sanity check
# setdiff(rownames(AvgExp.ms), rownames(AvgExp.hu))
# setdiff(rownames(AvgExp.hu), rownames(AvgExp.ms))
# setdiff(rownames(AvgExp.ms), rownames(AvgExp.pig))
# setdiff(rownames(AvgExp.pig), rownames(AvgExp.ms))

# Add ms/pig in cluster names, and reorder some columns
AvgExp.ms <- AvgExp.ms[, c("Stage0", "iNKTp", "iNKT2", "iNKT17", "iNKT1")]
colnames(AvgExp.ms) <- paste0("ms_", colnames(AvgExp.ms))
colnames(AvgExp.pig) <- paste0("pig_", colnames(AvgExp.pig))
# AvgExp.hu <- AvgExp.hu[, c("hu_iNKT_4", "hu_iNKT_0", "hu_iNKT_1", "hu_iNKT_5", "hu_iNKT_6", "hu_iNKT_2", "hu_iNKT_3")]

# Merge Expression Tables (ms-hu)
geTable = merge(AvgExp.ms,AvgExp.hu, by='row.names', all=F)
rownames(geTable) = geTable$Row.names
geTable = geTable[,2:ncol(geTable)]
# Add pig expression table
geTable = merge(geTable, AvgExp.pig, by='row.names', all=F)
rownames(geTable) = geTable$Row.names
geTable = geTable[,2:ncol(geTable)]
```



Step 6: Correlation

```{r correlation}
#7a:  Correlation
corr.method <- "spearman" # or pearson?
Corr.Coeff.Table = cor(geTable,method=corr.method)


#7b:  Shuffle data
shuffled.cor.list = list()
pb   <- txtProgressBar(1, 100, style=3)
nPermutations <- 1000

# Calculate correlations on shuffled expression table
for (i in 1:nPermutations){
  # set.seed(123)
  # Just for 2 species
  # shuffled = apply(geTable[,1:ncol(AvgExp.ms)],1,sample)
  # shuffled2 = apply(geTable[,(ncol(AvgExp.ms)+1):ncol(geTable)],1,sample)
  # shuffled = cbind(t(shuffled),t(shuffled2)) # bind the 2 species back together
  # For 3 species: split dataframe by species to shuffle by species
  split <- sapply(c("ms", "hu", "pig"), function(x) geTable[startsWith(names(geTable), x)], simplify=FALSE)
  shuffled  = apply(split$ms,  1, sample)
  shuffled2 = apply(split$hu,  1, sample)
  shuffled3 = apply(split$pig, 1, sample)
  shuffled = cbind(t(shuffled),t(shuffled2), t(shuffled3)) # bind the 3 species back together
  # Correlation
  shuffled.cor = cor(shuffled,method=corr.method)
  shuffled.cor.list[[i]] = shuffled.cor
  rm(list=c('shuffled','shuffled2', 'shuffled3', 'shuffled.cor'))
  if ((i %% 100) ==0){
    setTxtProgressBar(pb, (i*100)/nPermutations)
  }
}
  
# Make empty p.value and corr mean matrixes
p.value.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
rownames(p.value.table) = colnames(geTable)
colnames(p.value.table) = colnames(geTable)

# shuffled.mean.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
# rownames(shuffled.mean.table) = colnames(geTable)
# colnames(shuffled.mean.table) = colnames(geTable)

# Get the mean "random" correlation from all permutations, and pvalue comparing the "observed" correlation with the "random" one
a = combn(1:ncol(geTable),2)
for (i in 1:ncol(a)){
  cor.scores = sapply(shuffled.cor.list,"[",a[1,i],a[2,i])
  # shuffled.mean.table[a[1,i],a[2,i]] = mean(cor.scores)
  # shuffled.mean.table[a[2,i],a[1,i]] = mean(cor.scores)
  p.value = mean(abs(cor.scores)>=abs(Corr.Coeff.Table[a[1,i],a[2,i]]))
  p.value.table[a[1,i],a[2,i]] = p.value
  p.value.table[a[2,i],a[1,i]] = p.value
  rm(list=c('cor.scores','p.value'))
  setTxtProgressBar(pb, (i*100)/ncol(a))
}
p.value.table

# neg.log10.p = -log10(p.value.table)
```




Step 7: Overlap in markers
For all pairs of cell-types generate list of genes that are at least 1.5x avg in both cells

```{r overlap-in-markers}
#from above a = combn(1:ncol(geTable),2)
marker.overlap.list = list()
for (i in 1:ncol(a)){
  datasubset = cbind(geTable[,a[1,i]],geTable[,a[2,i]])
  markers = rownames(geTable[datasubset[,1]>1.5 & datasubset[,2]>1.5,])
  marker.overlap.list[[i]] = markers
  names(marker.overlap.list)[i] = paste(colnames(geTable)[a[1,i]], colnames(geTable)[a[2,i]],sep='_')
  rm(list=c('datasubset','markers'))
}
```




Step 8: Plot
```{r plot-correlation}
# comp.intersect <- SpPermute(ExpressionTableSpecies1, species1=Species1, DEgenesSpecies1, ExpressionTableSpecies2, species2=Species2, DEgenesSpecies2, nPermutations=1000, genes.use= Method1,corr.method="spearman")

# Get the correlation and ptable with only mouse x rest
# comp_table <- t(Corr.Coeff.Table[1:ncol(ExpressionTable_ms), (ncol(ExpressionTable_ms)+1):nrow(Corr.Coeff.Table)])
# p_table <- t(p.value.table[1:ncol(ExpressionTable_ms), (ncol(ExpressionTable_ms)+1):nrow(Corr.Coeff.Table)])
# p_table <- (-p_table)

# Get the correlation and ptable with all species
comp_table <- Corr.Coeff.Table
p_table <- p.value.table

# Prepare colors
col1 <- colorRampPalette(c("darkblue", "white","darkred"))
text_col <- c(rep("#762a83", 5), rep("#5aae61", 7), rep("#e08214", 8))
names(text_col) <- colnames(Corr.Coeff.Table)
text_col <- text_col[corrMatOrder(Corr.Coeff.Table, order="hclust", hclust.method="ward.D")]

# Ms-pig
# comp_table <- comp_table[c("pig_iNKT2.0", "pig_iNKT2.2", "pig_iNKTc", "pig_iNKTt", "pig_iNKT2.1", "pig_iNKT2.3", "pig_iNKT-ISG", "pig_iNKT-sw"),]
# p_table <- p_table[c("pig_iNKT2.0", "pig_iNKT2.2", "pig_iNKTc", "pig_iNKTt", "pig_iNKT2.1", "pig_iNKT2.3", "pig_iNKT-ISG", "pig_iNKT-sw"),]

# Plot
title <- "Union between DEgenes"
jpeg("~/Projects/20220809_Thymic-iNKT-CrossSpecies/data/02_CorrelationComparion/all_species_union_hclust.jpeg", width=2000, height=2000, res=300)
corrplot(comp_table,
         # 2 species
         # order="original",
         # tl.col="black",
         # 3 species
         order="hclust",
         hclust.method = "ward.D",
         tl.col=text_col,
         diag=F, # no color in the diagonal
         tl.pos="lt", method="color",
         col=col1(200),
         # cl.lim=c(min(comp_table.intersect),max(comp_table.intersect)),
         is.corr=F, tl.cex=0.7, sig.level=0.05,
         p.mat=p_table, # significance
         # insig="pch", pch=19, pch.cex=0.25, pch.col="black",
         insig="blank",
         addrect = 3,
         main= paste(title,"-",length(rownames(geTable)), "genes", sep=" "),mar=c(3,1,5,1),cl.align.text="l") #%>%
  # corrRect(c(1,6,13, 20))
  # corrRect(namesMat=c("ms_Stage0", "hu_iNKT_4", "ms_iNKT1", "hu_iNKT_3"))
dev.off()

```










