---
title: "TEMPORARY SCRIPT - clean up seurat objects thymic/pbmc lineages"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  root.dir = "~/Projects/HumanThymusProject/"
)
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(Seurat)
library(SCpubr)

source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# integrated object
seur_integrated <- readRDS("./data_geo/seurat_objects/seurat_human_integrated_object_23_12_01.rds")

# thymic objects
seur_thym <- list(
  "cd4"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.cd4.RDS"),
  "cd8"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.cd8.RDS"),
  "nkt"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.nkt.RDS"),
  "mait"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.mait.RDS"),
  "gdt"=readRDS("./data/human-thymus/HumanThymus_23_PlotThymicGEPs/seurat_filtered_harmony_02_15_23_thymus.gd.RDS")
)

# pbmc objects
# seur_pbmc <- list(
#   "cd4"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd4.RDS"),
#   "cd8"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd8.RDS"),
#   "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.nkt.RDS"),
#   "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.mait.RDS"),
#   "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.gd.RDS")
# )
```




# CLEAN UP THYMIC OBJECTS

## Remove useless columns
```{r thym-rm-cols}
# check that column names are the same for all lineages
table(colnames(seur_thym$cd4@meta.data)==colnames(seur_thym$cd8@meta.data), useNA="ifany")
table(colnames(seur_thym$cd4@meta.data)==colnames(seur_thym$nkt@meta.data), useNA="ifany")
table(colnames(seur_thym$cd4@meta.data)==colnames(seur_thym$mait@meta.data), useNA="ifany")
table(colnames(seur_thym$cd4@meta.data)==colnames(seur_thym$gdt@meta.data), useNA="ifany")
table(colnames(seur_thym$nkt@meta.data)==colnames(seur_thym$gdt@meta.data), useNA="ifany")

# the only difference in column names is a clustering resolution (we don't need that column anyway)
colnames(seur_thym$cd4@meta.data)[colnames(seur_thym$cd4@meta.data)!=colnames(seur_thym$nkt@meta.data)]

# check out all metadata columns
print(colnames(seur_thym$cd4@meta.data))

# remove useless columns
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data[,c(13:18, 33:77)] <- NULL
    return(x)
    }
)
```


## Add clustering columns
```{r thym-add-clust-cols}
# sanity check that rownames (cell barcodes) are in same order in integrated object & individual thymic objects
# table(
#   rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$cd4@meta.data),]) == rownames(seur_thym$cd4@meta.data), useNA="ifany"
# )
# table(
#   rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$cd8@meta.data),]) == rownames(seur_thym$cd8@meta.data), useNA="ifany"
# )
# table(
#   rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$nkt@meta.data),]) == rownames(seur_thym$nkt@meta.data), useNA="ifany"
# )
# table(
#   rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$mait@meta.data),]) == rownames(seur_thym$mait@meta.data), useNA="ifany"
# )
# table(
#   rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$gdt@meta.data),]) == rownames(seur_thym$gdt@meta.data), useNA="ifany"
# )

# add column "clusters integrated data"
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data$clusters_integrated_data <- seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(x@meta.data), "clusters_integrated_data"]
    return(x)
    }
  )

# add column "clusters per lineage"
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data$clusters_per_lineage <- seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(x@meta.data), "clusters_per_lineage"]
    return(x)
    }
  )
```

Check that the `clusters_per_lineage` column corresponds to Fig 2.
```{r thym-dim-plots}
lapply(
  seur_thym,
  function(x){
    return(DimPlot(x, group.by = "clusters_per_lineage", label=T))
  }
)
```


## Rename metadata columns
Rename with the same column names used in the clean integrated seurat object (deposited on GEO).
```{r thym-rename-cols}
seur_thym <- lapply(
  seur_thym,
  function(x){
    colnames(x@meta.data)[4:12] <- c(
      "percent_mitochondrial",
      "tcell_lineage",
      "donor_sex",
      "donor_age_weeks",
      "donor_id",
      "batch_id",
      "sequencing_method",
      "tissue",
      "tcell_lineage_tissue"
    )
    return(x)
  }
)
```


## Save seurat objects
```{r thym-save}
saveRDS(seur_thym$cd4, "./data_github/seurat_objects_thymus/seurat_thymus_cd4.rds")
saveRDS(seur_thym$cd8, "./data_github/seurat_objects_thymus/seurat_thymus_cd8.rds")
saveRDS(seur_thym$nkt, "./data_github/seurat_objects_thymus/seurat_thymus_inkt.rds")
saveRDS(seur_thym$mait, "./data_github/seurat_objects_thymus/seurat_thymus_mait.rds")
saveRDS(seur_thym$gdt, "./data_github/seurat_objects_thymus/seurat_thymus_gdt.rds")
```




# CLEAN UP PBMC OBJECTS

## Analysis 3
```{r analysis-3}

```


## Analysis 4
```{r analysis-4}

```




# SESSION INFO
```{r}
sessionInfo()
```


