---
title: "Download and prepare Park Thymic Atlas"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  attr.output='style="max-height: 300px;"',
  root.dir = "~/Projects/HumanThymusProject/"
)
```




# IMPORT LIBRAIRIES
```{r import-librairies}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SingleCellExperiment)
library(scater)
library(SCpubr)

setwd("~/Projects/HumanThymusProject/")
source("~/Projects/HumanThymusProject/scripts/colors_universal.R")
```




# HUMAN THYMUS ATLAS

## Import data
The Human Thymus Atlas was downloaded as a `h5ad` file from [cellxgene](https://cellxgene.cziscience.com/collections/de13e3e2-23b6-40ed-a413-e9e12d7d3910), the one containing 255,901 cells.
```{r import-data-human, fig.height=10, fig.width=15}
# import h5ad object downloaded
sce_human_cellgene <- zellkonverter::readH5AD("~/Projects/HumanThymusProject/data_github/park_dataset/c6e08ab6-ab3b-41dc-8058-8e6442e081ec.h5ad")
print(sce_human_cellgene)

# check out metadata
colData(sce_human_cellgene)

# check umap and cell type annotation
plotReducedDim(sce_human_cellgene, dimred="X_umap", colour_by="cell_type")

# check out gene names
rowData(sce_human_cellgene)
table(duplicated(rowData(sce_human_cellgene)$feature_name), useNA="ifany") # none of the gene symbols are duplicated (ideal :)
```


The cell annotation doesn't correspond to Figure 1 of the [Park et al.](https://www.science.org/doi/10.1126/science.aay3224) paper. We are looking for different annotation levels such as the ones on the [cellatlas.io](https://app.cellatlas.io/thymus-development/dataset/8/scatterplot) website. We downloaded the `h5ad` file from cellatlas.io, let's import it.
```{r import-data-human-2, fig.height=8, fig.width=10}
# import h5ad file from cellatlas.io
sce_human_cellatlas <- zellkonverter::readH5AD("~/Projects/HumanThymusProject/data_github/park_dataset/HTA08_v01_A05_Science_human_fig1.h5ad")
print(sce_human_cellatlas)

# check out metadata
colData(sce_human_cellatlas) # much better!

# check umap and cell type annotation
plotReducedDim(sce_human_cellatlas, dimred="X_umap", colour_by="Anno_level_fig1")

# check out gene names
# rowData(sce_human_cellatlas) # rowData is empty
table(duplicated(rownames(sce_human_cellatlas)), useNA="ifany") # some gene symbols are duplicated (not ideal...)
```

In summary:

- `h5ad` file from [cellxgene](https://cellxgene.cziscience.com/collections/de13e3e2-23b6-40ed-a413-e9e12d7d3910) contains more information on feature names (with unique gene symbols);
- `h5ad` file from [cellatlas.io](https://app.cellatlas.io/thymus-development/dataset/8/scatterplot) contains more/better cell metadata information (with correct clustering annotation).

We will combine all we need into a new `SingleCellExperiment` object, and then convert it into a `Seurat` object.


## Create Seurat object
First, let's create a new `SingleCellExperiment` object.
```{r create-human-sce-object-clean}
# check that the cell IDs are the same from both sources
table(colnames(sce_human_cellgene) == colnames(sce_human_cellatlas), useNA="ifany")
# table(rownames(colData(sce_human_cellgene)) == rownames(colData(sce_human_cellatlas)), useNA="ifany")

# check that umap coordinates are the same from both sources
table(reducedDim(sce_human_cellgene, "X_umap") == reducedDim(sce_human_cellatlas, "X_umap"), useNA="ifany")

# create new SCE object
sce_human_clean <- SingleCellExperiment(
  list(counts=assay(sce_human_cellgene)),
  colData=colData(sce_human_cellatlas),
  rowData=rowData(sce_human_cellgene),
  reducedDims=list(umap=reducedDim(sce_human_cellgene, "X_umap"))
)
print(sce_human_clean)

# switch rownames to gene symbols
# table(duplicated(rowData(sce_human_clean)$feature_name), useNA="ifany") # last sanity check: no duplicates :)
rownames(sce_human_clean) <- rowData(sce_human_clean)$feature_name
```

Let's now convert into a seurat object.
```{r create-human-seurat-object, fig.width=10, fig.height=8}
# convert to seurat
seur_human <- Seurat::as.Seurat(sce_human_clean, data=NULL)

# remove genes which have 0 total count
seur_human <- seur_human[rowSums(seur_human)!=0,]

SCpubr::do_DimPlot(
  seur_human,
  group.by="Anno_level_fig1",
  legend.position="right"
)
```


## Save Seurat object
```{r seur-human-save, eval=FALSE, echo=TRUE}
saveRDS(seur_human, "./data_github/park_dataset/park_seurat_human.rds")
```




# MURINE THYMUS ATLAS

## Import data
The `h5ad` file was downloaded from the latest version of the [zenodo repository](https://zenodo.org/records/5500511) of the paper (v1.0.2). Let's import it as a `SingleCellExperiment` object once again, and convert it to a `Seurat` object.
```{r import-data-mouse, fig.height=8, fig.width=10}
sce_mouse <- zellkonverter::readH5AD("~/Projects/HumanThymusProject/data_github/park_dataset/HTA08_v02_A04_Science_mouse_total.h5ad")
print(sce_mouse)

# check out metadata
colData(sce_mouse)

# check umap and cell type annotation
plotReducedDim(sce_mouse, dimred="X_umap", colour_by="cell.types")

# check out gene names
table(duplicated(rownames(sce_mouse)), useNA="ifany") # none of the gene symbols are duplicated (ideal :)
```


## Create Seurat object
```{r create-seurat-object-mouse, fig.width=10, fig.height=8}
# create new SCE object
sce_mouse_clean <- SingleCellExperiment(
  list(counts=assay(sce_mouse)),
  colData=colData(sce_mouse),
  rowData=rowData(sce_mouse),
  reducedDims=list(umap=reducedDim(sce_mouse, "X_umap"))
)
print(sce_mouse_clean)

# convert to seurat
seur_mouse <- Seurat::as.Seurat(sce_mouse_clean, data=NULL)

# remove genes which have 0 total count
seur_mouse <- seur_mouse[rowSums(seur_mouse)!=0,]

# plot cell annotation
SCpubr::do_DimPlot(
  seur_mouse,
  group.by="cell.types",
  legend.position="right"
)
```


## Save Seurat object
```{r seur-mouse-save, eval=FALSE, echo=TRUE}
saveRDS(seur_mouse, "./data_github/park_dataset/park_seurat_mouse.rds")
```




# HUMAN THYMUS ATLAS (THYMOCYTES ONLY)

## Import data
The Human Thymus Atlas with only thymocytes (corresponding to Figure 2C in [Park et al.](https://www.science.org/doi/10.1126/science.aay3224) paper) was downloaded from two sources as a `h5ad` file:

- [cellxgene](https://cellxgene.cziscience.com/collections/de13e3e2-23b6-40ed-a413-e9e12d7d3910), the one containing 76,994 cells;
- [cellatlas.io](https://app.cellatlas.io/thymus-development/dataset/10/scatterplot).

Once again, we will compare them and see which is more useful for our purposes. First, `cellxgene`:
```{r import-data-human-thymocytes-cellgene, fig.height=10, fig.width=15}
# import h5ad object downloaded
sce_human_thymocytes_cellgene <- zellkonverter::readH5AD("~/Projects/HumanThymusProject/data_github/park_dataset/de8665a2-0476-4865-b4af-c7b8d3b1b87f.h5ad")
print(sce_human_thymocytes_cellgene)

# check out metadata
colData(sce_human_thymocytes_cellgene)

# check umap and cell type annotation
plotReducedDim(sce_human_thymocytes_cellgene, dimred="X_umap", colour_by="cell_type")

# check out gene names
rowData(sce_human_thymocytes_cellgene)
table(duplicated(rowData(sce_human_thymocytes_cellgene)$feature_name), useNA="ifany") # none of the gene symbols are duplicated (ideal :)
```

The cell annotation is not ideal, however the gene symbols are not duplicated, which is better. Second, data from `cellatlas.io`:
```{r import-data-human-thymocytes-cellatlasio, fig.height=10, fig.width=15}
# import h5ad object downloaded
sce_human_thymocytes_cellatlas <- zellkonverter::readH5AD("~/Projects/HumanThymusProject/data_github/park_dataset/HTA08_v01_A06_Science_human_tcells.h5ad")
print(sce_human_thymocytes_cellatlas)

# check out metadata
colData(sce_human_thymocytes_cellatlas)

# check umap and cell type annotation
plotReducedDim(sce_human_thymocytes_cellatlas, dimred="X_umap", colour_by="cell.types")

# check out gene names
rowData(sce_human_thymocytes_cellatlas)
rownames(sce_human_thymocytes_cellatlas)[1:5]
table(duplicated(rownames(sce_human_thymocytes_cellatlas)), useNA="ifany") # some of the gene symbols are duplicated (not great)
```

In summary, similarly to the entire Human Thymus Atlas (which we previously processed in this notebook):

- `h5ad` file from [cellxgene](https://cellxgene.cziscience.com/collections/de13e3e2-23b6-40ed-a413-e9e12d7d3910) contains more information on feature names (with unique gene symbols);
- `h5ad` file from [cellatlas.io](https://app.cellatlas.io/thymus-development/dataset/10/scatterplot) contains more/better cell metadata information (with correct clustering annotation).

We will combine all we need into a new `SingleCellExperiment` object, and then convert it into a `Seurat` object.


## Create Seurat object
First, let's create a new `SingleCellExperiment` object.
```{r create-human-thymocytes-sce-object-clean}
# check that the cell IDs are the same from both sources
table(colnames(sce_human_thymocytes_cellgene) == colnames(sce_human_thymocytes_cellatlas), useNA="ifany")
# table(rownames(colData(sce_human_thymocytes_cellgene)) == rownames(colData(sce_human_thymocytes_cellatlas)), useNA="ifany")

# check that umap coordinates are the same from both sources
table(reducedDim(sce_human_thymocytes_cellgene, "X_umap") == reducedDim(sce_human_thymocytes_cellatlas, "X_umap"), useNA="ifany")

# create new SCE object
sce_human_thymocytes_clean <- SingleCellExperiment(
  list(counts=assay(sce_human_thymocytes_cellgene)),
  colData=colData(sce_human_thymocytes_cellatlas),
  rowData=rowData(sce_human_thymocytes_cellgene),
  reducedDims=list(umap=reducedDim(sce_human_thymocytes_cellgene, "X_umap"))
)
print(sce_human_thymocytes_clean)

# switch rownames to gene symbols
# table(duplicated(rowData(sce_human_thymocytes_clean)$feature_name), useNA="ifany") # last sanity check: no duplicates :)
rownames(sce_human_thymocytes_clean) <- rowData(sce_human_thymocytes_clean)$feature_name

# rename "cell.types" column to "cell_types"
colnames(colData(sce_human_thymocytes_clean))[13] <- "cell_types"
```

Let's now convert into a seurat object.
```{r create-human-thymocytes-seurat-object, fig.width=10, fig.height=8}
# convert to seurat
seur_human_thymocytes <- Seurat::as.Seurat(sce_human_thymocytes_clean, data=NULL)

# remove genes which have 0 total count
seur_human_thymocytes <- seur_human_thymocytes[rowSums(seur_human_thymocytes)!=0,]

SCpubr::do_DimPlot(
  seur_human_thymocytes,
  group.by="cell_types",
  legend.position="right",
  colors.use = cols_park,
  label=T
)
```


## Save Seurat object
```{r seur-human-thymocytes-save, eval=FALSE, echo=TRUE}
saveRDS(seur_human_thymocytes, "./data_github/park_dataset/park_seurat_human_thymocytes.rds")
```




# SESSION INFO
```{r}
sessionInfo()
```